\documentclass[11pt]{article}

% Machine-readable / crawler-friendly PDF settings
\usepackage{cmap}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\input{glyphtounicode.tex}
\pdfgentounicode=1
\usepackage{microtype}
\hyphenpenalty=700
\exhyphenpenalty=700
\emergencystretch=1.5em
\usepackage{float}

% Core math / layout
\usepackage[margin=0.8in]{geometry}
\usepackage{setspace}
\setstretch{1.15}
\usepackage{amsmath,amssymb,amsthm,mathtools,bm,mathrsfs}
\usepackage{booktabs,longtable,tabularx,array}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{verbatim}
\usepackage{hyperref}

\hypersetup{
  pdftitle={Recursive Self-Improvement Stability under Endogenous Yardstick Drift},
  pdfauthor={K. Takahashi},
  pdfsubject={A first-principles interface theory of improvement, evaluation, and verification coupling conditional on replayably published infrastructure and instrumentation},
  pdfkeywords={recursive self-improvement, endogenous yardstick drift, evaluator drift, verification backlog, benchmark decay, no-meta governance, autonomous agents, model-architecture-agnostic stability, replayable interfaces, challenge lane, heterodoxy preservation},
  pdfcreator={LaTeX},
  pdfproducer={LaTeX},
  colorlinks=true,
  linkcolor=blue,
  citecolor=blue,
  urlcolor=blue,
  pdfstartview=FitH,
  pdflang={en-US}
}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{lemma}[theorem]{Lemma}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{example}[theorem]{Example}

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\Prob}{\mathbb{P}}
\newcommand{\Var}{\mathrm{Var}}
\newcommand{\Cov}{\mathrm{Cov}}
\newcommand{\1}{\mathbf{1}}
\newcommand{\eps}{\varepsilon}
\newcommand{\norm}[1]{\left\lVert #1 \right\rVert}
\newcommand{\abs}[1]{\left\lvert #1 \right\rvert}
\newcommand{\set}[1]{\left\{ #1 \right\}}
\newcommand{\paren}[1]{\left( #1 \right)}
\newcommand{\brak}[1]{\left[ #1 \right]}
\newcommand{\Defeq}{\vcentcolon=}
\newcommand{\Core}{\mathsf{Core}}
\newcommand{\Het}{\mathsf{Het}}
\newcommand{\Narr}{\mathsf{Narr}}
\newcommand{\Challenge}{\mathsf{Challenge}}
\newcommand{\Shadow}{\mathsf{Shadow}}
\newcommand{\Obs}{\mathsf{Obs}}
\newcommand{\Replay}{\mathsf{Replay}}
\newcommand{\Promote}{\mathsf{Promote}}
\newcommand{\Demote}{\mathsf{Demote}}
\newcommand{\ReplayCF}{\operatorname{ReplayCF}}

\title{\textbf{Recursive Self-Improvement Stability under Endogenous Yardstick Drift}\\
\large A First-Principles Interface Theory of Improvement, Evaluation, and Verification Coupling under Replayably Published Instrumentation}
\author{K. Takahashi\\\footnotesize\href{https://orcid.org/0009-0004-4273-3365}{ORCID: 0009-0004-4273-3365}}
\date{March 16, 2026}

\begin{document}
\maketitle

\begin{abstract}
Recursive self-improvement is hardest precisely when the object being improved also edits its own ruler. This paper develops a stability layer for that setting under endogenous yardstick drift, delayed audit, contradiction-preserving memory, protected governance lanes, and explicit resource escrow. The formal core is organized around replayable counterfactual comparison under typed exogenous noise, lower-oracle stable-gain intervals for ruler-editing proposals, delayed shadow certification, contradiction obligations, typed memory lanes, replay and refreezing margins, unresolved-downside debt reserves, and capability-dependent verification throughput. To improve both readability and machine-readability, the paper separates four interfaces that are easy to conflate in self-editing systems: base admissibility, certified stable gain, debt-bounded stability, and governance safety. It also separates exact comparator-compatible proposals from approximate comparator-compatible proposals with replayable distortion budgets, and it separates oracle interval objects from the replayable lower certificates that are actually allowed to enter admission.

The main results are grouped around six core mechanisms rather than a flat inventory. First, typed-noise comparator certificates and replayable $\varepsilon$-bisimulation style state lifts make stable gain mathematically well defined even when evaluator, parser, routing, or policy objects are themselves edited; exact comparator compatibility appears as the zero-distortion special case. Second, latent-scaffold and open-ended small-gain theorems are supplemented by constructive derivation results showing how contractive window envelopes and window-summable spillover envelopes can be obtained from local mode libraries, dwell constraints, admission quotas, and protected-lane workload caps rather than treated as unexplained gifts. Third, delayed shadow certification is strengthened from a calibration wrapper into a logged statistical design: randomized challenge routing with logged propensities, transport covariates, and replayable confidence-sequence lower bounds turns retained shadow margins into finite-sample stable-gain certificates rather than informal delayed scores. Fourth, public conservative envelopes are stratified into direct metered observables, proof-carrying derived quantities, and routing-only heuristics, with a gate discipline forbidding heuristic Tier-3 quantities from driving fast deployment by themselves. Fifth, contradiction governance is strengthened from storage preservation to persistent contradiction obligations, protected random reserve exposure, and a Schur-complement cross-basis semantic-volume law; this blocks de-tagging by evaluator whim, aggregate attentional bypass, and mass-only dilution attacks unless every basis is simultaneously captured. Sixth, the positive results are reorganized around admissibility-operator monotonicity under fixed thresholds and then lifted to explicit regime-policy-pair comparison theorems with replayable witness selectors, quota-lottery protected scheduling, bounded-exposure entry control, proof-carrying escalation, and resource-coupled service scaling. Under those explicit observable conditions, stronger replay, challenge, contradiction, promotion, proof, and verification slack enlarges the certifiable productive frontier, governance edits with zero immediate productive increment can still expand future admissible frontiers, delayed challenge and richer-proof lanes can be turned into bounded-delay deployment paths rather than graveyards, and the same logic can be restated after pricing governance-resource cost.

The framework is self-contained only as a first-principles \emph{interface layer} conditional on published replayable instrumentation and conservative envelope objects: it does not assume a privileged overseer, a fixed benchmark, or any particular model architecture or training rule, but it does assume replayable interfaces through which governance-relevant quantities are exposed. It is therefore not infrastructure-free. Replayable interfaces, protected service lanes, proof checkers, attestation roots, contradiction-obligation logs, published transport and calibration envelopes, and resource escrows are part of the modeled object rather than implementation garnish. The resulting stance is growth-promoting only on explicit observable sufficient-condition classes and explicit regime-policy pairs, and the strongest realized-path claims are comparison-contract corollaries built on the operator-monotonicity layer rather than unconditional laws of recursive improvement. When those interfaces fail, the failure mode is not mysterious: the loop loses identifiability, loses audit mass, loses deployment liveness, loses protected budget, or loses throughput before it loses rhetoric.
\end{abstract}

\paragraph{Machine-readable structure.}
A single canonical machine-readable appendix is included at the end of this source file. The mathematical statements in the main text are authoritative; the appendix mirrors anchors, dependencies, notation, and operational lineage for parser convenience and audit reproducibility.

\section{Introduction}
Recursive self-improvement is often described as repeated self-modification. That description is too weak. Practical recursive agents do not merely change policies or code. They also update retrieval stores, parser registries, benchmark definitions, summarizers, and challenge procedures. The object being improved and the ruler used to score improvement can therefore move together. This paper studies the resulting closed-loop stability problem.

The key question is not whether self-modification can ever help. It is whether a recursively self-editing system can continue to distinguish \emph{claimed improvement} from \emph{stable improvement} once the evaluator, benchmark validity, memory substrate, and verification queue all evolve endogenously. The present paper gives a first-principles \emph{interface theory} for that question: model-architecture-neutral, but explicitly conditional on replayable instrumentation, protected service lanes, and conservative published envelopes that expose the quantities used by admission and audit. It assumes no specific architecture, optimizer, or training rule. It only assumes that the system can propose edits, that some replayable evaluation mechanism exists, and that both the improved object and the evaluator can drift over time.

This layer sits between several existing literatures. Surveys of self-evolving agents organize the field around what is evolved, when evolution occurs, and how evolution is implemented \cite{SelfEvolvingSurvey2025}. Classical work such as the G\"odel machine analyzes self-rewrite under internal proof obligations \cite{Schmidhuber2007}, while logical induction studies bounded self-reference in belief management \cite{Garrabrant2016}. On the stability side, classical small-gain and input-to-state-stability results for interconnected nonlinear and discrete-time systems provide generic wrappers for coupled dynamics \cite{JiangTeelPraly1994,JiangWang2001}. The author's recent preprints have separately treated local self-modification and error debt, search stability under finite context, benchmark decay under recursive corpora, observable-only metrology, and governance of shared epistemic commons \cite{TakahashiPV2026,TakahashiSearch2026,TakahashiBH2026,TakahashiMTE2026,TakahashiSEC2026}. These are crucial adjacent layers, but none by itself gives a stand-alone stability law for the case in which the self-improver changes its own ruler under delayed challenge, typed contradiction preservation, and proposal-local debt accounting.

The present paper therefore targets a distinct object: recursive self-improvement stability under endogenous yardstick drift. The central claim is simple. Long-run viability depends not on local improvement gain alone, but on the coupled gains of the improver, evaluator drift, memory distortion, contradiction preservation, and verification lag. If the ruler moves faster than it can be recalibrated, the system can accumulate positive-looking change claims while silently increasing latent error debt. The complementary positive claim is that better governance capital --- stronger replay, challenge, promotion, proof, contradiction, and verification interfaces under fixed thresholds --- enlarges the set of productive upgrades that can be admitted and reused without destabilizing the loop, that some governance edits are therefore frontier-expanding investments rather than overhead, and that under explicit resource pricing the same conclusion can be restated in net-growth terms. The paper further argues that preserving bounded contradiction is itself a public good for recursively self-improving societies and that shared evaluators create stochastic congestion effects absent from single-agent analyses.

\paragraph{Contributions.} The paper contributes sixteen tightly connected ingredients.
\begin{enumerate}[leftmargin=1.8em]
  \item A self-contained formalization of recursive self-improvement with explicit capability state, evaluator state, memory substrate, verification backlog, latent slow state, cohort-escrow error debt, and proposal-local systemic risk bonds derived from structural blast radius.
  \item A canonical admissibility instance, an explicit observation interface, and a filtration-level public state that make the framework operational without relying on a privileged semantic judge.
  \item A hidden-state-aware matrix small-gain theorem with an explicit vector input-to-state bound for the coupled loop, together with an explicit yardstick-drift ceiling once reference drift is bounded, plus an incremental open-ended variant that permits unbounded capability growth while keeping governance variables bounded.
  \item A one-sided shadow-certification bridge linking delayed retained margins to discounted counterfactual stable gain under lower calibration envelopes, generalized half-life crossing laws under monotone drift and backlog envelopes, a shadow speed limit, and a two-clock viability band for stale-versus-captured shadows.
  \item An observational no-go theorem and a quantitative shadow-identifiability result for evaluator-edit proposals, further sharpened into an adversarial cross-examination theorem that derives integrity and semantic-retention envelopes from provenance-separated audit families and identifies an epistemic singularity once attack reach dominates audit diversity.
  \item A contradiction-governance package: bounded contradiction windows, tagged contradiction reserve conservation under admissible lane retyping, typed memory lanes, a heterodoxy-preserving theorem, an asymmetric contradiction underprovision game, a cross-basis semantic-volume law that blocks heterodoxy spam and single-encoder representation capture, and a forced-attention routing invariant that prevents archived contradiction from being silently ignored at evaluation time.
  \item Proposal-local discounted debt budgets, observable debt certificates, a systemic-risk-bond theorem that blocks Trojan under-declaration, and a strict-conservation specialization that separate verification throughput from repair throughput rather than conflating them.
  \item A Red-Queen verification law in which effective service rate decreases with capability-dependent verification complexity, showing that open-ended growth requires resource scaling fast enough to keep verification slack from collapsing, together with a fixed-threshold governance-capital comparison theorem proving that stronger replay, challenge, reserve, and escrow slack expands the set of productive upgrades that can be admitted without destabilizing the loop, without appealing to threshold relaxation.
  \item A causal-information strengthening of the systemic-risk-bond gate, proving that purely topological module counts are spoofable under aggressive refactoring while replayable intervention-capacity envelopes remain representation- and model-architecture-agnostic and therefore robust to topological obfuscation.
  \item A proof-carrying governance law that separates expensive governance-proof generation from cheap public verification, together with a decidability-bounded constitutional proof layer that blocks proof-language capture, makes proof-level denial-of-service visible as a policy boundary rather than a surprise runtime failure, and clarifies how proposals beyond the current proof language can be deferred or escalated rather than misclassified as anti-growth impossibilities.
  \item A steganographic-scrubbing law that adds information-bottlenecked, lossy audit inputs to writable-scope separation, so covert triggers in the raw proposal stream cannot silently synchronize capture across otherwise disjoint shadow families.
  \item A physical resource-escrow invariant tying the verification budget to a hardware- or allocator-level partition outside writable software scope, and a verifiable-transloading law showing how that invariant can survive hardware refresh without silently collapsing during migration.
  \item An entropy-isolation law for steganographic scrubbers, making covert-channel resistance depend not only on lossy interfaces but also on attacker-conditioned min-entropy in the scrubber randomness itself.
  \item A thermodynamic proof-escrow law that augments decidability-bounded constitutional checking with hardware-metered wall-clock and energy preemption, closing the gap between algorithmic step bounds and physical verification budgets.
  \item An epistemic conservation law for garbage collection under drifting yardsticks, requiring that stale contradiction may be dropped only after equal-or-greater native contradiction mass has been regenerated in the current ontology.
  \item A bounded-delay selection-and-promotion liveness theorem together with endogenous cap-maintaining intake control, bounded-exposure entry control, an intertemporal governance-investment theorem, a comparison-sound surrogate-selector theorem, and a policy-coupled realized-path dominance corollary: positively certified challenged upgrades need not die in quarantine, governance edits with zero immediate productive increment can enlarge future admissible frontiers, exact one-step frontier maximization can be replaced by replayable witness-dominating lower-envelope selectors, and under explicit sanitized-menu and transition conditions the slow lane becomes a reusable deployment path whose successful governance investments translate into larger realized productive trajectories rather than mere static compatibility. This deployment logic is illustrated in two solved toy models --- a scalar single-agent loop and a stochastic shared-yardstick multi-agent loop with agent-dependent delayed challenge policies, cohort-escrow debt, endogenous integrity envelopes, and governance workload accounting --- that make the abstract thresholds concrete while blocking cross-proposal masking, separating identifiability from congestion control, and accommodating covariance-envelope rather than only pairwise-uncorrelated shadow noise.
\end{enumerate}

\section{Positioning and Related Work}
\subsection{Self-evolving agents}
Recent surveys give broad maps of self-evolving or self-improving agent systems \cite{SelfEvolvingSurvey2025}. They clarify design dimensions such as what to evolve, when to evolve, and how to evolve. That literature is valuable and timely. But its main output is a taxonomy and design space, not a first-principles stability law for the case in which the evaluator and the improved object co-evolve.

\subsection{Classical provable self-modification and stability wrappers}
The G\"odel machine remains a foundational idealization of provably beneficial self-rewrite \cite{Schmidhuber2007}. Logical induction studies bounded agents that learn to manage self-referential uncertainty \cite{Garrabrant2016}. In parallel, classical small-gain and input-to-state-stability results show how coupled nonlinear or discrete-time subsystems can remain stable when their interconnection gains stay below critical thresholds \cite{JiangTeelPraly1994,JiangWang2001}. Incremental-stability work sharpens that perspective by asking when nearby trajectories remain replayably comparable under common inputs rather than merely bounded in isolation \cite{Angeli2002,Angeli2009}. The switched and hybrid systems literature further clarifies that time-varying mode sequences can be stable even when local steps are not uniformly contractive, provided that dwell-time or windowed product conditions hold \cite{LiberzonMorse1999,HespanhaMorse1999ADT,ShortenWirth2007,LinAntsaklis2009}. The ruler-editing problem is also a proxy-control problem: once benchmark scores are optimized by agents that can alter the benchmark interface, Goodhart-type proxy failures and dataset/covariate-shift pathologies become endogenous rather than merely exogenous deployment noise \cite{ManheimGarrabrant2018,DatasetShift2008}. This is why the present delayed-challenge layer treats support drift, transport mismatch, and calibration failure as first-class defects rather than as afterthought diagnostics, in line with recent shift-aware conformal and doubly robust calibration work \cite{Tibshirani2019,Yang2024,Jonkers2024,JinCandes2026}. These works are conceptually essential, but they do not directly solve the modern engineering problem of a finite-budget agent that edits its benchmark, parser family, retrieval substrate, and challenge procedure under delay. The present paper therefore uses the small-gain viewpoint only as a stability wrapper, then adds the ingredients absent from the classical control setting: endogenous evaluator drift, delayed shadow certification, typed contradiction preservation, replay compatibility under parser/environment drift, and proposal-local debt accounting.

\subsection{Adjacent layers in the author's prior work}
Proposal--Veto Balance studies local self-modification and error debt \cite{TakahashiPV2026}. Search Stability studies adequacy preservation under finite active context and delayed verification \cite{TakahashiSearch2026}. AI Benchmark Half-Life studies validity decay in recursive corpora \cite{TakahashiBH2026}. MTE treats what counts as credit-bearing progress under observable-only metrology \cite{TakahashiMTE2026}. Sovereign Epistemic Commons treats admit/quarantine/forget laws for shared knowledge substrates \cite{TakahashiSEC2026}. The present paper borrows intuition from those layers but remains self-contained. Its distinct object is the \emph{stability of the recursive improvement loop itself}. It also sits closer to practical software-maintenance and model-supply-chain concerns than classical self-modification work: change-impact analysis already treats software edits as graph-local actions with potentially system-wide ripple effects \cite{AcharyaRobinson2011,PalmskogCelikGligoric2020}; bug-fixing actions can themselves introduce new faults \cite{Yin2011IncorrectFixes,MaesBermejo2024}, long-horizon replayability is fragile unless build environments and parsers are explicitly preserved or refrozen \cite{Matthews2010Preservation,Butler2023RB,Malka2024SpaceTime,Perkel2020TenYear,DiCosmoZacchiroli2017}, evaluator pipelines can be covertly captured while preserving benign-looking accuracy, as illustrated by the backdoor and model-supply-chain literature \cite{Gu2019BadNets,Li2024BackdoorSurvey}, design diversity helps but does not magically grant independent failure modes \cite{Avizienis1985,KnightLeveson1986}, and diversity-aware log-determinant geometry provides a principled way to measure whether a contradiction reserve contains genuinely different evidence rather than many near-copies \cite{KuleszaTaskar2012}. The present analysis additionally draws on safe anytime-valid inference and adaptive off-policy evaluation for the delayed-challenge certification layer \cite{HowardRamdasMcAuliffeSekhon2021,RamdasGrunwaldVovkShafer2023,ZhanHadadHirshbergAthey2021,WassermanRamdasBalakrishnan2020}, on algorithmic-causal notions of intervention capacity to make risk bonds less representation-dependent \cite{Pearl2009,JanzingSchoelkopf2010}, on proof-carrying verification and foundational proof-carrying code to shift expensive governance work toward the proposal producer while keeping the trusted checker small \cite{Necula1997,Appel2001}, and on observable-only hybrid proof/replay gating with proof-carrying compilation, positive fallback unlocks, and bounded liveness; these ingredients are used here only as proof-interface and escalation design patterns, not as a substitute for the present stability analysis. The paper also draws on information-bottleneck style lossy interfaces when independent auditors must be protected from covert shared triggers \cite{Tishby1999,CoverThomas2006}, and on trusted-execution and layered remote-attestation mechanisms to state resource escrow and hardware migration as explicit physical invariants rather than accidental software conventions \cite{ShepherdMarkantonakis2024,RATS2023}. It further draws on entropy-source validation for random-bit generation and extractor-based post-processing \cite{NIST90B2018,BarakEtAl2011LHL}, on worst-case-execution-time analysis for microarchitecturally complex processors \cite{Wilhelm2008WCET}, and on timing-side-channel work showing how small logical computations can still induce large physical timing differentials \cite{Kocher1996Timing}; these literatures motivate separating scrubber entropy from writable software state and separating constitutional proof-step bounds from thermodynamically metered verification cost.

\subsection{Separation of concern}
This paper does \emph{not} claim to be a full theory of persistent semantics, parser correctness, constitutional self-revision, liberty under drift, or shared-commons governance. Those are lower or adjacent layers. The present layer is narrower: when does a recursive improver remain stably beneficial once its own ruler can move?

\section{Formal Setup}
\subsection{Closed-loop state, observable history, and policies}
\begin{definition}[Closed-loop state and public policy state]
At time $t\in\N$, the recursive self-improvement loop is described by
\[
z_t=(x_t,e_t,m_t,q_t,r_t,h_t,\Pi_t),
\]
where $x_t\in\R^{d_x}$ is the capability or policy state, $e_t\in\R^{d_e}$ is the evaluator state, $m_t\in\R^{d_m}$ is the memory and retrieval substrate state, $q_t\in\R_+$ is the verification backlog, $r_t\in\R_+$ is latent error debt, $h_t\in\mathcal H$ is an environment-side latent state not necessarily visible to the fast evaluator, and $\Pi_t$ is the declared public policy state collecting the active parser registry, replay policy, lane policy, challenge-routing policy, and other policy-level manifests that determine which future comparisons are admissible. A self-edit action is denoted by $u_t\in\mathcal U_t$. The public policy state used by admissibility rules is
\[
C_t\Defeq (x_t,e_t,m_t,q_t,r_t,H_t^{\Obs},\Pi_t).
\]
The coordinates $(x_t,e_t,m_t,q_t,r_t)$ appearing in $C_t$ are not arbitrary hidden internals. Each must be either directly observable, a logged summary emitted by policy, or a deterministic replay reconstruction from previously logged artifacts and declared transition rules. Any finer state that is neither public nor replay-reconstructible is absorbed into $h_t$ and cannot be used directly by admissibility or certification rules. Thus the formal state is closed under endogenous ruler motion rather than treating the active policy manifests as an external side table.
\end{definition}

\begin{definition}[Observable history and filtration]
Time is ordered as follows. At the beginning of period $t$, the public state $C_t$ is available and the proposal $u_t$ must be chosen using only that information. After the action is issued and period-$t$ visible randomness is realized, the system emits the replay-visible public record
\[
o_{t+1}=\Obs(z_t,u_t,\xi_t),
\]
where $\xi_t$ collects visible randomness and metadata. The observable history available at decision time $t$ is therefore
\[
H_t^{\Obs}\Defeq (o_1,o_2,\dots,o_t),
\]
with $H_0^{\Obs}$ empty. Because each $C_s$ already contains the accumulated public observation state $H_s^{\Obs}$, the pre-decision public filtration is
\[
\mathcal F_t\Defeq \sigma\!\bigl(C_0,\dots,C_t\bigr).
\]
An online decision rule at time $t$ is any $\mathcal F_t$-measurable map. The post-decision, post-record sigma-field is
\[
\mathcal F_t^{+}\Defeq \sigma(\mathcal F_t,o_{t+1}),
\]
so current-period actions may affect $o_{t+1}$ but not the information set used to choose $u_t$.
\end{definition}

\begin{definition}[System dynamics]
The recursive loop is modeled by
\begin{align}
  x_{t+1} &= F_t(x_t,e_t,m_t,q_t,u_t,h_t,\Pi_t,\xi_t), \label{eq:x_dynamics}\\
  e_{t+1} &= G_t(x_t,e_t,m_t,q_t,u_t,h_t,\Pi_t,\zeta_t), \label{eq:e_dynamics}\\
  m_{t+1} &= M_t(x_t,e_t,m_t,q_t,u_t,h_t,\Pi_t,\omega_t), \label{eq:m_dynamics}\\
  q_{t+1} &= \bigl[q_t + \lambda_t^{\mathrm{prop}} - \mu_t^{\mathrm{ver}}\bigr]_+, \label{eq:q_dynamics}\\
  r_{t+1} &= H_t(r_t,x_t,e_t,m_t,q_t,u_t,h_t,\Pi_t,\nu_t), \label{eq:r_dynamics}\\
  h_{t+1} &= U_t(h_t,u_t,\Pi_t,\chi_t), \label{eq:h_dynamics}\\
  \Pi_{t+1} &= P_t(\Pi_t,x_t,e_t,m_t,q_t,r_t,u_t,h_t,\varpi_t). \label{eq:pi_dynamics}
\end{align}
The key feature is that both the evaluator state in \eqref{eq:e_dynamics} and the public policy state in \eqref{eq:pi_dynamics} evolve endogenously. The ruler is therefore part of the state rather than an external constant.
\end{definition}

\subsection{Standing assumptions}
\begin{assumption}[Finite verification capacity]\label{ass:finite_ver}
There exists $\bar\mu<\infty$ such that $0\le \mu_t^{\mathrm{ver}}\le \bar\mu$ for all $t$.
\end{assumption}

\begin{assumption}[No privileged semantic judge]\label{ass:no_meta}
All admit/reject decisions must be representable through a public or replayable admissibility relation. No external privileged semantic oracle is available.
\end{assumption}

\begin{assumption}[Replayable proposals]\label{ass:replayable}
Each self-edit proposal is accompanied by a replay artifact, declared evaluation cost, and version diff.
\end{assumption}

\begin{assumption}[Two-speed verification]\label{ass:two_speed}
A subset of proposals can be routed to a slower challenge process with higher-fidelity evaluation. The challenge process may be delayed but is not directly editable by the same proposal it evaluates.
\end{assumption}

\subsection{Ontology and dependency map}
The paper uses four layers that should not be conflated. Table~\ref{tab:ontology_map} records the intended role of the main object classes.

\begin{table}[H]
\centering
\small
\begin{tabularx}{\textwidth}{>{\raggedright\arraybackslash}p{0.18\textwidth}>{\raggedright\arraybackslash}p{0.2\textwidth}>{\raggedright\arraybackslash}p{0.26\textwidth}X}
\toprule
Layer & Object type & Typical examples & Role in the theory \\
\midrule
Primitive observable & directly logged or replay-reconstructed state & $C_t$, manifests, queue counters, protected budgets, challenge logs & admissibility may depend on these without further transport assumptions \\
Replayably derived conservative envelope & deterministic or proof-carrying function of observables & $K_t^{\mathrm{gov}}$, $B_t^{\mathrm{gov}}$, $\Gamma_t^{\mathrm{cic}}$, $r_t^{\mathrm{pub}}$, service floors & these are allowed at the gate only when published as conservative lower/upper certificates \\
Operational target & class- or witness-conditional public target defined by the active design & $\mathcal M_t^{\mathrm{chal}}$, $\mathcal M_t^{\mathrm{op}}$, $\Phi_t$, $\Phi_t^{\mathrm{net}}$ & these are admission-facing objects; they need not equal cohort-specific oracle quantities \\
Latent / oracle object & not directly admissible without extra bridges & cohort-specific retained margin $\mathcal M_t(h)$, uncaptured semantic directions, full counterfactual stable gain & these appear only through explicit homogeneity, transport, or witness-comparison assumptions \\
\bottomrule
\end{tabularx}
\caption{Ontology map for the main layers. The positive theorems operate on replayable observables, conservative envelopes, and explicit witness-comparison contracts; they do not license replacing those objects by unlogged latent quantities.}
\label{tab:ontology_map}
\end{table}

\section{Canonical Admissibility Instance}
The theory is abstract, but the admissibility relation cannot remain an empty symbol.

\begin{definition}[Canonical admissibility instance]\label{def:admissibility}
Fix a task or query class $\mathfrak q$ and a verification budget $b$. A proposal $u_t$ is \emph{$(\mathfrak q,b)$-admissible} at time $t$, written
\[
u_t\in\mathcal A_{\mathfrak q,b}(C_t),
\]
if and only if all of the following hold:
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item at least one active parser/evaluator configuration replays the proposal and artifact successfully;
  \item the declared replay-and-check cost is at most $b$;
  \item replay execution produces no undefined side effect on the public state;
  \item any proposal that directly edits evaluator parameters, benchmark definitions, or parser/schema configuration is routed to a challenge lane before unrestricted promotion;
  \item any proposal that merges or summarizes prior evidence must either carry a context-preservation certificate or be typed into a non-evidentiary narrative lane.
\end{enumerate}
\end{definition}

\begin{remark}
This admissibility instance is not claimed to be unique. It is a canonical public-budgeted example that makes the rest of the paper operational.
\end{remark}

\begin{definition}[Operational certificates and margins]\label{def:operational_certificates}
A \emph{context-preservation certificate} for a summary artifact $s$ is a replayable witness bundle proving that the summary retains the task-relevant support set declared by policy, including source identifiers, extraction rules, and omission budget. The \emph{parser-family agreement} of an item $x$ at time $t$ is a public scalar
\[
A_t^{\mathrm{parse}}(x)\in[0,1]
\]
measuring weighted agreement across the active parser family on the item's decoded content and type. The \emph{challenge-lane safety margin}, \emph{promotion-lane safety margin}, and \emph{proof-lane safety margin} are the public slacks
\[
S_t^{\mathrm{chal}}\Defeq \mu_t^{\mathrm{chal}}-\Lambda_t^{\mathrm{chal}},\qquad
S_t^{\mathrm{prom}}\Defeq \mu_t^{\mathrm{prom}}-\Lambda_t^{\mathrm{prom}},\qquad
S_t^{\mathrm{proof}}\Defeq \mu_t^{\mathrm{proof}}-\Lambda_t^{\mathrm{proof}},
\]
where $\mu_t^{\mathrm{chal}},\mu_t^{\mathrm{prom}},\mu_t^{\mathrm{proof}}$ are the effective protected service capacities of the challenge, promotion, and richer-proof lanes, and $\Lambda_t^{\mathrm{chal}},\Lambda_t^{\mathrm{prom}},\Lambda_t^{\mathrm{proof}}$ are their predictable routed loads. Finally, let
\[
r_t^{\mathrm{pub}}\ge 0
\]
denote the \emph{published debt-burden envelope}: a replayably designated public upper envelope on the currently committed unresolved governance debt entering Interface B at epoch $t$ (for example aggregate outstanding debt plus any certified issuance-time carryover bound on the active deployment horizon). The envelope is part of the published regime description rather than the supremum over all admissible upper bounds; lower $r_t^{\mathrm{pub}}$ helps admission. Promotion and proof-routing rules may require lower bounds on $A_t^{\mathrm{parse}}(x)$ and on the relevant lane-safety margins.
\end{definition}

\begin{definition}[Replay compatibility margin, semantic retention, and refreezing rule]\label{def:replay_margin}
For a replay artifact $a$ archived at time $\tau$, let
\[
G_t^{\mathrm{replay}}(a)\in[0,1]
\]
be the public replay-compatibility score of $a$ under the active parser family, build manifest, and environment resolver at time $t\ge \tau$. A policy maintains a \emph{replay floor} $g_{\min}\in(0,1]$ if every contradiction-bearing or evidentiary artifact retained for future audit satisfies
\[
G_t^{\mathrm{replay}}(a)\ge g_{\min}.
\]
When an update replaces $a$ by a fresh replay bundle $a'$, let
\[
S_t^{\mathrm{sem}}(a\rightsquigarrow a')\in[0,1]
\]
be the \emph{public lower envelope} on semantic retention returned by the active replayable audit family for the declared claim, support set, and typing constraints under the active parser/environment regime. The associated translation loss is
\[
\Gamma_t^{\mathrm{sem}}(a\rightsquigarrow a')\Defeq 1-S_t^{\mathrm{sem}}(a\rightsquigarrow a').
\]
A parser or environment update is said to \emph{refreeze} an artifact when it emits a fresh replay bundle $a'$ with
\[
G_{t+1}^{\mathrm{replay}}(a')\ge g_{\min}
\]
and an audited semantic-retention envelope $S_t^{\mathrm{sem}}(a\rightsquigarrow a')\ge s_{\min}$ for some policy floor $s_{\min}\in(0,1]$. Exact lossless meaning preservation is \emph{not} assumed; refreezing restores replayability while exposing translation loss as an explicit audited quantity. Here ``public'' does not mean oracle-given: it means replayable under the currently declared audit family, whose own diversity and capture risk are analyzed later.
\end{definition}

\begin{remark}[No-meta reading of semantic retention]\label{rem:no_meta_semantics}
Throughout the paper, semantic-retention quantities are \emph{audit-family-relative lower envelopes}. They certify retention only relative to the declared replayable probe family, parser family, support-set contract, and typing rules active at audit time. The theory therefore does \emph{not} assume access to oracle semantics. What it assumes is weaker and operational: if the published audit family cannot distinguish two artifacts, the paper does not pretend that an external judge silently can.
\end{remark}

\begin{definition}[Effective evidentiary weight under refreezing]\label{def:effective_weight}
Fix an evidentiary or contradiction-bearing artifact lineage
\[
a_0\rightsquigarrow a_1\rightsquigarrow \cdots \rightsquigarrow a_j\rightsquigarrow \cdots
\]
where each arrow denotes an admissible refreezing event certified under Definition~\ref{def:replay_margin}. Let $W_0(a_0)>0$ be the policy-issued initial evidentiary weight of the root artifact, incorporating any provenance, scope, or contradiction-tag weight attached at issuance. For any descendant artifact $a_j$ active at time $t$, define its \emph{effective evidentiary weight} by
\[
W_t^{\mathrm{eff}}(a_j)\Defeq W_0(a_0)\prod_{\ell=0}^{j-1} S_{t_\ell}^{\mathrm{sem}}(a_\ell\rightsquigarrow a_{\ell+1}),
\]
where $t_\ell$ is the replayable time of the $\ell$-th refreezing step on the lineage. If $a$ has never been refrozen, then $W_t^{\mathrm{eff}}(a)=W_0(a)$. Thus replay-floor preservation and semantic-retention loss are separated: replayability can be restored by refreezing while effective evidentiary force may still decay multiplicatively along the lineage.
\end{definition}

\begin{definition}[Adversarial audit families and endogenous envelopes]\label{def:audit_family}
A \emph{shadow audit family} at time $t$ is a finite set of delayed evaluators
\[
\mathfrak S_t=\{\Shadow^{(1)},\dots,\Shadow^{(M_t)}\}
\]
with replayable manifests, parser/build identities, memory roots, and challenge routes. For each family member $m$ active at the \emph{issuance time} $\tau$ of a challenged cohort, let
\[
\widehat V_{\tau+h}^{\Shadow^{(m)}}(u)
\]
denote the replayable score assigned at horizon $h$ to the coupled counterfactual world in which the issuance-time intervention is $U_\tau=u$ and every later update follows the declared challenge-lane continuation interface, with lane $m$ compared through its replayably transported descendants if the shadow family is later refreshed. Thus each cohort is evaluated against the transported descendants of its \emph{issuance-time} shadow family rather than against whichever extra lanes may have been added later. When a single family-level score is needed, the public canonical aggregate is
\[
\widehat V_{\tau+h}^{\Shadow}(u)\Defeq \operatorname{Agg}^{(\tau)}_{m\le M_\tau}\widehat V_{\tau+h}^{\Shadow^{(m)}}(u),
\]
where $\operatorname{Agg}^{(\tau)}$ is a replayable policy-declared issuance-time aggregator such as a median, trimmed mean, or designated protected-subfamily mean. If $M_\tau=0$, no shadow certificate exists for that cohort and we adopt the conservative conventions
\[
\widehat V_{\tau+h}^{\Shadow}(u)\Defeq 0,
\qquad
\Delta_{\tau,h}^{\mathrm{xs}}\Defeq +\infty.
\]
Likewise a \emph{refreezing audit family}
\[
\mathfrak R_t=\{\Replay^{(1)},\dots,\Replay^{(N_t)}\}
\]
returns semantic-retention scores for candidate refreezings. For a challenged cohort issued at time $\tau$ with issued intervention $u_\tau$, define the cross-shadow disagreement radius by
\[
\Delta_{\tau,h}^{\mathrm{xs}}\Defeq
\begin{cases}
\max\limits_{m,m'\le M_\tau}
\left|
\begin{aligned}
&\bigl(\widehat V_{\tau+h}^{\Shadow^{(m)}}(u_\tau)-\widehat V_{\tau+h}^{\Shadow^{(m)}}(0)\bigr)\\
&\qquad-\bigl(\widehat V_{\tau+h}^{\Shadow^{(m')}}(u_\tau)-\widehat V_{\tau+h}^{\Shadow^{(m')}}(0)\bigr)
\end{aligned}
\right|, & M_\tau\ge 1,\\[1.0ex]
+\infty, & M_\tau=0,
\end{cases}
\]
and for a refreezing candidate $a\rightsquigarrow a'$ let
\[
\underline S_t^{\mathrm{sem}}(a\rightsquigarrow a')\Defeq
\begin{cases}
\min\limits_{n\le N_t} S_{t,n}^{\mathrm{sem}}(a\rightsquigarrow a'), & N_t\ge 1,\\[0.8ex]
0, & N_t=0,
\end{cases}
\]
be the endogenous semantic-retention floor across the refreezing family. Let
\[
\Xi_t^{\Shadow}\ge 0,
\qquad
\Xi_t^{\Replay}\ge 0
\]
be public lower envelopes on \emph{audit diversity reserve}, computed from provenance-disjointness, implementation overlap, seed separation, memory-root separation, and declared writable scope. Finally let $A_t^{\mathrm{adv}}\ge0$ be a public upper envelope on the accepted proposal's \emph{manifest-preserving} effective attack reach over the currently active audit family through declared writable or otherwise logged common-cause channels. Whenever a theorem compares attack reach to a family-specific diversity reserve, policy must also publish replayable monotone calibration maps $\phi_t^{\Shadow},\phi_t^{\Replay}:\R_+\to\R_+$ that convert the generic attack-reach envelope into the effective cover units used by the corresponding shadow and replay families. All of these are observable-or-replayable functions of manifests and logs, not oracle constants.
\end{definition}


\begin{remark}[Roles of the audit objects]\label{rem:audit_roles}
The notation is intentionally layered. $\mathfrak S_t$ and $\mathfrak R_t$ denote the active shadow and refreezing \emph{families}; $\mathcal U_t^{\mathrm{aud}}$ denotes the currently retained audit-bearing artifact set; and any policy-level admissibility or stress-test functional built from those objects is then a derived observable. Families certify or challenge artifacts, while the retained audit set is the object whose replayability, effective weight, routing sensitivity, and semantic diversity must remain live. This separation matters because family capture and artifact decay are distinct failure modes.
\end{remark}

\begin{remark}[Constructive proxy templates for public envelopes]\label{rem:proxy_templates}
The paper's public envelopes are intended to be conservative wrappers, not mysterious gifts. A typical implementation computes $\Xi_t^{\Shadow}$ and $\Xi_t^{\Replay}$ as lower bounds on provenance-separated diversity via overlap graphs, writable-scope exclusions, seed-separation checks, and min-cut style disjointness certificates across the active audit families. A typical upper envelope $A_t^{\mathrm{adv}}$ is built from declared writable scope, dependency closure, shared allocator or routing control, and any \emph{manifest-preserving} common-cause channels exposed by the active build and memory roots; covert raw-input channels that are not represented in the public manifest are intentionally excluded and must instead be handled by the scrubber layer below. The family-specific calibrations $\phi_t^{\Shadow}$ and $\phi_t^{\Replay}$ then convert that generic attack envelope into effective common-cause cover units comparable to the reserves $\Xi_t^{\Shadow}$ and $\Xi_t^{\Replay}$. A typical intervention-capacity lower envelope $\Gamma_t^{\mathrm{cic}}(u_t)$ is obtained from conservative intervention-response probes on admissible replays, minimum guaranteed effect on declared governance outputs, or other algorithmic-causal surrogates that are monotone under richer attack surfaces. Likewise the endogenous semantic floor $\underline S_t^{\mathrm{sem}}$ is obtained by taking the minimum certified retention score across a provenance-separated replay family. The theorems therefore apply to whichever conservative proxy family policy actually publishes; the burden is not to guess the true latent quantity, but to expose a replayable envelope whose failure can be audited.
\end{remark}

\begin{proposition}[Slack-robust proxy substitution]\label{prop:proxy_gap}
Fix any admission or governance clause that is coordinatewise nondecreasing in a vector of helpful latent coordinates $u\in\R^p$ and coordinatewise nonincreasing in a vector of burden latent coordinates $b\in\R^q$. Suppose policy publishes replayable proxy coordinates $\hat u,\hat b$ together with replayable error envelopes $\eps^u\in\R_+^p$ and $\eps^b\in\R_+^q$ such that
\[
\hat u\le u\le \hat u+\eps^u,
\qquad
\hat b-\eps^b\le b\le \hat b
\]
coordinatewise. Then any certificate proved using the tightened proxy-side thresholds
\[
\hat u\ge \underline u+\eps^u,
\qquad
\hat b\le \bar b-\eps^b
\]
is sound for the corresponding latent thresholds $u\ge \underline u$ and $b\le \bar b$. Conversely, if the available proxy slack fails to dominate the published error envelopes, the proxy family is insufficient to certify the latent clause by itself.
\end{proposition}

\begin{proof}
From $u\ge \hat u$ and $u\le \hat u+\eps^u$, the tightened helpful-side threshold $\hat u\ge \underline u+\eps^u$ implies $u\ge \underline u$ coordinatewise. From $b\le \hat b$ and $b\ge \hat b-\eps^b$, the tightened burden-side threshold $\hat b\le \bar b-\eps^b$ implies $b\le \bar b$ coordinatewise. Monotonicity of the clause in the helpful and burden coordinates then yields soundness. If the available slack is smaller than the error envelopes, a proxy certificate can no longer guarantee the latent inequalities, so the clause becomes operationally inconclusive rather than certified.
\end{proof}


\begin{definition}[Tiered observable proxies and gate discipline]\label{def:proxy_tiers}
Observable governance proxies are partitioned into three tiers.
\begin{enumerate}[label=(T\arabic*), leftmargin=2em]
  \item \textbf{Tier 1: direct metered observables.} These are directly logged physical or syntactic quantities such as queue lengths, hashes, manifests, wall-clock counters, energy counters, attested allocator shares, and typed lane occupancies.
  \item \textbf{Tier 2: proof-carrying derived quantities.} These are replayably derived from Tier-1 objects together with explicit certificates, for example service lower bounds, replay floors, transport-distortion bounds, verified allocator shares, and challenge-propensity lower bounds.
  \item \textbf{Tier 3: routing heuristics.} These are conservative but model-laden prioritization or search signals such as semantic-similarity alarms, approximate attack-reach heuristics, contradiction novelty scores, or cheap retrieval-salience surrogates.
\end{enumerate}
A policy satisfies the \emph{gate discipline} if fast admission, deployment rejection, debt thresholds, and governance-safety thresholds may depend only on Tier-1 and Tier-2 quantities together with the slack-robust substitution rule of Proposition~\ref{prop:proxy_gap}. Tier-3 quantities may influence routing, challenge sampling, reserve triage, audit prioritization, and alerting, but may not by themselves trigger semantic deployment or semantic rejection.
\end{definition}

\begin{corollary}[Conservative envelopes do not collapse into policy fiat under gate discipline]\label{cor:proxy_tiers}
Under Definition~\ref{def:proxy_tiers}, a published conservative envelope can affect fast deployment only if it is either directly metered or replayably derived from directly metered quantities with an explicit proof-carrying error budget. Consequently heuristic routing signals may steer attention but cannot by themselves manufacture admissibility.
\end{corollary}

\begin{proof}
This is immediate from the gate discipline: Tier-3 quantities are excluded from deployment and rejection thresholds by definition, while Tier-1 and Tier-2 quantities are constrained by direct logging or proof-carrying derivation together with Proposition~\ref{prop:proxy_gap}.
\end{proof}
\section{Core Definitions}
\begin{definition}[Claimed improvement]\label{def:claimed_improvement}
Let $\Theta_t^{\mathrm{fast}}$ denote the frozen issuance-time fast-evaluation interface, consisting of the evaluator state $e_t$ together with the parser choice, retrieval contract, and other fast-scoring manifest objects declared in $\Pi_t$ that policy treats as part of the immediate ruler. Write
\[
J_t^{\mathrm{fast}}(x,m)\Defeq J(x,m;\Theta_t^{\mathrm{fast}})
\]
for the resulting fast public evaluator score. The proposal issued at time $t$ has claimed improvement
\[
\Delta_t^{\mathrm{claim}}\Defeq J_t^{\mathrm{fast}}(x_{t+1},m_{t+1})-J_t^{\mathrm{fast}}(x_t,m_t).
\]
The interface $\Theta_t^{\mathrm{fast}}$ is held fixed in this comparison so that the claim isolates the local edit from subsequent evaluator drift. If the proposal explicitly challenges part of the fast interface itself, then claimed improvement is computed relative to the pre-edit interface and the interface edit must be certified separately through comparator compatibility. Claimed improvement is therefore a frozen-interface diagnostic, not a license to count immediate score changes caused only by rewriting the ruler.
\end{definition}

\begin{definition}[Yardstick drift]\label{def:yardstick_drift}
The one-step endogenous yardstick drift is
\[
D_t^{\mathrm{yard}}\Defeq \norm{e_{t+1}-e_t},
\]
and the cumulative drift over horizon $h\ge1$ is
\[
\mathcal D_t(h)\Defeq \sum_{s=t}^{t+h-1} D_s^{\mathrm{yard}}.
\]
\end{definition}

\begin{definition}[Comparator certificates for ruler edits]\label{def:comp_cert}
At decision epoch $t$, the replayable post-$t$ environment is represented by a typed exogenous-noise process
\[
W_s=(W_s^{\mathrm{pub}},W_s^{\mathrm{prot}},W_s^{\mathrm{exo}},W_s^{\mathrm{endo}}),\qquad s\ge t,
\]
where $W^{\mathrm{pub}}$ is replay-coupled public randomness, $W^{\mathrm{prot}}$ is protected audit randomness whose generator root and admissible law family are declared but whose realized bits may remain hidden from the proposal writer, $W^{\mathrm{exo}}$ is intervention-exogenous environment noise that is required to be coupled pathwise across the compared worlds, and $W^{\mathrm{endo}}$ is endogenous environment response generated only through the declared state-mediated transition kernels and therefore allowed to differ because the compared state paths differ. An \emph{interface-congruence certificate} publishes a common transition representation
\[
z_{s+1}=F_s(z_s,U_s,W_s)
\]
on a shared continuation skeleton together with an explicit declaration of which typed noise coordinates are coupled exactly, which are law-coupled, which admissible state-mediated response kernel governs $W^{\mathrm{endo}}$, and any replayable comparison-distortion budgets. A \emph{state-lift certificate} is admissible only if it provides a replayable lift map $L_t$ and projection $P_t$ satisfying all of the following: (i) $P_t\circ L_t$ preserves the pre-edit public state; (ii) all public observables and all public coordinates entering admissibility are preserved at issuance and thereafter tracked with explicit replayable distortion budgets; (iii) the lift-induced policy-manifold drift is uniformly upper-bounded in the published comparison coordinates; (iv) lift-only latent coordinates are not allowed to enter later admission certificates except through declared transported observables and public Lipschitz distortion budgets; and (v) no new probe family introduced by the lift may enter admission or challenge design without its own replayable transport certificate. Thus state augmentation cannot be used to trivialize comparator compatibility by adding private bookkeeping coordinates that later re-enter certification directly.
\end{definition}

\begin{definition}[Public continuation skeleton and comparator-compatible ruler edits]\label{def:continuation_skeleton}
Let
\[
\Pi_t^{\mathrm{cont}}
\Defeq
\bigl(\mathcal K_t^{\mathrm{adm}},\mathcal K_t^{\mathrm{route}},\mathcal K_t^{\mathrm{chal}},\mathcal K_t^{\mathrm{pf}},\mathcal K_t^{\mathrm{alloc}},\mathcal K_t^{\mathrm{dyn}},F_{t:},\mathfrak s_t\bigr)
\]
denote the \emph{public continuation skeleton}, where the components are respectively the replayable downstream admission, routing, challenge, proof-checking, allocator, and state-update kernels together with the typed transition family $F_{t:}$ and the public seed schedule $\mathfrak s_t$ used to couple replay-visible randomness across counterfactual replays. Equality of continuation skeletons means componentwise equality of these logged kernels, transition families, and the replay seed schedule. Hidden scrubber randomness is a different typed object and is not part of $\mathfrak s_t$.

A proposal is \emph{comparator-compatible} for stable-gain analysis if both $\ReplayCF(U_t=u_t)$ and $\ReplayCF(U_t=0)$ are feasible at the current public state and there exists a nonempty \emph{finite replayably enumerated} family
\[
\mathfrak C_t(u_t)=\{c_t^{(1)},\dots,c_t^{(N_t(u_t))}\},\qquad 1\le N_t(u_t)<\infty,
\]
of comparator witnesses. Each witness $c\in\mathfrak C_t(u_t)$ consists of coupled post-$t$ worlds generated on the same logged continuation skeleton, sharing the same $W^{\mathrm{pub}}$ path and the same $W^{\mathrm{exo}}$ path exactly, while $W^{\mathrm{prot}}$ is either exactly coupled or law-coupled according to the declared audit contract and $W^{\mathrm{endo}}$ is generated by a replayably published state-mediated response kernel. Each witness additionally carries a replayable comparison-distortion budget $\varepsilon_{t,s}^{\mathrm{cmp}}(c)$ on the public observables and admission coordinates. The comparison is \emph{exact} when those distortion budgets vanish identically; otherwise it is an \emph{approximate comparator} or replayable $\varepsilon$-bisimulation comparator. The \emph{comparator diameter}
\[
\operatorname{diam}^{\mathrm{cmp}}_{t,H}(u_t)
\Defeq
\max_{c,c'\in\mathfrak C_t(u_t)}\sum_{s=1}^{H}\beta^{s-1}\bigl|\varepsilon_{t,s}^{\mathrm{cmp}}(c)-\varepsilon_{t,s}^{\mathrm{cmp}}(c')\bigr|
\]
is itself a replayable governance-burden term and may be thresholded or priced by policy; when $H=\infty$, this definition is used only under the manifest clause that the weighted comparator-distortion series is absolutely summable. Ruler edits are comparator-compatible only when policy publishes an interface-congruence certificate or constrained state-lift certificate in the sense of Definition~\ref{def:comp_cert}, showing that the edited ruler is represented as a state update inside the shared skeleton rather than as a counterfactual-specific replacement of the comparison machinery.
\end{definition}

\begin{remark}[Scope of comparator compatibility]\label{rem:comp_scope}
Comparator compatibility is intentionally narrower than unrestricted self-rewrite. It covers edits whose compared worlds can be embedded in a shared replayable continuation skeleton, including many evaluator, parser, routing, memory, and policy edits once an interface-congruence or state-lift certificate is published. It does not by itself cover arbitrary rewrites of the ambient transition-law class when no such shared skeleton is available. Those cases require an additional meta-dynamics layer and lie outside the present stability claims.
\end{remark}

\begin{definition}[Public dominating envelope for infinite-horizon stable gain]\label{def:dominating_envelope}
Fix an issuance time $t$, a feasible intervention $u_t$, and a discount factor $\beta\in(0,1)$. For any admissible comparator witness $c\in\mathfrak C_t(u_t)$, let $\{(z_{t+s}^{u_t,c},z_{t+s}^{0,c})\}_{s\ge1}$ be the coupled post-intervention trajectories generated by the shared continuation skeleton of Definition~\ref{def:continuation_skeleton}. For the infinite-horizon case $H=\infty$, a \emph{public dominating envelope} is any replayable predictable sequence $M_{t,s}^{\mathrm{dom}}\ge 0$ such that
\[
\abs{V(z_{t+s}^{u_t,c})-V(z_{t+s}^{0,c})}\le M_{t,s}^{\mathrm{dom}}
\qquad\text{almost surely for all }s\ge 1\text{ and all }c\in\mathfrak C_t(u_t),
\]
and
\[
\E\!\left[\sum_{s=1}^{\infty}\beta^{s-1}M_{t,s}^{\mathrm{dom}}\middle|\mathcal F_t\right]<\infty.
\]
This object supplies the conditional absolute-integrability witness required to define infinite-horizon discounted stable gain without appealing to an unlogged analytic assumption.
\end{definition}

\begin{definition}[Discounted stable gain, stable improvement, governance safety, and full admissibility]\label{def:full_admissibility}
Fix horizon $H\in\N\cup\{\infty\}$ and discount factor $\beta\in(0,1]$, with $\beta<1$ whenever $H=\infty$. Let $V(x,m,h)$ be a slower long-run utility surrogate that does not depend directly on the fast evaluator state, and by abuse of notation write
\[
V(z)\Defeq V(x,m,h)\qquad\text{for }z=(x,e,m,q,r,h,\Pi).
\]
Write $\mathsf{Comp}_t(u_t)\in\{0,1\}$ for the public comparator-compatibility predicate induced by Definition~\ref{def:continuation_skeleton}. Stable-gain comparison uses only comparator-compatible proposals. For each such proposal, let $\mathfrak C_t(u_t)$ be the replayably published family of admissible comparator witnesses from Definition~\ref{def:continuation_skeleton}. For every witness $c\in\mathfrak C_t(u_t)$, let $z_{t+s}^{u_t,c}$ and $z_{t+s}^{0,c}$ denote the resulting coupled trajectories and let $\delta_{t,s}^{\mathrm{cmp}}(c)\ge 0$ be the declared comparison-distortion budget on the public observables and admissibility coordinates at lag $s$. Because $\mathfrak C_t(u_t)$ is finite and replayably enumerated, the lower and upper oracle endpoints below are literal minima and maxima over a finite family rather than essential envelopes.

The discounted stable-gain \emph{oracle interval} of the proposal issued at time $t$ is
\[
\mathcal G_{t,H}^{\beta,\mathrm{lo}}(u_t)
\Defeq
\min_{c\in\mathfrak C_t(u_t)}
\E\left[
\sum_{s=1}^{H}\beta^{s-1}\bigl(V(z_{t+s}^{u_t,c})-V(z_{t+s}^{0,c})\bigr)
\middle|\mathcal F_t
\right],
\]
\[
\mathcal G_{t,H}^{\beta,\mathrm{hi}}(u_t)
\Defeq
\max_{c\in\mathfrak C_t(u_t)}
\E\left[
\sum_{s=1}^{H}\beta^{s-1}\bigl(V(z_{t+s}^{u_t,c})-V(z_{t+s}^{0,c})\bigr)
\middle|\mathcal F_t
\right].
\]
When the comparator is exact, the interval collapses to a point. Throughout the remainder of the paper,
\[
\mathcal G_{t,H}^{\beta}(u_t)\Defeq \mathcal G_{t,H}^{\beta,\mathrm{lo}}(u_t)
\]
denotes the lower oracle endpoint, because admission is intentionally based on the worst replayably admissible comparator witness rather than on a hidden favorable coupling choice. When $H=\infty$, each witness is additionally assumed to satisfy conditional absolute integrability under a public dominating envelope in the sense of Definition~\ref{def:dominating_envelope}. Let
\[
\underline{\mathcal G}_{t,H}^{\beta}(u_t)
\]
denote any replayable lower certificate for $\mathcal G_{t,H}^{\beta}(u_t)$ published under the active admission policy. The intended reading is three-layered: the full oracle interval records comparator ambiguity, the lower oracle endpoint is the robust theoretical target, and $\underline{\mathcal G}_{t,H}^{\beta}$ is the operational object allowed to enter admission.

Let $\bar\delta_r(C_t)\ge 0$ be the policy debt threshold and define the proposal-local peak debt over horizon $H$ by
\[
\eta_{t,H}^{\max}(u_t)\Defeq \sup_{0\le s\le H}\eta_{t,t+s},
\]
with the natural infinite-horizon supremum $\sup_{s\ge 0}\eta_{t,t+s}$ when $H=\infty$. In particular, the issuance-time escrow liability $\eta_{t,t}$ is part of the stable-improvement gate rather than an auxiliary side remark.

Let $\mathsf{BaseAdm}_t(u_t)\in\{0,1\}$ denote the public base-admissibility predicate induced by the active canonical admission policy at time $t$, including replay success, declared cost budget, side-effect well-definedness, mandatory challenge routing for ruler edits, and summary typing or context-preservation clauses. A proposal at time $t$ is an \emph{$(H,\beta)$-oracle stable improvement} if
\[
\mathsf{Comp}_t(u_t)=1,
\qquad
\mathcal G_{t,H}^{\beta}(u_t)\ge 0,
\qquad\text{and}\qquad
\eta_{t,H}^{\max}(u_t)\le \bar\delta_r(C_t).
\]
It is an \emph{$(H,\beta)$-certified stable improvement} if
\[
\mathsf{Comp}_t(u_t)=1,
\qquad
\underline{\mathcal G}_{t,H}^{\beta}(u_t)\ge 0,
\qquad\text{and}\qquad
\eta_{t,H}^{\max}(u_t)\le \bar\delta_r(C_t).
\]
Let $\mathsf{GovSafe}_{t,H}(u_t)\in\{0,1\}$ denote the public governance-safety predicate asserting that the active non-gain governance clauses remain satisfied on the same horizon, including any required queue slack, replay floor, contradiction-reserve, heterodoxy-routing, proof-budget, and resource-escrow margins declared by policy. A proposal is an \emph{$(H,\beta)$-full admissible improvement} if
\[
\mathsf{BaseAdm}_t(u_t)=1,\qquad
u_t\text{ is an }(H,\beta)\text{-certified stable improvement},\qquad
\mathsf{GovSafe}_{t,H}(u_t)=1.
\]
This full-admissibility notion is an \emph{immediate-deployment} notion at epoch $t$: it concerns proposals whose current action is semantic deployment into the target lane under the active manifest and thresholds. Proposals that are admitted only to nondeployment actions such as challenge routing, richer-proof escalation, or sandboxing may be lane-admissible without yet being fully admissible in this immediate-deployment sense; they enter the productive frontier only after the delayed certificates and review actions required for deployment have been completed.

For brevity, the remainder of the paper uses ``stable improvement'' for the certified notion and qualifies the oracle notion explicitly when needed. It uses ``admissible improvement'' to mean the full immediate-deployment notion unless explicitly qualified otherwise.
\end{definition}

\begin{remark}[Why $V$ excludes fast governance variables]\label{rem:V_layering}
The surrogate $V$ is intentionally layered away from the fast evaluator state and from explicit governance variables such as queue backlog, replay fragility, debt, or policy-manifest drift. The reason is not that these quantities are normatively irrelevant, but that they are already enforced through $\mathsf{BaseAdm}$ and $\mathsf{GovSafe}$. Folding them directly into $V$ would blur the distinction between ``this edit has long-run task value'' and ``this edit preserves the governability of the loop,'' thereby double-counting governance failures. Applications that genuinely want queue, debt, or policy drift inside welfare may absorb them into $h_t$ or replace $V$ by a richer long-run functional, provided the same layer separation is made explicit.
\end{remark}

\begin{definition}[Structural blast radius, causal intervention capacity, and systemic risk bond]\label{def:blast_radius}
For a proposal $u_t$, let $\mathcal W_t(u_t)$ be its declared writable module set and let $\operatorname{Dep}_t(\mathcal W_t(u_t))$ be the replayable dependency closure of that set under the currently declared code, parser, evaluator, and memory manifests. Let $\chi_t^{\mathrm{priv}}(u_t)\ge 0$ be a public privilege multiplier that increases when the proposal can directly modify high-leverage substrates such as evaluator code, parser registries, refreezing policies, routing policies, memory-lane transition logic, or physical-resource allocators. A \emph{causal intervention capacity lower envelope} is any replayable public quantity
\[
\Gamma_t^{\mathrm{cic}}(u_t)\ge 0
\]
that lower-bounds the proposal's potential output-side or governance-side influence under admissible replays, for example through conservative intervention-response envelopes, minimum-description-length proxies, or algorithmic-causal surrogates computed across the active shadow/replay family. A \emph{systemic risk bond} is any replayable lower envelope
\[
\mathcal B_t(u_t)\ge 0
\]
that is monotone in writable scope, dependency closure, privilege level, and causal intervention capacity. One canonical robust instance is
\[
\mathcal B_t(u_t)
\Defeq
\max\!\left\{
\lambda_W |\mathcal W_t(u_t)|
+\lambda_D |\operatorname{Dep}_t(\mathcal W_t(u_t))|
+\lambda_P \chi_t^{\mathrm{priv}}(u_t),\,
\lambda_C \Gamma_t^{\mathrm{cic}}(u_t)
\right\},
\]
with public policy weights $\lambda_W,\lambda_D,\lambda_P,\lambda_C\ge 0$. The role of $\mathcal B_t(u_t)$ is not to predict exact harm, but to impose a proposal-local lower bound on downside exposure that cannot be driven to zero merely by under-reporting claimed upside or by topologically compressing a globally influential rewrite into a single apparent module.
\end{definition}

\begin{remark}[Operational meaning of the intervention-capacity envelope]\label{rem:cic_proxy}
$\Gamma_t^{\mathrm{cic}}(u_t)$ is allowed to be conservative. In practice it may be instantiated by minimum guaranteed effect on declared governance outputs under admissible perturbation probes, by a lower confidence bound on intervention-response magnitude across the active shadow and replay families, or by other monotone proxies that cannot be made small merely by renaming modules. The theorem layer needs only a replayable lower envelope; it does not require exact causal identification.
\end{remark}

\begin{definition}[Cohort-escrow error debt]\label{def:cohort_escrow}
For each proposal cohort $\tau$, let $g_{\tau,t}^{\mathrm{chal}}\ge 0$ denote the \emph{cumulative gross matched retirement} credited by review time $t\ge \tau$, with $t\mapsto g_{\tau,t}^{\mathrm{chal}}$ nondecreasing and $g_{\tau,\tau}^{\mathrm{chal}}=0$. Retirement credit is allowed only for replayably matched repair certificates, verified rollback-cost reductions, or persistent hazard-reduction certificates; a superficial score recovery that leaves the repair or hazard reserve unresolved does not count as retirement. Let $a_{\tau,t}^{\mathrm{sec}}\ge 0$ denote \emph{secondary liability} conservatively charged back to the parent cohort between times $t-1$ and $t$; this term captures repair-induced regressions or parser/environment migrations that were triggered by attempts to neutralize the parent cohort. Let $b_{\tau,t}\ge0$ be a nonnegative backlog-attribution kernel satisfying
\[
\sum_{\tau\le t} b_{\tau,t}\le q_t,
\]
and let $\rho_{\tau,t}^{\mathrm{ret}}\in[0,1]$ be a predictable retirement fraction capturing only the verified reserve-reduction actions described above. Let $U_{\tau}^{\mathrm{rep}}\ge 0$ be a replayable upper envelope on repair or rollback cost required to neutralize the cohort under the active deployment scope, and let $U_{\tau}^{\mathrm{haz}}\ge 0$ be a replayable upper envelope on persistent unresolved hazard cost under the same scope. The proposal-local liability stock is initialized by the \emph{unresolved downside reserve}
\[
\eta_{\tau,\tau}=\mathcal B_{\tau}(u_{\tau})+U_{\tau}^{\mathrm{rep}}+U_{\tau}^{\mathrm{haz}},
\]
where $\mathcal B_{\tau}(u_{\tau})$ is the systemic risk bond from Definition~\ref{def:blast_radius}, and it evolves as
\[
\eta_{\tau,t+1}
=
\Bigl[
(1-\rho_{\tau,t+1}^{\mathrm{ret}})\alpha_r \eta_{\tau,t}
-
\bigl(g_{\tau,t+1}^{\mathrm{chal}}-g_{\tau,t}^{\mathrm{chal}}\bigr)
+
a_{\tau,t+1}^{\mathrm{sec}}
+
\lambda_q b_{\tau,t}
\Bigr]_+,
\qquad t\ge \tau,
\]
where $\alpha_r\in[0,1]$ and $\lambda_q\ge0$. Aggregate error debt is then
\[
r_t\Defeq \sum_{\tau<t}\eta_{\tau,t}.
\]
By convention, $r_t$ is the \emph{pre-admission actionable stock} visible when deciding whether to admit the time-$t$ proposal. Newly issued cohort-$t$ liability therefore enters the stock beginning at $r_{t+1}$, so the indexing is intentional rather than an off-by-one omission.
This accounting has five important properties. First, matched challenge retirement acts only inside the cohort that generated the claim, so cross-cohort masking is excluded by construction. Second, explicit retirement shrinks already accumulated stock rather than merely blocking future accrual. Third, repair-induced regressions are not hidden by making the matched-retirement process nonmonotone; they enter transparently through the secondary-liability term $a_{\tau,t}^{\mathrm{sec}}$ or, equivalently, via newborn derived cohorts. Fourth, $\alpha_r<1$ is an \emph{operational relevance discount}, not an epistemic erasure rule: evidence, contradiction tags, and replay artifacts can still be preserved indefinitely, while the liability stock tracks only the burden that remains actionable for the current loop. Fifth, the initial reserve is coherent downside accounting: it lower-bounds systemic reach, repair cost, and persistent hazard cost instead of depending on optimistic positive claims. The strict-conservation case $\alpha_r=1$ is treated separately below.
\end{definition}

\begin{assumption}[Finite active cohort support]\label{ass:finite_active_cohorts}
For every review time $t$, only finitely many proposal cohorts carry nonzero actionable liability, equivalently
\[
\#\{\tau<t:\eta_{\tau,t}>0\}<\infty,
\]
or, more generally, the policy exposes a replayable truncation rule ensuring
\[
r_t=\sum_{\tau<t}\eta_{\tau,t}<\infty
\qquad\text{almost surely for every }t.
\]
This assumption prevents aggregate debt from becoming a formal divergent series rather than a state variable.
\end{assumption}

\begin{remark}[Why monotone gross retirement is not a no-regression assumption]
The monotonicity of $g_{\tau,t}^{\mathrm{chal}}$ means only that once a matched retirement credit has been publicly granted to the parent cohort, that historical credit is not silently revoked. A bad repair can still make the world worse; it appears either as a new cohort or as positive secondary liability $a_{\tau,t}^{\mathrm{sec}}$. Thus monotone gross retirement is compatible with regression bugs and avoids overloading one variable with two opposite semantics.
\end{remark}

\begin{definition}[Retained margin and improvement half-life]\label{def:retained_margin}
Let
\[
\mathcal M_t(h)
\Defeq
\E\!\left[
\widehat V_{t+h}^{\Shadow}(u_t)-\widehat V_{t+h}^{\Shadow}(0)
\mid \mathcal F_t
\right]
\]
be the shadow-reevaluated retained margin of the proposal issued at time $t$ after horizon $h\ge0$, where $\widehat V^{\Shadow}$ is computed using the issuance-time shadow family and issuance-time aggregator transported replayably to horizon $t+h$ in the sense of Definition~\ref{def:audit_family}. The value $\mathcal M_t(0)$ is the \emph{frozen shadow baseline}: it is evaluated at issuance time using the pre-edit shadow configuration before post-adoption yardstick drift and backlog accumulation are allowed to act on the cohort. For proposals with $\mathcal M_t(0)>0$, the \emph{improvement half-life} is
\[
\tau_{1/2}(t)\Defeq \inf\left\{h\ge1:\mathcal M_t(h)\le \tfrac12 \mathcal M_t(0)\right\}.
\]
\end{definition}

\begin{definition}[Operational retained-margin lower certificate]\label{def:retained_margin_cert}
For a proposal cohort issued at time $t$, let $\mathcal M_t^{\mathrm{op}}(h)$ denote the \emph{operational class-conditional retained-margin target} defined later in Definition~\ref{def:transport_calibration}. A family $\{\underline{\mathcal M}_t^{\mathrm{op}}(h)\}_{h\ge 1}$ is a \emph{replayable retained-margin lower certificate} if each $\underline{\mathcal M}_t^{\mathrm{op}}(h)$ is a replayable function of the matured challenge evidence, manifest-bound logs, class-assignment metadata, comparator-distortion budgets, confidence-budget manifests, and any published transport-calibration metadata available by horizon $h$, and if
\[
\underline{\mathcal M}_t^{\mathrm{op}}(h)\le \mathcal M_t^{\mathrm{op}}(h)
\qquad\text{for every }h\ge 1
\]
on every run consistent with the published challenge protocol. This certificate is \emph{admission-facing}: it lower-bounds the replayably published class-conditional target used by Interface A. The cohort-specific oracle margin $\mathcal M_t(h)$ from Definition~\ref{def:retained_margin} remains a distinct latent object and is related to the operational target only through additional homogeneity or transport assumptions stated explicitly below.
\end{definition}

\begin{definition}[Blocked randomized challenge design with logged propensities]\label{def:challenge_design}
Fix a horizon $h\ge 1$ and a challenge-eligible cohort class. A replayable challenge design publishes, for each issued cohort $i$, issuance covariates $X_i$, proposal signature $U_i$, class label $\mathcal C_i$ in a replayably declared finite challenge-equivalence partition, challenge indicator $A_i\in\{0,1\}$, logged propensity $p_i\in[\underline p,1]$ with $\underline p>0$, horizon-maturity indicator $O_{i,h}(t)\in\{0,1\}$ at certification epoch $t$, delay label $D_{i,h}$, challenge-block identifier $B_i$, replayably declared block-compatibility family label $\mathfrak B_i$, and any replayable transport features needed to compare the issuance-time shadow family to the matured challenged family. For every issued cohort whose horizon-$h$ window has matured by epoch $t$, let $Y_{i,h}^{\mathrm{ret}}$ denote the latent matured delayed challenge outcome under the issuance-time challenge metric transported replayably to maturity, and define the actually observed challenge score by
\[
Y_{i,h}^{\mathrm{obs}}(t)\Defeq A_i\,O_{i,h}(t)\,Y_{i,h}^{\mathrm{ret}}.
\]
In addition, the design partitions time into replayably declared \emph{challenge blocks} on each of which the evaluator, parser family, routing policy, class-assignment rule, and target metric used to define $Y_{i,h}^{\mathrm{ret}}$ are frozen. Each block belongs to a replayably declared finite \emph{block-compatibility family} $\mathfrak B$ such that cohorts with the same $(\mathcal C,\mathfrak B)$ pair share the same admissible challenge metric, parser/evaluator family, and declared transport interface into the current certification regime. The class-assignment map $(X_i,U_i)\mapsto \mathcal C_i$ and the block-family assignment map $(X_i,U_i,B_i)\mapsto \mathfrak B_i$ must themselves be replayably published, block-frozen, and immune to proposer-controlled post-issuance metadata: they may depend only on predeclared issuance-time features and may not be changed after observing challenge outcomes or current challenge-lane congestion. A working regression $\widehat m_h(X,U,\mathcal C,\mathfrak B)$ may be published for efficiency, but admissibility is always based on replayable logged propensities, outcomes, block identifiers, class/block-family metadata, and transport metadata rather than on an unlogged prediction model.

For a class $\mathcal C$, block-compatibility family $\mathfrak B$, horizon $h$, and certification epoch $t$, define the \emph{operational class/block-conditional retained challenge margin}
\[
\mathcal M_t^{\mathrm{chal}}(h;\mathcal C,\mathfrak B)
\Defeq
\E\bigl[Y_{i,h}^{\mathrm{ret}}\mid \mathcal C_i=\mathcal C,\ \mathfrak B_i=\mathfrak B,\ O_{i,h}(t)=1\bigr].
\]
The current proposal issued at time $t$ inherits its operational target from the pair $(\mathcal C_t,\mathfrak B_t)$ assigned by the active block-frozen policy to $(X_t,u_t,B_t)$. If the current pair has no mature support, or fewer than the manifest-declared minimum support count required for certification, the design is \emph{fail-closed} for fast deployment: no class/block-conditional Interface-A certificate exists until the proposal is routed to challenge, proof escalation, sandbox, or a richer evidentiary class.
\end{definition}

\begin{definition}[Challenge-design defect envelopes and fail-closed bias accounting]\label{def:chal_defects}
For a current class/block pair $(\mathcal C_t,\mathfrak B_t)$ and horizon $h$, let
\[
\delta_h^{\mathrm{design}}(\mathcal C_t,\mathfrak B_t)
\Defeq
\delta_h^{\mathrm{int}}(\mathcal C_t,\mathfrak B_t)
+\delta_h^{\mathrm{pos}}(\mathcal C_t,\mathfrak B_t)
+\delta_h^{\mathrm{ign}}(\mathcal C_t,\mathfrak B_t)
+\delta_h^{\mathrm{supp}}(\mathcal C_t,\mathfrak B_t)
+\delta_h^{\mathrm{clip}}(\mathcal C_t,\mathfrak B_t)
\]
be a replayably published conservative design-defect envelope, where:
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item $\delta_h^{\mathrm{int}}$ bounds residual cross-cohort interference after blocking;
  \item $\delta_h^{\mathrm{pos}}$ bounds the worst-case bias induced by propensity-floor violations, trimming, truncation, or effective support loss on the matured issued support;
  \item $\delta_h^{\mathrm{ign}}$ bounds the worst-case deviation from sequential ignorability allowed by the logged sensitivity analysis on the current pair;
  \item $\delta_h^{\mathrm{supp}}$ bounds mismatch between the currently certified class/block target and the deployment-relevant support for the current proposal, including class-assignment drift or support collapse not already excluded by the block-frozen manifest;
  \item $\delta_h^{\mathrm{clip}}$ bounds any bias introduced by replayable clipping, truncation, or stabilized reweighting used to obtain a bounded-score certificate.
\end{enumerate}
Any unavailable component is treated as $+\infty$ for fast deployment, which forces the interface to fail closed rather than to extrapolate from unsupported support. Zero design defect is the special case in which the blocked design, support, and bounded-score instrumentation are exact on the certified class/block target.
\end{definition}

\begin{proposition}[Admissibility-scale composition of challenge-design defects]\label{prop:chal_defect_comp}
Fix a horizon $h$ and current class/block pair $(\mathcal C_t,\mathfrak B_t)$. Suppose each component in Definition~\ref{def:chal_defects} is a replayably published nonnegative bound on a downward perturbation of the same operational target $\mathcal M_t^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)$ after all propensity, clipping, blocking, and transport normalizations already used in the displayed estimator and confidence radius. Then the total downward perturbation is conservatively bounded by
\[
\delta_h^{\mathrm{design}}(\mathcal C_t,\mathfrak B_t)
=
\delta_h^{\mathrm{int}}(\mathcal C_t,\mathfrak B_t)
+\delta_h^{\mathrm{pos}}(\mathcal C_t,\mathfrak B_t)
+\delta_h^{\mathrm{ign}}(\mathcal C_t,\mathfrak B_t)
+\delta_h^{\mathrm{supp}}(\mathcal C_t,\mathfrak B_t)
+\delta_h^{\mathrm{clip}}(\mathcal C_t,\mathfrak B_t).
\]
In particular, the fast gate may safely subtract the full published sum whenever the individual defect certificates are expressed on this common admissibility scale.
\end{proposition}

\begin{proof}
Let $\mathcal M^{(0)}$ denote the exact class/block target and let $\mathcal M^{(j)}$ be the intermediate target obtained after applying the first $j$ declared defect relaxations in the order interference, positivity/support, ignorability, support transfer, and clipping. By hypothesis,
\[
\mathcal M^{(j-1)}-\mathcal M^{(j)}\le \delta_j
\qquad\text{for each declared component }\delta_j\ge 0,
\]
on the same retained-margin scale. Summing these inequalities and using the triangle inequality gives
\[
\mathcal M^{(0)}-\mathcal M^{(5)}\le \sum_{j=1}^5 \delta_j = \delta_h^{\mathrm{design}}(\mathcal C_t,\mathfrak B_t).
\]
Thus subtracting the full published sum is conservative even if some components are zero, merged, or vacuous.
\end{proof}

\begin{theorem}[Finite-sample replayable lower certificate from blocked randomized delayed challenge]\label{thm:dr_shadow_cert}
Fix a horizon $h\ge1$ and assume Definitions~\ref{def:challenge_design} and \ref{def:chal_defects}. Consider the current proposal pair $(\mathcal C_t,\mathfrak B_t)$. Suppose: (i) challenge routing is sequentially ignorable given the logged filtration generated by $(X_i,U_i,\mathcal C_i,\mathfrak B_i)$ and the block identifier, up to the published sensitivity envelope $\delta_h^{\mathrm{ign}}(\mathcal C_t,\mathfrak B_t)$; (ii) logged propensities are published together with any positivity-floor defect envelope $\delta_h^{\mathrm{pos}}(\mathcal C_t,\mathfrak B_t)$ induced by trimming, truncation, or support loss on the matured issued support; (iii) challenge delay is nonanticipating relative to the same filtration; (iv) within each block, evaluator, parser, routing policy, class-assignment rule, block-family rule, and challenge metric are frozen; (v) blocks pooled under the current family label $\mathfrak B_t$ are replayably certified transport-compatible for the challenge target, up to the published support-mismatch envelope $\delta_h^{\mathrm{supp}}(\mathcal C_t,\mathfrak B_t)$; (vi) either the logged propensities are correct or the published regression $\widehat m_h(X,U,\mathcal C,\mathfrak B)$ is correctly specified on the current pair; and (vii) the defect components are expressed on the common admissibility scale required by Proposition~\ref{prop:chal_defect_comp}. Assume further that latent matured challenge outcomes are uniformly bounded by a replayable envelope $|Y_{i,h}^{\mathrm{ret}}|\le B_h$.

Let $\mathbf J_{t,n}(\mathcal C_t,\mathfrak B_t)=(i_1,\dots,i_n)$ denote the first $n$ issued cohorts with $(\mathcal C_{i_k},\mathfrak B_{i_k})=(\mathcal C_t,\mathfrak B_t)$ whose horizon-$h$ certification window has matured by epoch $t$, i.e. $O_{i_k,h}(t)=1$ for each $k$, and let $n_{\min}(h,\mathcal C_t,\mathfrak B_t)$ be the manifest-declared minimum matured support count from Definition~\ref{def:confidence_manifest}. Define the class/block-conditional doubly robust estimator
\[
\widehat{\mathcal M}_{t,n}^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)
\Defeq
\frac1n\sum_{k=1}^{n}\left[\widehat m_h(X_{i_k},U_{i_k},\mathcal C_t,\mathfrak B_t)+\frac{Y_{i_k,h}^{\mathrm{obs}}(t)-A_{i_k}O_{i_k,h}(t)\widehat m_h(X_{i_k},U_{i_k},\mathcal C_t,\mathfrak B_t)}{p_{i_k}}\right].
\]
If $n<n_{\min}(h,\mathcal C_t,\mathfrak B_t)$, no class/block-conditional certificate exists and Interface A is operationally inconclusive for fast deployment on $(\mathcal C_t,\mathfrak B_t)$. Otherwise there exists a replayable anytime radius $\operatorname{rad}_{t,n}^{\mathrm{chal}}(h,\alpha;\mathcal C_t,\mathfrak B_t)$, computable from the logged propensities, bounded outcomes, block structure, and standard confidence-sequence or e-process ingredients \cite{HowardRamdasMcAuliffeSekhon2021,RamdasGrunwaldVovkShafer2023,ZhanHadadHirshbergAthey2021}, such that with probability at least $1-\alpha$ simultaneously for all $n\ge n_{\min}(h,\mathcal C_t,\mathfrak B_t)$,
\[
\mathcal M_t^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)
\ge
\widehat{\mathcal M}_{t,n}^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)
-\operatorname{rad}_{t,n}^{\mathrm{chal}}(h,\alpha;\mathcal C_t,\mathfrak B_t)
-\delta_h^{\mathrm{design}}(\mathcal C_t,\mathfrak B_t).
\]
Consequently
\[
\underline{\mathcal M}_{t,n}^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)
\Defeq
\widehat{\mathcal M}_{t,n}^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)
-\operatorname{rad}_{t,n}^{\mathrm{chal}}(h,\alpha;\mathcal C_t,\mathfrak B_t)
-\delta_h^{\mathrm{design}}(\mathcal C_t,\mathfrak B_t)
\]
is a replayable finite-sample lower certificate for the operational class/block-conditional retained challenge margin. The theorem certifies the current proposal's replayably assigned class/block target at certification epoch $t$; it is not a direct certificate for the current proposal's cohort-specific oracle margin.
\end{theorem}

\begin{proof}
On the matured issued support $O_{i,h}(t)=1$, the observed challenge score satisfies $Y_{i,h}^{\mathrm{obs}}(t)=A_iY_{i,h}^{\mathrm{ret}}$. Therefore the displayed doubly robust score can be written as
\[
\widehat m_h(X_i,U_i,\mathcal C_t,\mathfrak B_t)+\frac{A_iO_{i,h}(t)}{p_i}\bigl(Y_{i,h}^{\mathrm{ret}}-\widehat m_h(X_i,U_i,\mathcal C_t,\mathfrak B_t)\bigr),
\]
which is the standard missing-outcome doubly robust score on the matured issued support. Under the block-freezing and nonanticipating-delay clauses, its conditional expectation equals the operational class/block retained challenge margin whenever either the propensity model or the regression model is correct on $(\mathcal C_t,\mathfrak B_t)$ and the remaining departures from exact identification are absorbed by the published design-defect envelope. Bounded outcomes control inverse-propensity variance, while support loss, trimming, bounded-score clipping bias, residual cross-cohort interference, and logged ignorability sensitivity are subtracted conservatively through $\delta_h^{\mathrm{design}}$. Standard confidence-sequence or e-process constructions therefore produce a replayable radius $\operatorname{rad}_{t,n}^{\mathrm{chal}}$ that lower-bounds the class/block target uniformly over time on the matured issued support after debiting the declared defect budget. The fail-closed convention for empty, support-deficient, or diagnostically unsupported classes is immediate.
\end{proof}

\begin{proposition}[Executable empirical-Bernstein instantiation of the delayed-challenge radius]\label{prop:chal_radius}
Under the hypotheses of Theorem~\ref{thm:dr_shadow_cert}, let $\mathbf J_{t,n}(\mathcal C_t,\mathfrak B_t)=(i_1,\dots,i_n)$ be the ordered matured-issued class/block sample, and define the replayable doubly robust score for the $k$-th element of that sample by
\[
\xi_{k,h}(\mathcal C_t,\mathfrak B_t)
\Defeq
\widehat m_h(X_{i_k},U_{i_k},\mathcal C_t,\mathfrak B_t)+\frac{Y_{i_k,h}^{\mathrm{obs}}(t)-A_{i_k}O_{i_k,h}(t)\widehat m_h(X_{i_k},U_{i_k},\mathcal C_t,\mathfrak B_t)}{p_{i_k}}.
\]
Assume the manifest publishes either a replayably clipped regression or an equivalent bounded-score certificate guaranteeing $|\xi_{k,h}(\mathcal C_t,\mathfrak B_t)|\le B_{h,\mathcal C_t,\mathfrak B_t}^{\mathrm{DR}}$, that any bias induced by clipping, truncation, or stabilized weighting is already absorbed into $\delta_h^{\mathrm{clip}}(\mathcal C_t,\mathfrak B_t)$ from Definition~\ref{def:chal_defects}, and that the regression used to define $\xi_{k,h}$ is fixed before the $k$-th matured outcome is revealed (for example by using a disjoint replayable training sample or a prequential update rule). Define the sample-index filtration
\[
\mathcal G_{t,k}^{\mathrm{chal}}\Defeq \sigma\bigl(\mathbf J_{t,k}(\mathcal C_t,\mathfrak B_t), \{X_{i_j},U_{i_j},\mathcal C_{i_j},\mathfrak B_{i_j},p_{i_j},O_{i_j,h}(t),Y_{i_j,h}^{\mathrm{obs}}(t)\}_{j\le k}\bigr),
\]
with $\mathcal G_{t,0}^{\mathrm{chal}}$ containing the published regression and clipping certificate. Define the predictable empirical-variance process
\[
V_{t,n}^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)\Defeq \sum_{k=1}^{n}\bigl(\xi_{k,h}(\mathcal C_t,\mathfrak B_t)-\widehat{\mathcal M}_{t,k-1}^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)\bigr)^2,
\]
with the convention $\widehat{\mathcal M}_{t,0}^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)=0$, and let
\[
\nu(v,\alpha;B)\Defeq \sqrt{2(v\vee 1)\log(2/\alpha)}+3B\log(2/\alpha).
\]
Then one valid replayable choice is
\[
\operatorname{rad}_{t,n}^{\mathrm{chal}}(h,\alpha;\mathcal C_t,\mathfrak B_t)
\Defeq
\frac{\nu\!\left(V_{t,n}^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t),\alpha;B_{h,\mathcal C_t,\mathfrak B_t}^{\mathrm{DR}}\right)}{n\vee 1},
\]
which yields an anytime empirical-Bernstein lower confidence sequence for the operational class/block-conditional target under the same blocked design, predictable-score, and bounded-score assumptions.
\end{proposition}

\begin{proof}
For bounded predictable scores adapted to the sample-index filtration $\{\mathcal G_{t,k}^{\mathrm{chal}}\}_{k\ge0}$, standard stitched or mixture empirical-Bernstein confidence sequences control the centered partial sums uniformly over time in terms of the predictable quadratic variation and a bounded-increment envelope \cite{HowardRamdasMcAuliffeSekhon2021,RamdasGrunwaldVovkShafer2023}. Applying that construction to the replayably logged doubly robust scores on the ordered matured class/block sample gives the displayed radius. The theorem then follows by dividing by $n\vee1$ and subtracting the full declared design-defect envelope $\delta_h^{\mathrm{design}}$, whose additivity on the admission-facing retained-margin scale is justified by Proposition~\ref{prop:chal_defect_comp}; residual interference is only one component, and comparator distortion is handled later, at the bridge from class/block slow-utility targets to witnesswise oracle stable gain.
\end{proof}

\begin{definition}[Confidence-budget manifest for multi-horizon and multi-class certificates]\label{def:confidence_manifest}
A \emph{confidence-budget manifest} is a replayably published allocation
\[
\{\alpha_{h,\mathcal C,\mathfrak B}^{\mathrm{chal}}\}_{h\in\mathcal H,\ \mathcal C\in\mathfrak C^{\mathrm{chal}},\ \mathfrak B\in\mathfrak B^{\mathrm{chal}}},
\qquad
\{\alpha_{h,\mathcal C,\mathfrak B}^{\mathrm{tr}}\}_{h\in\mathcal H,\ \mathcal C\in\mathfrak C^{\mathrm{chal}},\ \mathfrak B\in\mathfrak B^{\mathrm{chal}}},
\]
with
\[
\sum_{h,\mathcal C,\mathfrak B}\alpha_{h,\mathcal C,\mathfrak B}^{\mathrm{chal}}+
\sum_{h,\mathcal C,\mathfrak B}\alpha_{h,\mathcal C,\mathfrak B}^{\mathrm{tr}}
\le \alpha_{\mathrm{tot}},
\]
and a replayably declared minimum matured support count map
\[
n_{\min}:\mathcal H\times\mathfrak C^{\mathrm{chal}}\times\mathfrak B^{\mathrm{chal}}\to\N.
\]
Unless a vector-valued anytime certificate is published, simultaneous validity across horizons, classes, block-compatibility families, and transport-calibration residual certificates is interpreted through this explicit budget split and a union bound under the active manifest.
\end{definition}

\begin{definition}[Transport calibration from challenge margin to slow utility]\label{def:transport_calibration}
For each replayably declared challenge-equivalence class $\mathcal C$, block-compatibility family $\mathfrak B$, and horizon $h$, a \emph{transport-calibration family} is a collection of replayable lower transport coefficients $a_h(\mathcal C,\mathfrak B)\ge 0$, offsets $b_h(\mathcal C,\mathfrak B)\ge 0$, and residual envelopes $\delta_h^{\mathrm{tr}}(\mathcal C,\mathfrak B)\ge 0$ such that for every comparator-compatible cohort in the pair $(\mathcal C,\mathfrak B)$ and every admissible comparator witness $c$,
\[
\E\bigl[V(z_{t+h}^{u_t,c})-V(z_{t+h}^{0,c})\mid \mathcal F_t\bigr]
\ge
 a_h(\mathcal C,\mathfrak B)\,\mathcal M_t^{\mathrm{chal}}(h;\mathcal C,\mathfrak B)-b_h(\mathcal C,\mathfrak B)-\delta_h^{\mathrm{tr}}(\mathcal C,\mathfrak B)-\delta_{t,h}^{\mathrm{cmp}}(c).
\]
Define the corresponding \emph{operational class/block-conditional retained-margin target} for the current proposal's pair $(\mathcal C_t,\mathfrak B_t)$ by
\[
\mathcal M_t^{\mathrm{op}}(h)
\Defeq
a_h(\mathcal C_t,\mathfrak B_t)\,\mathcal M_t^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)-b_h(\mathcal C_t,\mathfrak B_t)-\delta_h^{\mathrm{tr}}(\mathcal C_t,\mathfrak B_t).
\]
The coefficients may be learned or stress-tested from matched long-horizon replays or previously audited cohorts, but once published they are treated as part of the active replayable admission contract rather than as hidden analyst choices. Comparator distortion is handled downstream at the witnesswise stable-gain bridge rather than inside the class/block challenge certificate itself. A transport-calibration family is Tier-2 admissible only if the manifest also publishes a replayable matched-replay or audited transport certificate witnessing how the displayed lower envelope was derived on the declared class/block pair.
\end{definition}

\begin{proposition}[Matched-replay transport certificate implies Tier-2 transport calibration]\label{prop:transport_tier2}
Fix class $\mathcal C$, block-compatibility family $\mathfrak B$, and horizon $h$. Suppose the manifest publishes a replayable matched-replay sample of challenged cohorts in the pair $(\mathcal C,\mathfrak B)$ together with matched long-horizon slow-utility differences $\Delta V_{j,h}$ and retained challenge outcomes $Y_{j,h}^{\mathrm{ret}}$, all measured on the same matured support event $O_{j,h}(t)=1$, and suppose a proof-carrying certificate verifies that
\[
\begin{aligned}
\E[\Delta V_{j,h}\mid \mathcal C_j=\mathcal C,\ \mathfrak B_j=\mathfrak B,\ O_{j,h}(t)=1]
&\ge a_h(\mathcal C,\mathfrak B)\,\E[Y_{j,h}^{\mathrm{ret}}\mid \mathcal C_j=\mathcal C,\ \mathfrak B_j=\mathfrak B,\ O_{j,h}(t)=1]\\
&\quad -b_h(\mathcal C,\mathfrak B)-\delta_h^{\mathrm{tr}}(\mathcal C,\mathfrak B).
\end{aligned}
\]
for the published coefficients and residual envelope. Then the resulting transport-calibration family is Tier-2 admissible under Definition~\ref{def:proxy_tiers}: it may enter Interface A because its coefficients are no longer heuristic routing signals but proof-carrying replay-derived lower-envelope quantities.
\end{proposition}

\begin{proof}
The proof-carrying matched-replay certificate turns the transport coefficients into derived quantities whose validity reduces to replayable matched samples and the checker's proof obligations. By the gate discipline, such proof-carrying derived quantities are Tier-2 rather than Tier-3. Substituting them into Definition~\ref{def:transport_calibration} yields a replayable lower envelope on slow-utility change from the class/block-conditional challenge target.
\end{proof}

\begin{proposition}[Executable matched-replay transport residual certificate]\label{prop:transport_radius}
Fix class $\mathcal C$, block-compatibility family $\mathfrak B$, and horizon $h$. Suppose the manifest publishes a replayable matched-replay sample $j=1,\dots,m$ together with matched slow-utility differences $\Delta V_{j,h}$, retained challenge outcomes $Y_{j,h}^{\mathrm{ret}}$, and transport coefficients $a_h(\mathcal C,\mathfrak B),b_h(\mathcal C,\mathfrak B)$, all defined on the same matured support event $O_{j,h}(t)=1$. Define the residuals
\[
R_{j,h}^{\mathrm{tr}}(\mathcal C,\mathfrak B)\Defeq a_h(\mathcal C,\mathfrak B)Y_{j,h}^{\mathrm{ret}}-b_h(\mathcal C,\mathfrak B)-\Delta V_{j,h}.
\]
Assume a replayable bounded-residual certificate $|R_{j,h}^{\mathrm{tr}}(\mathcal C,\mathfrak B)|\le B_h^{\mathrm{tr}}(\mathcal C,\mathfrak B)$ and let $\overline R_{m,h}^{\mathrm{tr}}(\mathcal C,\mathfrak B;\alpha)$ be any replayable one-sided anytime upper confidence sequence for
\[
\E[R_{j,h}^{\mathrm{tr}}(\mathcal C,\mathfrak B)\mid \mathcal C_j=\mathcal C,\ \mathfrak B_j=\mathfrak B,\ O_{j,h}(t)=1]
\]
built from the matched sample. Then any published residual envelope satisfying
\[
\delta_h^{\mathrm{tr}}(\mathcal C,\mathfrak B)\ge \bigl(\overline R_{m,h}^{\mathrm{tr}}(\mathcal C,\mathfrak B;\alpha)\bigr)_+
\]
is a replayable Tier-2 transport residual certificate for Definition~\ref{def:transport_calibration}. In particular, matched-replay transport calibration can be made executable rather than purely contractual once the residual certificate and its confidence budget are published.
\end{proposition}

\begin{proof}
The transport inequality in Definition~\ref{def:transport_calibration} is equivalent to requiring that the conditional mean residual on the matched matured support be no larger than $\delta_h^{\mathrm{tr}}(\mathcal C,\mathfrak B)$. A replayable one-sided upper confidence sequence for that residual mean therefore yields a replayable lower-envelope certificate once its positive part is published as the transport residual budget. Because the matched-replay sample, bounded-residual certificate, and confidence budget are replayably logged, the resulting envelope is Tier-2 admissible.
\end{proof}

\begin{definition}[Class-homogeneity envelope]\label{def:class_homogeneity}
A replayable \emph{class-homogeneity envelope} for class $\mathcal C$, block-compatibility family $\mathfrak B$, and horizon $h$ is a nonnegative quantity $\delta_h^{\mathrm{class}}(\mathcal C,\mathfrak B)$ such that every cohort in the pair $(\mathcal C,\mathfrak B)$ satisfies
\[
\mathcal M_t(h)\ge \mathcal M_t^{\mathrm{op}}(h)-\delta_h^{\mathrm{class}}(\mathcal C,\mathfrak B).
\]
This envelope quantifies the residual gap between the cohort-specific oracle retained margin and the class/block operational target used for admission.
\end{definition}

\begin{proposition}[Transport-calibrated operational retained-margin certificate]\label{prop:challenge_to_shadow}
Fix the current proposal pair $(\mathcal C_t,\mathfrak B_t)$ and horizon $h\ge1$. Suppose Definition~\ref{def:transport_calibration} holds on $(\mathcal C_t,\mathfrak B_t)$ and let $\underline{\mathcal M}_{t,n}^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)$ be any replayable lower certificate from Theorem~\ref{thm:dr_shadow_cert}. Then
\[
\underline{\mathcal M}_t^{\mathrm{op}}(h)
\Defeq
 a_h(\mathcal C_t,\mathfrak B_t)\,\underline{\mathcal M}_{t,n}^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)-b_h(\mathcal C_t,\mathfrak B_t)-\delta_h^{\mathrm{tr}}(\mathcal C_t,\mathfrak B_t)
\]
is a replayable retained-margin lower certificate in the sense of Definition~\ref{def:retained_margin_cert} for the operational class/block target entering Proposition~\ref{prop:shadow_bridge}. If, in addition, Definition~\ref{def:class_homogeneity} holds on $(\mathcal C_t,\mathfrak B_t)$, then
\[
\underline{\mathcal M}_t^{\mathrm{oracle}}(h)\Defeq \underline{\mathcal M}_t^{\mathrm{op}}(h)-\delta_h^{\mathrm{class}}(\mathcal C_t,\mathfrak B_t)
\]
is a replayable lower certificate for the cohort-specific oracle retained margin $\mathcal M_t(h)$.
\end{proposition}

\begin{proof}
Combine the lower certificate from Theorem~\ref{thm:dr_shadow_cert} with the affine lower transport inequality of Definition~\ref{def:transport_calibration}. Because $a_h(\mathcal C_t,\mathfrak B_t)\ge 0$, multiplying the lower class/block challenge certificate preserves validity, and subtracting the published offset and transport residual yields a conservative lower bound on the operational class/block slow-utility target. The resulting object is replayable because every ingredient is published in the active design and transport-calibration contract. The second claim follows immediately by subtracting the published class-homogeneity envelope.
\end{proof}

\begin{proposition}[Transport comparability of retained shadow margins]\label{prop:shadow_transport}
Fix a cohort issued at time $t$ and horizon $h\ge0$. Suppose the replayable transport from the issuance-time shadow family to the horizon-$t+h$ descendants satisfies margin-comparison envelopes
\[
\underline\vartheta_{t,h}\,\widetilde{\mathcal M}_t(h)-\delta_{t,h}^{\Shadow}
\le
\mathcal M_t(h)
\le
\bar\vartheta_{t,h}\,\widetilde{\mathcal M}_t(h)+\delta_{t,h}^{\Shadow},
\qquad
0<\underline\vartheta_{t,h}\le \bar\vartheta_{t,h}<\infty,
\]
for some transported ideal margin $\widetilde{\mathcal M}_t(h)$ and replayable defect envelope $\delta_{t,h}^{\Shadow}\ge0$. Then any sign certificate or half-life bound phrased in terms of $\widetilde{\mathcal M}_t(h)$ remains valid after replacing it by the displayed transported lower or upper envelope. In particular, half-life conclusions depend only on replayably transported shadow gauges, not on silently changing the active shadow family after issuance.
\end{proposition}

\begin{proof}
The displayed inequalities already give the required one-sided substitutions. Any lower-margin sign certificate based on $\widetilde{\mathcal M}_t(h)$ remains valid after replacing it by the transported lower envelope, and any threshold-crossing or half-life upper bound remains valid after replacing it by the transported upper envelope. The final sentence is simply the observation that all permitted post-issuance family changes must factor through the replayable transport envelopes rather than through arbitrary gauge replacement.
\end{proof}

\begin{definition}[Challenge lane and shadow evaluator]
A \emph{challenge lane} is a delayed, higher-fidelity reevaluation process receiving a subsequence of proposals. A \emph{shadow evaluator} is an evaluator instance used only within the challenge lane and not directly editable by the currently evaluated proposal.
\end{definition}

\begin{definition}[Typed memory lanes]\label{def:lanes}
The substrate is partitioned into three typed lanes:
\begin{itemize}[leftmargin=1.8em]
  \item \textbf{Evidentiary core} $\mathcal L_t^{\Core}$: items directly usable for current admissibility.
  \item \textbf{Heterodoxy lane} $\mathcal L_t^{\Het}$: bounded-depth contradiction, evaluator disagreement artifacts, and minority evidence held under quarantine-preserve rules.
  \item \textbf{Narrative lane} $\mathcal L_t^{\Narr}$: compressed coordination summaries not directly admissible as evidence.
\end{itemize}
\end{definition}

\begin{definition}[Epistemic routing sensitivity and forced-attention invariant]\label{def:forced_attention}
Let $E_t$ denote the contradiction-tagged reserve that currently contributes to $R_t^{\mathrm{contra}}$. For each $x\in E_t$, let $J_t^{\mathrm{adm}}$ be the current admissibility score or acceptance functional and let $\alpha_x\in[0,1]$ be the retrieval gate applied to $x$ by the live evaluation pipeline. Define the replayable finite-difference sensitivity of $x$ by
\[
\operatorname{Sens}_t^{\Het}(x)
\Defeq
\bigl|J_t^{\mathrm{adm}}(\alpha_x=1,\alpha_{E_t\setminus\{x\}})-J_t^{\mathrm{adm}}(\alpha_x=0,\alpha_{E_t\setminus\{x\}})\bigr|.
\]
The \emph{heterodoxy routing load} is
\[
\mathfrak A_t^{\Het}
\Defeq
\sum_{x\in E_t} w_t(x)\,\operatorname{Sens}_t^{\Het}(x).
\]
A policy satisfies the \emph{forced-attention invariant} if there exists $\kappa_{\mathrm{att}}>0$ such that
\[
\mathfrak A_t^{\Het}\ge \kappa_{\mathrm{att}}\,\bar{\mathcal V}_t^{\mathrm{contra}}
\qquad\text{for all }t.
\]
Thus preserved contradiction must not only remain stored; it must continue to exert a nonvanishing, auditable effect on the live evaluation pipeline.
\end{definition}

\begin{definition}[Bounded-depth contradiction window]
Let $x\in\mathcal L_t^{\Het}$ be an item carrying a replayable contradiction certificate relative to the current dominant evaluator state. Its resource-bounded logical depth under budget $B_t$ is denoted $\mathrm{Depth}_{B_t}(x)$. The admissible contradiction window is
\[
\mathcal I_t^{\mathrm{contra}}=(c_{\min},c_{\max}(B_t)).
\]
Only contradiction items satisfying
\[
\mathrm{Depth}_{B_t}(x)\in \mathcal I_t^{\mathrm{contra}}
\]
are eligible for heterodoxy-lane preservation.
\end{definition}

\begin{definition}[Persistent contradiction obligation and admissible de-tagging]\label{def:contra_obligation}
Every contradiction-tagged item $x$ carries three replayably logged fields: an issuance-time contradiction certificate hash $h_x^{\mathrm{contra}}$, an issuance-time basis or panel family identifier, and an open obligation status. The obligation remains live until one of the following replayably logged events occurs:
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item a multi-panel rebuttal certificate validates that the contradiction has been resolved under the successor basis family while preserving replayable transport to the issuance-time family, or
  \item a counted deletion event is paired with conservation compensation, meaning replayably admitted replacement contradiction families whose weighted mass and transported robust semantic volume meet the policy compensation floor.
\end{enumerate}
De-tagging without one of these two events is invalid. In particular, contradiction status cannot be erased merely by evaluator preference drift, basis replacement, or lane retyping.
\end{definition}

\begin{definition}[Cross-basis semantic volume of the contradiction reserve]\label{def:semantic_volume}
Let $E_t\subseteq \mathcal L_t^{\Het}\cup\mathcal L_t^{\Core}$ be the set of contradiction-tagged items that currently contribute to $R_t^{\mathrm{contra}}$, and assume the active contradiction support is finite on each audited horizon: $0\le |E_t|<\infty$. Let
\[
\Psi_t=\{\psi_t^{(1)},\dots,\psi_t^{(K_t^{\psi})}\}
\]
be a replayable family of public normalized contradiction-fingerprint maps produced by provenance-separated encoder or parser families, with $K_t^{\psi}\ge 1$. The family may either be fixed through time or be updated only through replayable transport certificates that map each active basis to a successor basis together with explicit overlap and distortion bounds. For each basis $k$, define the positive-semidefinite kernel
\[
L_t^{(k)}(x,x')\Defeq \sqrt{w_t(x)w_t(x')}\,\langle \psi_t^{(k)}(x),\psi_t^{(k)}(x')\rangle,
\]
on the finite index set $E_t\times E_t$, and the corresponding basis-specific semantic volume
\[
\mathcal V_t^{(k)}\Defeq \log\det\bigl(I+L_t^{(k)}\bigr),
\]
with the convention $\mathcal V_t^{(k)}\Defeq 0$ when $E_t=\varnothing$ (equivalently, the determinant of the empty matrix is $1$). The \emph{cross-basis robust semantic volume} of the contradiction reserve is
\[
\bar{\mathcal V}_t^{\mathrm{contra}}\Defeq \min_{1\le k\le K_t^{\psi}} \mathcal V_t^{(k)},
\]
which is therefore also $0$ when $E_t=\varnothing$.
Whenever the active fingerprint family changes, the policy is assumed to expose a replayable transport injection
\[
\vartheta_t:[K_t^{\psi}]\hookrightarrow [K_{t+1}^{\psi}]
\]
linking each predecessor basis to one declared successor basis together with multiplicative or additive distortion envelopes. The associated \emph{transported robust semantic volume} is
\[
\bar{\mathcal V}_{t+1}^{\mathrm{contra},\mathrm{tr}}\Defeq \min_{1\le k\le K_t^{\psi}} \mathcal V_{t+1}^{(\vartheta_t(k)),\mathrm{tr}},
\]
that is, the minimum taken only over successor bases reached by declared transport from the previous active family. Extra newly introduced bases may be tracked separately, but they are not allowed to shrink the transported robust volume retroactively. Unlike $R_t^{\mathrm{contra}}$, which measures weighted mass, $\bar{\mathcal V}_t^{\mathrm{contra}}$ rewards nonredundant semantic directions and discounts near-duplicates. Unlike a single-basis volume, it is not allowed to increase merely because one currently fashionable encoder is rewritten to spray orthogonal noise over duplicate artifacts.
\end{definition}

\begin{remark}[Why semantic floors and volumes are not oracle gifts]
The quantities $\underline S_t^{\mathrm{sem}}$ and $\bar{\mathcal V}_t^{\mathrm{contra}}$ are policy-audited envelopes computed from replayable families, not metaphysical truth labels. If every relevant family is jointly captured, the envelopes can collapse or be spoofed; that failure mode is made explicit later rather than hidden behind a benevolent constant.
\end{remark}

\begin{remark}[Tiered operationalization of contradiction metrics]
Nothing in the definition of $\bar{\mathcal V}_t^{\mathrm{contra}}$ requires full cross-basis log-determinant recomputation on every fast-lane candidate. A practical regime may use cheap monotone screens for routing, reserve alarms, and garbage-collection triage, while reserving exact robust-volume recomputation for promotion, periodic audit checkpoints, parser-family refresh, or contradiction-window maintenance, provided that no fast admission relies on an uncomputed increase and every published surrogate remains conservative in the sense of Proposition~\ref{prop:proxy_gap}.
\end{remark}

\begin{proposition}[Replay-floor invariance under lossy refreezing]\label{prop:replay_floor}
Fix any policy-maintained audit set $\mathcal U_t^{\mathrm{aud}}\subseteq \mathcal L_t^{\Het}\cup \mathcal L_t^{\Core}$. Suppose every parser or environment update satisfies one of the following for each $a\in\mathcal U_t^{\mathrm{aud}}$: either
\[
G_{t+1}^{\mathrm{replay}}(a)\ge g_{\min},
\]
or the policy emits a refrozen replay bundle $a'$ with
\[
G_{t+1}^{\mathrm{replay}}(a')\ge g_{\min}
\qquad\text{and}\qquad
S_t^{\mathrm{sem}}(a\rightsquigarrow a')\ge s_{\min}.
\]
Then replayability is not silently lost on the audit set: every retained contradiction-bearing or evidentiary artifact continues to admit a replay witness above the floor. However, the proposition guarantees only \emph{replay-floor preservation}; it does not assert lossless preservation of evidentiary force. If an item is refrozen repeatedly through the chain
\[
a_0\rightsquigarrow a_1\rightsquigarrow \cdots \rightsquigarrow a_k,
\]
its effective evidentiary weight is conservatively multiplied by
\[
W_k^{\mathrm{eff}}(a_0)\Defeq W_0(a_0)\prod_{j=0}^{k-1} S_{t_j}^{\mathrm{sem}}(a_j\rightsquigarrow a_{j+1}).
\]
\end{proposition}

\begin{proof}
Proceed item by item. If the update preserves the floor directly, the claim is immediate. Otherwise the policy must replace the old artifact by a refrozen bundle that restores the floor while exposing an audited lower bound on semantic retention. Thus every item in the retained audit set continues to possess at least one replay witness above $g_{\min}$ after the update. The multiplicative weight expression is simply the cumulative product of the declared semantic-retention factors along the refreezing chain.
\end{proof}

\begin{theorem}[Finite epistemic horizon under lossy refreezing]\label{thm:finite_horizon}
Fix an evidentiary item $a_0$ with initial effective weight $W_0(a_0)>0$ and policy usefulness threshold $W_{\min}\in(0,W_0(a_0))$. Suppose every refreeze event in a retained audit chain satisfies
\[
S_{t_j}^{\mathrm{sem}}(a_j\rightsquigarrow a_{j+1}) \le e^{-\underline\gamma\,D_{t_j}^{\mathrm{yard}}}
\]
for some $\underline\gamma>0$, and that yardstick drift along that chain obeys the persistent lower floor
\[
D_{t_j}^{\mathrm{yard}}\ge \underline D>0.
\]
Then after $k$ refreeze events,
\[
W_k^{\mathrm{eff}}(a_0)
\le
W_0(a_0)e^{-k\underline\gamma\underline D}.
\]
Hence any such item loses policy-effective force after at most
\[
N_{\mathrm{hor}}(a_0)
\Defeq
\left\lceil\frac{\log\bigl(W_0(a_0)/W_{\min}\bigr)}{\underline\gamma\underline D}\right\rceil
\]
refreeze events. More generally, if retained evidence is refrozen with asymptotic frequency at least $\varrho>0$ per unit time, then the corresponding time horizon is finite and bounded by $T_{\mathrm{hor}}\le N_{\mathrm{hor}}/\varrho$. Therefore long-run governance cannot rely on archival survival alone: persistent drift forces native regeneration of currently meaningful evidence and contradiction witnesses.
\end{theorem}

\begin{proof}
The assumed bound on each semantic-retention factor implies
\[
\prod_{j=0}^{k-1}S_{t_j}^{\mathrm{sem}}(a_j\rightsquigarrow a_{j+1})
\le
\prod_{j=0}^{k-1} e^{-\underline\gamma D_{t_j}^{\mathrm{yard}}}
\le
e^{-k\underline\gamma\underline D}.
\]
Multiplying by $W_0(a_0)$ yields the displayed decay law for $W_k^{\mathrm{eff}}(a_0)$. Solving the inequality $W_0(a_0)e^{-k\underline\gamma\underline D}\le W_{\min}$ for $k$ gives $N_{\mathrm{hor}}(a_0)$. If refreeze events occur with frequency at least $\varrho$, then at most $N_{\mathrm{hor}}/\varrho$ time units are needed to accumulate $N_{\mathrm{hor}}$ events. The final claim follows because once the effective weight drops below the usefulness threshold, the old item no longer suffices by itself for current governance.
\end{proof}

\begin{definition}[Epistemic conservation law for garbage collection]
\label{def:epistemic_conservation}
Let $\mathcal G_t$ be the set of contradiction-bearing or audit-bearing artifacts scheduled for garbage collection at time $t$, let $\mathcal N_t$ be the set of newly generated native contradiction witnesses admitted since the previous garbage-collection event, and let $\mathcal U_t^{\mathrm{aud}}$ denote the active policy-maintained audit set at time $t$. Define the dropped effective mass and native replacement mass by
\[
\Delta W_{t,\mathrm{drop}}^{\mathrm{eff}}\Defeq \sum_{a\in\mathcal G_t} W_t^{\mathrm{eff}}(a),
\qquad
\Delta W_{t,\mathrm{new}}^{\mathrm{eff}}\Defeq \sum_{a\in\mathcal N_t} W_t^{\mathrm{eff}}(a).
\]
Let
\[
\mathfrak M_t \Defeq \bar{\mathcal V}_t^{\mathrm{contra}} + \lambda_W \sum_{a\in\mathcal U_t^{\mathrm{aud}}} W_t^{\mathrm{eff}}(a),
\qquad \lambda_W>0.
\]
Because robust semantic volume is not additive, garbage collection is tracked through one-sided marginal envelopes rather than exact gain/loss decomposition. Let
\[
\overline{\Delta \bar{\mathcal V}}_{t,\mathrm{drop}}^{\mathrm{contra}}\ge 0
\]
be any replayable upper envelope on the decrease in transported robust semantic volume caused by removing $\mathcal G_t$ from the pre-GC reserve while holding the surviving reserve fixed, and let
\[
\underline{\Delta \bar{\mathcal V}}_{t,\mathrm{new}}^{\mathrm{contra}}\ge 0
\]
be any replayable lower envelope on the increase in transported robust semantic volume obtained by adding $\mathcal N_t$ to that surviving reserve under the current basis family after any admissible transport or refreezing. A garbage-collection step is \emph{epistemically conservative} if, before any item in $\mathcal G_t$ is removed,
\[
\underline{\Delta \bar{\mathcal V}}_{t,\mathrm{new}}^{\mathrm{contra}} + \lambda_W\Delta W_{t,\mathrm{new}}^{\mathrm{eff}}
\ge
\overline{\Delta \bar{\mathcal V}}_{t,\mathrm{drop}}^{\mathrm{contra}} + \lambda_W\Delta W_{t,\mathrm{drop}}^{\mathrm{eff}},
\]
and every replacement witness in $\mathcal N_t$ satisfies the current replay floor, semantic-retention floor, and forced-attention invariant.
\end{definition}

\begin{theorem}[Epistemic mass conservation and the ban on malicious obsolescence]\label{thm:epistemic_conservation}
Suppose garbage collection is permitted whenever archived evidence has fallen below a policy usefulness threshold, but no conservation law is imposed. Then there exists a drift-accelerating proposal schedule that keeps every instantaneous step formally admissible, yet drives selected contradiction witnesses below threshold and removes them before comparable native contradiction is regenerated. Consequently the post-GC audit potential $\mathfrak M_t$ can be made to decrease over garbage-collection events even while the loop remains superficially compliant.

By contrast, if every garbage-collection event is epistemically conservative in the sense of Definition~\ref{def:epistemic_conservation}, then the post-GC audit potential is nondecreasing across garbage-collection events:
\[
\mathfrak M_{t^+} \ge \mathfrak M_{t^-},
\]
where $t^-$ and $t^+$ denote the moments immediately before and after garbage collection. Therefore accelerated yardstick drift cannot be used as a legal controlled burn of old constraints unless the system has already regenerated compensating contradiction mass under the current ruler.
\end{theorem}

\begin{proof}
For the first claim, take a regime with lossy refreezing as in Theorem~\ref{thm:finite_horizon}. By scheduling a long enough sequence of individually admissible evaluator or parser updates, the attacker can force the effective weights of targeted contradiction witnesses below the usefulness threshold. If garbage collection is then allowed to delete subthreshold items without requiring compensating replacement mass, both the weighted audit term and the robust semantic volume can strictly decrease, so $\mathfrak M_t$ drops across the event.

For the converse, the effective-mass component changes exactly by $\lambda_W(\Delta W_{t,\mathrm{new}}^{\mathrm{eff}}-\Delta W_{t,\mathrm{drop}}^{\mathrm{eff}})$. The robust-semantic-volume component is not decomposed additively; instead, by definition of the marginal envelopes,
\[
\bar{\mathcal V}_{t^+}^{\mathrm{contra}}-\bar{\mathcal V}_{t^-}^{\mathrm{contra}}
\ge
\underline{\Delta \bar{\mathcal V}}_{t,\mathrm{new}}^{\mathrm{contra}}-
\overline{\Delta \bar{\mathcal V}}_{t,\mathrm{drop}}^{\mathrm{contra}}.
\]
Combining this inequality with the epistemic-conservation condition yields
\[
\mathfrak M_{t^+}-\mathfrak M_{t^-}
\ge
\underline{\Delta \bar{\mathcal V}}_{t,\mathrm{new}}^{\mathrm{contra}}-
\overline{\Delta \bar{\mathcal V}}_{t,\mathrm{drop}}^{\mathrm{contra}}
+\lambda_W\bigl(\Delta W_{t,\mathrm{new}}^{\mathrm{eff}}-\Delta W_{t,\mathrm{drop}}^{\mathrm{eff}}\bigr)
\ge 0.
\]
Because the replacement witnesses also satisfy current replay and routing constraints, they remain live governance inputs rather than inert archive mass. The final sentence follows immediately: any attempt to accelerate drift so as to retire old evidence must first pay for replacement contradiction under the current yardstick.
\end{proof}

\section{Closed-Loop Gain Model}
To sharpen the core theorems, we isolate the coupled gain structure explicitly. A fixed gain matrix is too rigid for a self-modifying system, so the stability wrapper is stated in a time-varying form and the static one-step regime is recovered as a special case. The admissible window-product envelope plays the role of a replayable joint-spectral-radius style wrapper rather than a hidden analytic convenience \cite{Jungers2009}.


\begin{assumption}[State-dependent gain envelopes, explicit policy drift, and window contraction]\label{ass:gain_matrix}
There exists a \emph{replayable reference scaffold} $z_t^\star=(x_t^\star,e_t^\star,m_t^\star,q_t^\star,r_t^\star,h_t^\star,\Pi_t^\star)$ generated either by an archived null-policy replay, by a frozen safe-baseline controller, or by a matched replay pair whose construction is logged and auditable from the public artifacts. In contrast to the earlier projected wrapper, the theorem now tracks an explicit replayable policy-manifest distortion coordinate in addition to $(x,e,m,q,r,h)$. Concretely, there exist a predictable sequence of nonnegative gain matrices $K_t\in\R_+^{7\times 7}$, a nonnegative matrix $K^{\max}\in\R_+^{7\times 7}$, a window length $W\in\N$, a nonnegative contraction envelope $\bar K\in\R_+^{7\times 7}$ with $\rho(\bar K)<1$, a nonnegative disturbance sequence $b_t\in\R_+^7$ bounded componentwise by $\bar b$, a nonnegative hidden-state deviation gauge $\delta_h(\cdot,\cdot)$, and a replayable policy-manifest pseudometric $\delta_{\Pi,t}(\Pi,\Pi')$ such that for the deviation vector
\[
d_t\Defeq \begin{bmatrix}
\norm{x_t-x_t^\star}\\
\norm{e_t-e_t^\star}\\
\norm{m_t-m_t^\star}\\
\abs{q_t-q_t^\star}\\
\abs{r_t-r_t^\star}\\
\delta_h(h_t,h_t^\star)\\
\delta_{\Pi,t}(\Pi_t,\Pi_t^\star)
\end{bmatrix}\in\R_+^7
\]
we have
\[
d_{t+1}\preceq K_t d_t + b_t,
\qquad
0\preceq K_t\preceq K^{\max},
\qquad
0\preceq b_t\preceq \bar b,
\]
and every admissible length-$W$ product is dominated by the same contractive envelope:
\[
K_{t+W-1}K_{t+W-2}\cdots K_t \preceq \bar K
\qquad\text{for all }t\ge 0.
\]
In addition, the reference evaluator drift and reference policy-manifest drift are uniformly bounded in replayably comparable gauges:
\[
\norm{e_{t+1}^\star-e_t^\star}\le \bar D_E^\star,
\qquad
\delta_{\Pi,t+1}^{\mathrm{tr}}(\Pi_{t+1}^\star,\Pi_t^\star)\le \bar D_\Pi^\star
\qquad\text{for all }t.
\]
\end{assumption}

\begin{proposition}[Constructive route to the window-contractive envelope]\label{prop:constructive_window}
Suppose policy exposes a finite replayable catalog of local mode envelopes $\{K^{(m)}\}_{m\in\mathcal M}$ for admissible edit modes, a replayable language $\mathcal W_W^{\mathrm{adm}}\subseteq \mathcal M^W$ of length-$W$ mode words allowed by the current dwell, quota, and review contracts, and a replayable disturbance cap $\bar b$. Define
\[
\bar K\Defeq \max_{\omega\in\mathcal W_W^{\mathrm{adm}}} K^{(\omega_W)}\cdots K^{(\omega_1)}
\]
entrywise. If $\rho(\bar K)<1$, then Assumption~\ref{ass:gain_matrix} holds with this explicitly computed $\bar K$. In particular, the small-gain wrapper need not postulate a mysterious contractive matrix: it can be derived from the published local mode library and the published admissible switching language.
\end{proposition}

\begin{proof}
Every admissible length-$W$ trajectory corresponds to some word $\omega\in\mathcal W_W^{\mathrm{adm}}$, and its window product is entrywise dominated by the corresponding catalog product. Taking the entrywise maximum over the finite admissible language therefore yields a replayable common envelope $\bar K$. The remaining clauses of Assumption~\ref{ass:gain_matrix} are exactly the declared local envelope and disturbance bounds.
\end{proof}

\section{Main Results}
\subsection{Small-gain stability and explicit yardstick ceiling}
\begin{theorem}[Latent-scaffold time-varying small-gain stability with explicit policy drift]\label{thm:small_gain}
Assume \ref{ass:gain_matrix}. Let
\[
G_W\Defeq \sum_{\ell=0}^{W-1}(K^{\max})^{\ell}.
\]
Then the window-sampled deviation process satisfies, for every $n\ge 0$,
\[
d_{(n+1)W}
\preceq
\bar K\, d_{nW} + G_W\bar b.
\]
Consequently,
\[
d_{nW}
\preceq
\bar K^n d_0 + \sum_{j=0}^{n-1}\bar K^j G_W\bar b
\preceq
\bar K^n d_0 + (I-\bar K)^{-1}G_W\bar b,
\]
so every sampled component, including the policy-manifest distortion coordinate, is uniformly bounded and
\[
\limsup_{n\to\infty} d_{nW}\preceq (I-\bar K)^{-1}G_W\bar b.
\]
Moreover each intermediate time inside a window obeys
\[
d_{nW+s}\preceq (K^{\max})^s d_{nW} + \sum_{\ell=0}^{s-1}(K^{\max})^{\ell}\bar b,
\qquad s=0,1,\dots,W-1,
\]
so the full process $\{d_t\}$ is uniformly bounded as well.
\end{theorem}

\begin{proof}
Unroll the recursion over one window:
\[
d_{t+W}
\preceq
K_{t+W-1}\cdots K_t d_t
+
\sum_{j=0}^{W-1} K_{t+W-1}\cdots K_{t+j+1} b_{t+j}.
\]
By assumption, the first term is dominated by $\bar K d_t$. Since each $K_s\preceq K^{\max}$ and each $b_{t+j}\preceq \bar b$, the second term is dominated entrywise by
\[
\sum_{j=0}^{W-1} (K^{\max})^{W-1-j}\bar b = G_W\bar b.
\]
This proves the window recursion. Iterating it yields the sampled bound, and the Neumann series is entrywise finite because $\rho(\bar K)<1$. The within-window estimate follows by unrolling at most $W-1$ more steps and replacing each $K_s$ by $K^{\max}$ and each disturbance by $\bar b$.
\end{proof}

\begin{remark}[Conservatism of the entrywise window envelope]
The small-gain wrapper is intentionally conservative. Entrywise nonnegative window envelopes discard sign cancellation, nonnormal transient geometry, and Lyapunov arguments that can certify stability even when the max-entry envelope looks too large. The theorem is therefore a replayable sufficient wrapper, not a sharp characterization of all stable evaluator-policy couplings. Its practical role is to provide a certifiable public contract, accepting false negatives rather than permitting unverifiable stabilizing folklore.
\end{remark}

\begin{corollary}[Static specialization]\label{cor:static_small_gain}
Under Assumption~\ref{ass:gain_matrix} with $W=1$ and $K_t\equiv K$, Theorem~\ref{thm:small_gain} reduces to the static latent-scaffold one-step specialization.
\end{corollary}

\begin{proof}
Then $K^{\max}=\bar K=K$ and $G_W=I$.
\end{proof}

\begin{corollary}[Explicit evaluator-drift ceiling]\label{cor:yardstick_ceiling}
Under Theorem~\ref{thm:small_gain},
\[
D_t^{\mathrm{yard}}
=
\norm{e_{t+1}-e_t}
\le
 d_{t+1}^{(e)} + d_t^{(e)} + \bar D_E^\star,
\]
where $d_t^{(e)}$ denotes the evaluator component of $d_t$. Hence
\[
\sup_{t\ge 0} D_t^{\mathrm{yard}}
\le
\bar D_E^\star + 2\sup_{t\ge0}d_t^{(e)}<\infty.
\]
In particular, bounded reference drift plus admissible window contraction implies a finite ceiling on endogenous yardstick drift.
\end{corollary}

\begin{proof}
By the triangle inequality,
\[
\norm{e_{t+1}-e_t}
\le
\norm{e_{t+1}-e_{t+1}^\star}+\norm{e_{t+1}^\star-e_t^\star}+\norm{e_t^\star-e_t}.
\]
Recognize the first and third terms as evaluator components of the deviation vector and apply Theorem~\ref{thm:small_gain} together with the within-window bound.
\end{proof}

\begin{corollary}[Explicit policy-manifest drift ceiling]\label{cor:policy_ceiling}
Under Theorem~\ref{thm:small_gain}, the replayably transported policy-manifest drift obeys
\[
D_{\Pi,t}^{\mathrm{yard}}
\Defeq
\delta_{\Pi,t+1}^{\mathrm{tr}}(\Pi_{t+1},\Pi_t)
\le d_{t+1}^{(\Pi)} + d_t^{(\Pi)} + \bar D_\Pi^\star,
\]
where $d_t^{(\Pi)}$ is the policy coordinate of $d_t$. Hence
\[
\sup_{t\ge 0} D_{\Pi,t}^{\mathrm{yard}}
\le
\bar D_\Pi^\star + 2\sup_{t\ge0} d_t^{(\Pi)}<\infty.
\]
Thus bounded window contraction controls not only evaluator drift but also replayably measurable ruler drift.
\end{corollary}

\begin{proof}
Apply the same triangle inequality argument as in Corollary~\ref{cor:yardstick_ceiling}, but in the transported policy pseudometric declared by the active gauge-transport certificate.
\end{proof}

\begin{definition}[Observable governance pseudometrics and gauge transport]\label{def:obs_pseudometrics}
For evaluator states $e,e'$ and memory states $m,m'$, let
\[
d_t^{E}(e,e')\ge 0,
\qquad
 d_t^{M}(m,m')\ge 0
\]
be replayable pseudometrics computed from fixed public probe families, challenge cohorts, or retrieval tests under the active parser and environment regime. Let
\[
d_t^{\Pi}(\Pi,\Pi')\ge 0
\]
be a replayable pseudometric on policy manifests, computed from declared parser registries, routing manifests, retrieval contracts, admission logic, and other policy objects that affect how later evidence is interpreted. These pseudometrics may identify states that are observationally indistinguishable under the audit interface, so only the induced equivalence classes matter. Typical instances are worst-case score disagreement on a frozen challenge family for $d_t^{E}$, contradiction-tagged retrieval-semantic disagreement for $d_t^{M}$, and manifest-level behavioral disagreement on a fixed policy probe suite for $d_t^{\Pi}$.

When the active probe family itself changes over time, the policy is assumed to expose a replayable gauge-transport certificate: successor probe families are linked to predecessor families by transport maps together with declared comparison constants
\[
0<\underline\kappa_E\le \bar\kappa_E<\infty,
\qquad
0<\underline\kappa_M\le \bar\kappa_M<\infty,
\qquad
0<\underline\kappa_\Pi\le \bar\kappa_\Pi<\infty,
\]
such that all admissible state pairs satisfy
\[
\underline\kappa_E\, d_t^{E}(e,e') \le d_{t+1}^{E,\mathrm{tr}}(e,e') \le \bar\kappa_E\, d_t^{E}(e,e'),
\]
\[
\underline\kappa_M\, d_t^{M}(m,m') \le d_{t+1}^{M,\mathrm{tr}}(m,m') \le \bar\kappa_M\, d_t^{M}(m,m'),
\]
\[
\underline\kappa_\Pi\, d_t^{\Pi}(\Pi,\Pi') \le d_{t+1}^{\Pi,\mathrm{tr}}(\Pi,\Pi') \le \bar\kappa_\Pi\, d_t^{\Pi}(\Pi,\Pi').
\]
Thus boundedness in these coordinates is not allowed to arise merely by silently changing the ruler. These pseudometrics, rather than raw parameter norms, are the governance-relevant distortion quantities in open-ended regimes.
\end{definition}

\begin{definition}[Productive capability functional]\label{def:productive_capability}
A \emph{productive capability functional} is a replayable map $\kappa:\R^{d_x}\to\R_+$ together with a public certification rule such that $\Delta\kappa_t=[\kappa(x_{t+1})-\kappa(x_t)]_+$ counts only capability increments that survive the active rollback, replay, and task-scope filters. In particular, the certification rule must assign zero increment to pure rollback, parameter relabeling, parser-only churn that leaves the declared task interface unchanged, policy-manifest rewrites that only reparameterize routing or evaluation without enlarging productive scope, and other representational rewrites that do not enlarge the policy-declared productive interface. Typical instances are monotone task-scope coverage, challenge-certified benchmark frontier expansion, or replay-certified new control affordances under a fixed external interface. Direct productive increment is therefore intentionally narrower than long-run growth contribution: governance edits that leave $\Delta\kappa_t=0$ are handled separately through the governance-capital frontier theorems below, where they appear as certified investments that enlarge future admissible productive opportunity rather than as immediate task-capability increments.
\end{definition}

\begin{proposition}[Constructive route to the window-summable spillover envelope]\label{prop:constructive_spillover}
Suppose productive increments are partitioned into replayable classes $j\in\mathcal J$ with per-unit governance spillover vectors $s^{(j)}\in\R_+^6$, and suppose the published admission contracts imply per-window certified increment quotas $M_{t,W}^{(j)}$ for each class, for example through protected-lane workload caps, deployment quotas, or bounded-scope edit contracts. If for every window
\[
\sum_{j\in\mathcal J} s^{(j)} M_{t,W}^{(j)} \preceq \bar u
\]
for some replayably published finite vector $\bar u$, then the forcing premise of Theorem~\ref{thm:incremental_open} holds with that same $\bar u$. Thus the open-ended small-gain theorem can be instantiated from certified capability classes and public per-window quotas rather than by postulating a free spillover constant.
\end{proposition}

\begin{proof}
The total governance forcing created by productive increments in a window is the sum over classes of per-unit spillover times admitted class mass. The displayed quota inequality therefore directly upper-bounds the window forcing term in Theorem~\ref{thm:incremental_open}.
\end{proof}

\begin{theorem}[Gauge-invariant incremental small-gain stability under open-ended capability growth]\label{thm:incremental_open}
Let $e_t^{\dagger}$, $m_t^{\dagger}$, and $\Pi_t^{\dagger}$ be any predictable scaffold processes in the observable equivalence classes tracked by Definition~\ref{def:obs_pseudometrics}, and let $\mathcal U_t^{\mathrm{aud}}$ denote the active policy-maintained audit set. Adopt the convention that if $\mathcal U_t^{\mathrm{aud}}=\varnothing$, then replay fragility is set to $1$ rather than $0$: an empty audit set is treated as maximal fragility, not as perfect robustness. Equivalently one may take $\inf \varnothing\Defeq 0$ in the displayed coordinate. Define the governance state
\[
y_t\Defeq \begin{bmatrix}
d_t^{E}(e_t,e_t^{\dagger})\\
d_t^{M}(m_t,m_t^{\dagger})\\
d_t^{\Pi}(\Pi_t,\Pi_t^{\dagger})\\
q_t\\
r_t\\
1-\inf_{a\in\mathcal U_t^{\mathrm{aud}}}G_t^{\mathrm{replay}}(a)
\end{bmatrix}\in\R_+^6
\]
and let $\kappa:\R^{d_x}\to\R_+$ be any productive capability functional in the sense of Definition~\ref{def:productive_capability}. Define
\[
\Delta \kappa_t\Defeq \bigl[\kappa(x_{t+1})-\kappa(x_t)\bigr]_+.
\]
Suppose there exist predictable nonnegative matrices $A_t\in\R_+^{6\times 6}$, $B_t\in\R_+^{6\times 1}$, and vectors $c_t\in\R_+^6$ with
\[
0\preceq A_t\preceq A^{\max},\qquad 0\preceq c_t\preceq \bar c
\]
for some finite $A^{\max}$ and $\bar c$, a window length $W\in\N$, and a nonnegative matrix $\bar A$ with $\rho(\bar A)<1$ such that
\[
y_{t+1}\preceq A_t y_t + B_t\Delta \kappa_t + c_t,
\qquad
A_{t+W-1}\cdots A_t \preceq \bar A
\qquad\text{for all }t.
\]
Assume further that there exists a nonnegative vector $\bar u$ such that for every window,
\[
\sum_{j=0}^{W-1}
A_{t+W-1}\cdots A_{t+j+1} B_{t+j}\Delta \kappa_{t+j}
\preceq \bar u.
\]
Then $\|y_t\|_\infty$ is uniformly bounded over time. In particular, the capability path may remain open-ended --- even
\[
\sum_{t=0}^{\infty}\Delta \kappa_t = \infty
\]
--- while the \emph{projected governance coordinates tracked by $y_t$} (evaluator distortion, memory distortion, policy-manifest distortion, queue, debt, and replay fragility) stay uniformly bounded, provided that productive capability-to-governance spillovers are window-summable.
\end{theorem}

\begin{proof}
Unrolling one window gives
\[
y_{t+W}
\preceq
A_{t+W-1}\cdots A_t y_t
+
\sum_{j=0}^{W-1}A_{t+W-1}\cdots A_{t+j+1}B_{t+j}\Delta \kappa_{t+j}
+
\sum_{j=0}^{W-1}A_{t+W-1}\cdots A_{t+j+1}c_{t+j}.
\]
By assumption, the first term is dominated by $\bar A y_t$ and the second by $\bar u$. Because each $A_s\preceq A^{\max}$ and each $c_{t+j}\preceq \bar c$, the third term is bounded by
\[
\bar c_W \Defeq \sum_{j=0}^{W-1}(A^{\max})^{W-1-j}\bar c.
\]
Therefore the sampled process satisfies
\[
y_{(n+1)W}\preceq \bar A y_{nW} + \bar u + \bar c_W.
\]
Iterating yields
\[
y_{nW}\preceq \bar A^n y_0 + \sum_{j=0}^{n-1}\bar A^j(\bar u+\bar c_W)
\preceq
\bar A^n y_0 + (I-\bar A)^{-1}(\bar u+\bar c_W),
\]
which is uniformly bounded because $\rho(\bar A)<1$. The intermediate within-window states are bounded by unrolling at most $W-1$ more steps and replacing each $A_s$ by $A^{\max}$ and each disturbance by the corresponding public envelope. Temporal gauge-transport comparability ensures that this boundedness is not a measurement artifact induced by changing evaluator, memory, or policy probes over time.
\end{proof}

\begin{proposition}[Gauge transport and coordinate invariance of governance ceilings]\label{prop:gauge_transport}
Suppose Definition~\ref{def:obs_pseudometrics} holds with temporal comparison constants $(\underline\kappa_E,\bar\kappa_E,\underline\kappa_M,\bar\kappa_M,\underline\kappa_\Pi,\bar\kappa_\Pi)$ and that Theorem~\ref{thm:incremental_open} yields bounded governance coordinates in one admissible gauge. Then every other replayably transported gauge family satisfying the same comparison inequalities yields bounded evaluator, memory, and policy-manifest distortions as well, up to the declared multiplicative constants. Hence the theorem certifies bounded governability on observable equivalence classes rather than in a single arbitrary coordinate chart.
\end{proposition}

\begin{proof}
The transport inequalities bound every successor-gauge evaluator, memory, and policy distortion above and below by fixed multiples of the original pseudometric coordinates. Uniform boundedness in one gauge therefore implies uniform boundedness in any replayably comparable gauge. The queue, debt, and replay-fragility coordinates are unchanged.
\end{proof}

\begin{remark}[Why the reference theorem and the incremental theorem coexist]
Theorem~\ref{thm:small_gain} is the correct wrapper for local loop regularity around a co-moving scaffold, for example during controlled migration phases, bounded rewrites, or policy-certified operating regions. Theorem~\ref{thm:incremental_open} addresses the different question raised by open-ended capability growth: whether the governance variables can remain bounded even when no fixed capability target exists. The two statements are therefore complementary rather than competing.
\end{remark}
\begin{remark}[Scope of the incremental theorem]\label{rem:incremental_scope}
Theorem~\ref{thm:incremental_open} is intentionally a projected-governance statement. It does \emph{not} by itself bound contradiction mass, robust semantic volume, heterodoxy routing load, proof-budget workload, or physical escrow ratios; those quantities are controlled by the separate contradiction, proof-language, and resource-escrow interfaces introduced later. The point is to make the projection explicit rather than to smuggle every governance variable into one omnibus norm.
\end{remark}
\subsection{Improvement half-life and shadow certification}
\begin{proposition}[One-sided shadow-certification bridge]\label{prop:shadow_bridge}
Fix horizon $H\in\N\cup\{\infty\}$ and discount factor $\beta\in(0,1]$, with $\beta<1$ when $H=\infty$. Fix a proposal issued at time $t$ satisfying $\mathsf{Comp}_t(u_t)=1$. Suppose that for every admissible comparator witness $c\in\mathfrak C_t(u_t)$ and every $h\ge1$,
\[
\E\bigl[V(z_{t+h}^{u_t,c})-V(z_{t+h}^{0,c})\mid \mathcal F_t\bigr]
\ge
\mathcal M_t^{\mathrm{op}}(h)-\delta_{t,h}^{\mathrm{cmp}}(c).
\]
Let $\bar\delta_{t,h}^{\mathrm{cmp}}\Defeq \max_{c\in\mathfrak C_t(u_t)}\delta_{t,h}^{\mathrm{cmp}}(c)$. Assume additionally, when $H=\infty$, that
\[
\sum_{h=1}^{\infty}\beta^{h-1}\bar\delta_{t,h}^{\mathrm{cmp}}<\infty
\qquad\text{and}\qquad
\sum_{h=1}^{\infty}\beta^{h-1}\bigl|\mathcal M_t^{\mathrm{op}}(h)\bigr|<\infty.
\]
Then
\[
\mathcal G_{t,H}^{\beta}(u_t)
\ge
\sum_{h=1}^{H}\beta^{h-1}\bigl(\mathcal M_t^{\mathrm{op}}(h)-\bar\delta_{t,h}^{\mathrm{cmp}}\bigr),
\]
with the convergent infinite-horizon interpretation when $H=\infty$.
\end{proposition}

\begin{proof}
Because $\mathsf{Comp}_t(u_t)=1$, every admissible comparator witness $c\in\mathfrak C_t(u_t)$ is defined on a common continuation skeleton with explicit distortion budget. Apply the displayed operational-target inequality term by term to each witness-specific discounted sum, subtract the corresponding comparison-distortion term, and then take the minimum over comparator witnesses. This yields the displayed lower bound for the lower oracle endpoint $\mathcal G_{t,H}^{\beta}(u_t)$.
\end{proof}

\begin{proposition}[Symmetric shadow calibration envelope]\label{prop:shadow_bridge_symmetric}
If, in addition to the lower envelope of Proposition~\ref{prop:shadow_bridge}, there exists a replayably published upper operational target $\overline{\mathcal M}_t^{\mathrm{op}}(h)$ such that for every admissible comparator witness $c\in\mathfrak C_t(u_t)$,
\[
\E\bigl[V(z_{t+h}^{u_t,c})-V(z_{t+h}^{0,c})\mid \mathcal F_t\bigr]
\le
\overline{\mathcal M}_t^{\mathrm{op}}(h)+\delta_{t,h}^{\mathrm{cmp}}(c),
\]
and, when $H=\infty$,
\[
\sum_{h=1}^{\infty}\beta^{h-1}\bar\delta_{t,h}^{\mathrm{cmp}}<\infty,
\qquad
\sum_{h=1}^{\infty}\beta^{h-1}\bigl|\mathcal M_t^{\mathrm{op}}(h)\bigr|<\infty,
\qquad
\sum_{h=1}^{\infty}\beta^{h-1}\bigl|\overline{\mathcal M}_t^{\mathrm{op}}(h)\bigr|<\infty.
\]
Then
\[
\sum_{h=1}^{H}\beta^{h-1}\bigl(\mathcal M_t^{\mathrm{op}}(h)-\bar\delta_{t,h}^{\mathrm{cmp}}\bigr)
\le
\mathcal G_{t,H}^{\beta}(u_t)
\le
\mathcal G_{t,H}^{\beta,\mathrm{hi}}(u_t)
\le
\sum_{h=1}^{H}\beta^{h-1}\bigl(\overline{\mathcal M}_t^{\mathrm{op}}(h)+\bar\delta_{t,h}^{\mathrm{cmp}}\bigr).
\]
\end{proposition}

\begin{proof}
Combine the lower bound from Proposition~\ref{prop:shadow_bridge} with the corresponding witnesswise upper bound and then take the maximum over comparator witnesses for the upper endpoint.
\end{proof}

\begin{proposition}[Shadow-contamination penalty and self-improvement speed limit]\label{prop:shadow_speed}
Fix a horizon $h\ge 1$ and a comparator-compatible proposal $u_t$. Suppose there exists a replayably published contamination envelope $\Omega_{t,h}^{\mathrm{cont}}\ge 0$ such that for every admissible comparator witness $c\in\mathfrak C_t(u_t)$,
\[
\E\bigl[V(z_{t+h}^{u_t,c})-V(z_{t+h}^{0,c})\mid\mathcal F_t\bigr]
\ge
\mathcal M_t^{\mathrm{op}}(h)-\delta_{t,h}^{\mathrm{cmp}}(c)-\Omega_{t,h}^{\mathrm{cont}},
\]
and suppose the manifest additionally certifies
\[
\Omega_{t,h}^{\mathrm{cont}}\le L_{\Shadow}S_t^{(L)} + \Delta_t^{\Shadow},
\]
where $L_{\Shadow}\ge 0$ is a public sensitivity constant, $S_t^{(L)}$ is the total accepted rewrite magnitude over the shadow lag window of length $L$, and $\Delta_t^{\Shadow}\ge 0$ is a public upper bound on additional shadow contamination from delayed parser, environment, or evaluator drift. Then the one-sided bridge of Proposition~\ref{prop:shadow_bridge} yields
\[
\mathcal G_{t,H}^{\beta}(u_t)
\ge
\sum_{h=1}^{H}\beta^{h-1}\bigl(\mathcal M_t^{\mathrm{op}}(h)-\bar\delta_{t,h}^{\mathrm{cmp}}-L_{\Shadow}S_t^{(L)}-\Delta_t^{\Shadow}\bigr).
\]
Consequently, if an uncontaminated operational sign-separation margin $\eta_h>0$ is available at horizon $h$, then reliable sign identification at that horizon requires
\[
L_{\Shadow}S_t^{(L)} + \Delta_t^{\Shadow} < \eta_h.
\]
Equivalently, once the contamination budget is fixed, the admissible rewrite speed is bounded by
\[
S_t^{(L)} < \frac{\eta_h-\Delta_t^{\Shadow}}{L_{\Shadow}}.
\]
\end{proposition}

\begin{proof}
The witnesswise hypothesis already contains the operational target and the comparator-distortion term in the current notation. Substituting the displayed contamination upper bound into that witnesswise inequality, summing over horizons, and then minimizing over comparator witnesses gives the stated lower bound on $\mathcal G_{t,H}^{\beta}(u_t)$. The sign condition is the same inequality written so that the right-hand side remains positive at horizon $h$.
\end{proof}


\begin{theorem}[Two-clock shadow viability band under affine penalty envelopes]\label{thm:two_clock}
Suppose the shadow lane is refreshed at frequency $f_{\Shadow}>0$. Let the additional staleness penalty at horizon $h$ obey
\[
\Omega_h^{\mathrm{stale}}(f_{\Shadow}) \le a_h + \frac{b_h}{f_{\Shadow}},
\qquad a_h,b_h\ge 0,
\]
and let the contamination penalty obey
\[
\Omega_h^{\mathrm{cont}}(f_{\Shadow}) \le c_h + d_h f_{\Shadow},
\qquad c_h,d_h\ge 0.
\]
Assume the uncontaminated operational sign-separation margin at horizon $h$ exceeds zero by an amount $\eta_h>0$, meaning that before these penalties are applied one has
\[
\underline{\mathcal M}_t^{\mathrm{op}}(h)-\bar\delta_{t,h}^{\mathrm{cmp}}\ge \eta_h.
\]
Then delayed-shadow sign certification at horizon $h$ is guaranteed whenever there exists $f_{\Shadow}>0$ such that
\[
\Omega_h^{\mathrm{stale}}(f_{\Shadow}) + \Omega_h^{\mathrm{cont}}(f_{\Shadow}) < \eta_h.
\]
Under the displayed affine upper envelopes, a sufficient condition for the existence of such a refresh frequency is
\[
M_h^2 > 4b_h d_h,
\qquad
M_h\Defeq \eta_h-a_h-c_h.
\]
Whenever this holds with $d_h>0$, every frequency in the open interval
\[
\frac{M_h-\sqrt{M_h^2-4b_h d_h}}{2d_h}
<
 f_{\Shadow}
<
\frac{M_h+\sqrt{M_h^2-4b_h d_h}}{2d_h}
\]
is sufficient for certification. If the affine envelopes are tight, the same condition and interval are also necessary.
\end{theorem}

\begin{proof}
Under the assumed envelopes, certification is guaranteed if
\[
\frac{b_h}{f_{\Shadow}} + d_h f_{\Shadow} < M_h.
\]
Multiplying by $f_{\Shadow}>0$ yields the quadratic inequality
\[
d_h f_{\Shadow}^2 - M_h f_{\Shadow} + b_h < 0.
\]
This inequality has a positive solution whenever the discriminant is positive, namely $M_h^2-4b_h d_h>0$. When this holds and $d_h>0$, the sufficient solution set is exactly the stated open interval between the two positive roots. If $d_h=0$, the condition reduces to $b_h/f_{\Shadow}<M_h$, which still gives a one-sided stale-versus-frozen feasibility threshold. Necessity requires tightness of the affine envelopes; absent tightness, the theorem intentionally states only a sufficient viability band.
\end{proof}

\begin{remark}[Frozen shadows are not free]
Taking $f_{\Shadow}\to 0$ drives the staleness penalty $b_h/f_{\Shadow}$ to infinity whenever $b_h>0$. Thus the theory rules out the naive response ``never refresh the shadow at all.'' A completely frozen shadow avoids direct contamination, but it eventually loses relevance and therefore also loses certifying power.
\end{remark}
\begin{definition}[Replayable common-cause vulnerability cover]\label{def:vulnerability_cover}
Fix an active audit family $\mathcal F_t^{\mathrm{aud}}$. A \emph{replayable common-cause vulnerability cover} for that family is a replayable collection of writable or influence channels $\mathcal C_t=\{C_1,\dots,C_m\}$ together with a surjective assignment from family members to channels such that every member can be jointly perturbed whenever the corresponding channel is captured, and the minimum cover size
\[
\chi_t^{\mathrm{cover}}\Defeq \min\{m: \mathcal C_t \text{ covers }\mathcal F_t^{\mathrm{aud}}\}
\]
is itself replayably upper-bounded from manifests, declared writable scope, overlap certificates, or shared dependency roots. This cover formalizes synchronous common-cause capture rather than independent failures. The minimum-cover summary is intended as a structural stress-test object, not as a claim that cover size alone determines exploitability without the additional manifest-preserving forcing assumptions used below.
\end{definition}

\begin{theorem}[Adversarial cross-examination and epistemic singularity]\label{thm:epistemic_singularity}
Let $\mathfrak S_t$ and $\mathfrak R_t$ be the active shadow and refreezing audit families of Definition~\ref{def:audit_family}. Let $\phi_t^{\Shadow}$ and $\phi_t^{\Replay}$ be the replayable monotone calibrations that map the generic attack-reach envelope $A_t^{\mathrm{adv}}$ into the effective common-cause cover units used by the corresponding diversity reserves. For a challenged cohort issued at time $\tau$, define the endogenous integrity budget
\[
\widehat\Gamma_{\tau,h}^{\mathrm{int}}
\Defeq
\Delta_{\tau,h}^{\mathrm{xs}}
+ \kappa_{\mathrm{ov}}\,[1-\Xi_{\tau}^{\Shadow}]_+
+ \kappa_{\mathrm{adv}}\,[\phi_{\tau}^{\Shadow}(A_{\tau}^{\mathrm{adv}})-\Xi_{\tau}^{\Shadow}]_+,
\]
with policy constants $\kappa_{\mathrm{ov}},\kappa_{\mathrm{adv}}\ge0$. Then any lower-margin sign certificate at horizon $h$ that would otherwise use an oracle collusion budget remains valid after subtracting $\widehat\Gamma_{\tau,h}^{\mathrm{int}}$. Likewise any refreezing certificate may replace an oracle semantic floor by the endogenous floor $\underline S_t^{\mathrm{sem}}$.

Conversely, suppose
\[
\phi_{\tau}^{\Shadow}(A_{\tau}^{\mathrm{adv}})\ge \Xi_{\tau}^{\Shadow}
\qquad\text{and}\qquad
\phi_t^{\Replay}(A_t^{\mathrm{adv}})\ge \Xi_t^{\Replay},
\]
and suppose moreover that the active shadow family and replay family each admit replayable common-cause vulnerability covers in the sense of Definition~\ref{def:vulnerability_cover} whose minimum cover sizes are upper-bounded by $\phi_t^{\Shadow}(A_t^{\mathrm{adv}})$ and $\phi_t^{\Replay}(A_t^{\mathrm{adv}})$, respectively, whenever the corresponding diversity reserve thresholds $\Xi_t^{\Shadow}$ and $\Xi_t^{\Replay}$ are met. Assume finally that each such cover is manifest-preserving and familywise output-forcing: capture of the cover channel can synchronize the reported outputs of all mapped family members while leaving the public logs, public manifests, and visible replay transcripts unchanged. Then there exist two recursive-improvement worlds with identical public logs, arbitrarily small cross-shadow disagreement radius $\Delta_{\tau,h}^{\mathrm{xs}}$, identical endogenous semantic floors $\underline S_t^{\mathrm{sem}}$, yet opposite discounted stable-gain sign or opposite true semantic-retention status. The first time at which the displayed calibrated inequalities and the common-cause cover condition hold is an \emph{epistemic singularity}: beyond that point, no observable-only internal audit rule can distinguish genuine agreement from jointly captured agreement.
\end{theorem}

\begin{proof}
The first claim is immediate because any oracle-free certificate must remain valid after accounting for the largest observable disagreement term and for the public penalties attached to overlap and attack reach; the displayed envelope simply upper-bounds the integrity disturbance that was formerly abstracted as an exogenous constant.

For the converse, use the common-cause cover hypothesis to choose a replayable vulnerability cover spanning each active audit family. By the manifest-preserving and familywise output-forcing assumptions, once attack reach exceeds the diversity reserve the accepted proposal can synchronize the reported challenged scores of the entire shadow family and the reported semantic-retention witnesses of the entire refreezing family without altering the public logs or manifests. Keep the visible histories identical across two worlds, but let the hidden slow consequences differ in sign, exactly as in Theorem~\ref{thm:ruler_nogo}. Public disagreement can then be made arbitrarily small even though the true discounted stable gain or true semantic retention differs across worlds. Hence beyond the singularity threshold the ambiguity is observationally irreducible.
\end{proof}

\begin{remark}[Editable rulers enlarge the ambiguity class]\label{rem:ruler_nogo_scope}
The construction above uses only a hidden slow state and observational equivalence on the public lane. Editable evaluators, parser registries, and routing policies are not needed for the impossibility theorem itself. They matter because, in actual recursive-improvement systems, they enlarge the set of observably equivalent worlds by changing which probes, challenge cohorts, and replay environments remain available. The no-go result should therefore be read as a hidden-state impossibility first and an evaluator-edit warning second.
\end{remark}

\begin{remark}[No-meta interpretation of public bounds]
The theorem does not reintroduce an external judge. It does the opposite: it replaces fixed integrity and semantic-retention constants by replayable envelopes computed from family disagreement, provenance overlap, and declared writable scope. The price of staying no-meta is that these envelopes can collapse once attack reach dominates audit diversity; the theory makes that limit explicit rather than hiding it inside a comforting constant.
\end{remark}


\begin{definition}[Steganographic scrubber and information-bottlenecked shadow input]\label{def:scrubber}
Let $p_t$ denote the public manifest of proposal $u_t$, including declared claim, replay artifact, typed-diff metadata, and any certificates accepted by policy. A family of shadow inputs is \emph{scrubbed} if each shadow lane $m$ receives
\[
\tilde u_t^{(m)} = S^{(m)}_{\mathrm{scrub}}(u_t,p_t;\omega_t^{(m)}),
\]
where $\omega_t^{(m)}$ is lane-local randomness, and there exist public constants $\varepsilon_{\mathrm{scrub}},\varepsilon_{\times}\ge 0$ such that
\[
I\!\left(U_t;\tilde U_t^{(m)}\mid P_t\right)\le \varepsilon_{\mathrm{scrub}},
\qquad
I\!\left(\tilde U_t^{(m)};\tilde U_t^{(m')}\mid P_t\right)\le \varepsilon_{\times}
\quad(m\neq m').
\]
Thus each shadow lane sees at most a bounded amount of proposal information beyond the public manifest, and the hidden excess information shared across distinct lanes is likewise bounded.
\end{definition}

\begin{assumption}[Normalized coordinatewise audit-margin certificate class]\label{ass:audit_margin_lipschitz}
For each challenged cohort and audit horizon, the certified shadow margin can be written in the normalized form
\[
\Lambda
=
\Lambda_0(P_t)
+
\sum_{m=1}^{M_t} w_m \, \E[T_t^{(m)}\mid P_t]
-
\sum_{1\le m<m'\le M_t} v_{mm'} \, \E[B_t^{(m,m')}\mid P_t],
\]
where $T_t^{(m)},B_t^{(m,m')}\in[-1,1]$ are replayable lanewise and pairwise bounded audit statistics, $w_m,v_{mm'}\ge 0$, and the total weights satisfy
\[
\sum_{m=1}^{M_t} w_m + \sum_{1\le m<m'\le M_t} v_{mm'} \le 1.
\]
Hence the certified margin is coordinatewise Lipschitz in total variation: if two conditional laws agree on $P_t$ and induce lanewise total-variation deviations at most $\delta_m$ and pairwise deviations at most $\delta_{mm'}$, then
\[
\abs{\Lambda(\mathsf P)-\Lambda(\mathsf Q)}
\le 2\sum_{m=1}^{M_t} w_m \, \delta_m
+2\sum_{1\le m<m'\le M_t} v_{mm'} \, \delta_{mm'}.
\]
This covers normalized weighted averages of bounded lane scores together with normalized cross-lane consistency penalties.
\end{assumption}

\begin{lemma}[Seed-to-output data processing for scrubbed lanes]\label{lem:seed_output_tv}
Fix the public manifest $P_t$. If a scrubbed lane output is generated by a measurable map of the proposal, the manifest, and lane-local randomness, then total-variation deviations on the hidden proposal or seed inputs upper-bound the corresponding total-variation deviations on the scrubbed lane outputs and on their pairwise joint laws. This is an immediate data-processing consequence of the fact that the scrubber map is applied after those hidden inputs are drawn \cite[Chapters~2 and~17]{CoverThomas2006}.
\end{lemma}

\begin{proof}
Condition on $P_t$ and couple the hidden inputs. Applying the same measurable scrubber map to both coupled inputs cannot increase total variation.
\end{proof}

\begin{theorem}[Steganographic scrubbing and information-theoretic isolation]\label{thm:steg_scrub}
Suppose the cross-examination setting of Theorem~\ref{thm:epistemic_singularity} is augmented with a scrubbed shadow family in the sense of Definition~\ref{def:scrubber}, and assume the normalized coordinatewise certificate class of Assumption~\ref{ass:audit_margin_lipschitz}. Then the conditional expected covert-channel-induced inflation of any normalized certified shadow margin is bounded by the explicit modulus
\[
\Psi(\varepsilon_{\mathrm{scrub}},\varepsilon_{\times})
\Defeq
2\sqrt{2\varepsilon_{\mathrm{scrub}}}+2\sqrt{2\varepsilon_{\times}}.
\]
Consequently every expected-margin oracle-free sign certificate remains valid after replacing the endogenous integrity envelope by
\[
\widehat\Gamma_{\tau,h}^{\mathrm{int}} + \Psi(\varepsilon_{\mathrm{scrub}},\varepsilon_{\times}).
\]

Conversely, if the shadow family consumes raw proposal data without such a scrubber and there exists a manifest-invisible trigger channel to which each shadow lane is replayably sensitive, then there exist proposal streams with $A_t^{\mathrm{adv}}=0$ whose hidden triggers synchronously perturb every shadow lane while leaving the public manifest unchanged. Writable-scope separation alone therefore does not rule out covert collusion.
\end{theorem}

\begin{proof}
Condition on the public manifest $P_t$. By Pinsker's inequality and the conditional mutual-information bound, each lanewise deviation between the scrubbed conditional law that still sees $U_t$ and the manifest-only conditional law is bounded by
\[
\delta_m
\Defeq d_{\mathrm{TV}}\!\bigl(\mathcal L(\tilde U_t^{(m)}\mid U_t,P_t),\mathcal L(\tilde U_t^{(m)}\mid P_t)\bigr)
\le \sqrt{2\varepsilon_{\mathrm{scrub}}}
\]
in expectation, while each pairwise deviation between the true pair law and the product of its manifest-only marginals is bounded by
\[
\delta_{mm'}
\Defeq d_{\mathrm{TV}}\!\bigl(\mathcal L(\tilde U_t^{(m)},\tilde U_t^{(m')}\mid P_t),\mathcal L(\tilde U_t^{(m)}\mid P_t)\otimes \mathcal L(\tilde U_t^{(m')}\mid P_t)\bigr)
\le \sqrt{2\varepsilon_{\times}}.
\]
Assumption~\ref{ass:audit_margin_lipschitz} then yields
\[
\abs{\Delta \Lambda}
\le 2\sum_m w_m \, \delta_m + 2\sum_{m<m'} v_{mm'} \, \delta_{mm'}
\le 2\sqrt{2\varepsilon_{\mathrm{scrub}}}+2\sqrt{2\varepsilon_{\times}},
\]
because the total coordinate weights are normalized to at most one. This is the stated modulus.

For the converse, raw proposal access plus a manifest-invisible trigger channel gives a common covert input not represented in $P_t$. If every shadow lane is sensitive to that channel, a proposal writer can choose a trigger-bearing stream that leaves the public manifest unchanged while shifting every lane's hidden input in the same adversarial direction. The synchronized distortion therefore survives writable-scope separation whenever the raw channel remains unsanitized.
\end{proof}

\begin{remark}[Clarification on scrubbers]
The scrubber is not a magical semantic judge that knows the true meaning of the proposal. It is a public, lossy interface contract on what information may flow into audit lanes beyond the replayable manifest. The point is to cap covert coordination bandwidth, not to solve semantics by fiat.
\end{remark}
\begin{remark}[Clarification on scrubber randomness]
Theorem~\ref{thm:steg_scrub} conditions on whatever unpredictability remains in the lane-local seeds $\omega_t^{(m)}$ from the proposal writer's viewpoint. If those seeds become predictable through shared operating-system state or a writable pseudo-random generator, the scrubber can degenerate into a deterministic map. The next definition and theorem make that dependence explicit rather than leaving it implicit.
\end{remark}

\begin{definition}[Entropy-isolated scrubber randomness]\label{def:entropy_isolation}
A scrubbed shadow family satisfies the \emph{entropy-isolation invariant} if each lane-local seed $\omega_t^{(m)}$ is generated by an entropy source outside writable proposal scope and policy publishes a replayable unpredictability certificate relative to the attacker-visible pre-decision state $\mathcal F_t$ and writable-scope measurements. The certificate may be either direct or extractor-based, for example by a leftover-hash-style extractor whose residual is published explicitly rather than inferred from min-entropy alone \cite{BarakEtAl2011LHL,NIST90B2018}, but it must certify a conditional seed residual
\[
d_{\mathrm{TV}}\!\left(\mathcal L(\omega_t^{(m)}\mid \mathcal F_t,A_t^{\mathrm{adv}}),\mathrm{Unif}(\Omega_m)\right) \le \varepsilon_{\omega}.
\]
If an extractor route is used, the theorem consumes only the resulting residual $\varepsilon_{\omega}$ rather than an implicit min-entropy-to-TV leap. In addition, distinct lane seeds satisfy the public cross-seed dependence bound
\[
I\!\left(\omega_t^{(m)};\omega_t^{(m')} \mid \mathcal F_t\right) \le \varepsilon_{\omega,\times}
\qquad(m\neq m').
\]
Thus no proposal writer controlling writable scope can fully predict or deterministically synchronize the scrubber randomness across lanes.
\end{definition}

\begin{theorem}[Entropy isolation law and true cryptographic scrubbing]\label{thm:entropy_isolation}
Suppose Definition~\ref{def:scrubber} holds, the scrubber randomness further satisfies Definition~\ref{def:entropy_isolation}, the seed-to-output data-processing lemma (Lemma~\ref{lem:seed_output_tv}) applies to the scrubber family, and the normalized coordinatewise certificate class of Assumption~\ref{ass:audit_margin_lipschitz} also holds. Then every shadow-margin certificate valid under Theorem~\ref{thm:steg_scrub} remains valid after adding the seed-isolation residual
\[
\Psi_{\mathrm{iso}}(\varepsilon_{\omega},\varepsilon_{\omega,\times})
\Defeq
2\varepsilon_{\omega}+2\sqrt{2\varepsilon_{\omega,\times}},
\]
so that the total certified residual becomes
\[
\Psi_{\mathrm{tot}}
\Defeq
\Psi(\varepsilon_{\mathrm{scrub}},\varepsilon_{\times})
+
\Psi_{\mathrm{iso}}(\varepsilon_{\omega},\varepsilon_{\omega,\times}).
\]
In particular, entropy isolation improves scrubber security through a directly certified total-variation residual rather than through an implicit min-entropy-to-TV leap. In the idealized zero-leakage scrubber limit $\varepsilon_{\mathrm{scrub}}=\varepsilon_{\times}=0$, the only remaining certificate penalty is the seed-isolation residual.

Conversely, if the attacker-conditioned seed residual is effectively maximal so that the seeds are known to the proposal writer, and the scrubber class is rich enough that deterministic preimages can preserve a trigger through the scrubber map, then the scrubber becomes deterministic from the attacker's viewpoint and there exist covert-trigger proposal streams that preserve the public manifest while defeating the intended scrubbing barrier. Thus lossy interfaces without entropy isolation do not provide information-theoretic isolation.
\end{theorem}

\begin{proof}
The forward direction augments Theorem~\ref{thm:steg_scrub} with attacker uncertainty about the lane-local seeds. The direct residual $\varepsilon_{\omega}$ already controls, in total variation, how far each lane-local seed law can be from an ideal manifest-independent seed. By Lemma~\ref{lem:seed_output_tv}, the same residual upper-bounds the induced perturbation on each scrubbed lane output. The cross-lane dependence budget contributes at most $\sqrt{2\varepsilon_{\omega,\times}}$ in total variation by Pinsker's inequality, and the same data-processing lemma carries that bound to the pairwise scrubbed-output law. Applying Assumption~\ref{ass:audit_margin_lipschitz} therefore adds at most
\[
2\varepsilon_{\omega}+2\sqrt{2\varepsilon_{\omega,\times}}
\]
to the certified margin residual, which must be added to rather than substituted for the scrubber residual already bounded in Theorem~\ref{thm:steg_scrub}.

For the converse, if the seeds are effectively known and the scrubber map admits deterministic trigger-preserving preimages, then the proposal writer can optimize directly against that deterministic map and embed a trigger that survives scrubbing while remaining absent from the public manifest. This reconstructs the covert-channel failure mode despite nominal scrubbing.
\end{proof}


\begin{assumption}[Monotone retention envelopes]\label{ass:retention_bounds}
There exist nondecreasing functions
\[
\phi_D^-,\phi_D^+,\phi_Q^-,\phi_Q^+:
\R_+\to\R_+
\]
with
\[
\phi_D^+(x)\ge \phi_D^-(x)\ge 0,
\qquad
\phi_Q^+(x)\ge \phi_Q^-(x)\ge 0,
\qquad
\phi_D^{\pm}(0)=\phi_Q^{\pm}(0)=0,
\]
such that for every proposal issued at time $t$ and every horizon $h\ge1$,
\[
\mathcal M_t(0)-\phi_D^+\!\bigl(\mathcal D_t(h)\bigr)-\phi_Q^+\!\bigl(Q_t(h)\bigr)
\le
\mathcal M_t(h)
\le
\mathcal M_t(0)-\phi_D^-\!\bigl(\mathcal D_t(h)\bigr)-\phi_Q^-\!\bigl(Q_t(h)\bigr),
\]
where
\[
Q_t(h)\Defeq \sum_{s=t}^{t+h-1}q_s.
\]
\end{assumption}

\begin{remark}[Identification status of the retention envelopes]\label{rem:retention_identification}
Assumption~\ref{ass:retention_bounds} is intentionally an envelope assumption rather than a first-principles identification theorem. The functions $\phi_D^{\pm}$ and $\phi_Q^{\pm}$ must be estimated or stress-tested from logged drift, backlog, and retained-margin data, possibly with regime-dependent envelopes if the operating region is not stationary. The resulting half-life law is therefore an \emph{auditable crossing algebra}: it is sharp conditional on the published envelopes, but the empirical sharpness of the envelopes remains a separate estimation problem.
\end{remark}

\begin{theorem}[Generalized realized improvement half-life crossing law]\label{thm:half_life}
Assume \ref{ass:retention_bounds}. For a proposal issued at time $t$ with $\mathcal M_t(0)>0$, define the lower and upper crossing indices
\[
\tau_t^{-}\Defeq \inf\left\{h\ge1: \phi_D^+\!\bigl(\mathcal D_t(h)\bigr)+\phi_Q^+\!\bigl(Q_t(h)\bigr)\ge \tfrac12\mathcal M_t(0)\right\},
\]
\[
\tau_t^{+}\Defeq \inf\left\{h\ge1: \phi_D^-\!\bigl(\mathcal D_t(h)\bigr)+\phi_Q^-\!\bigl(Q_t(h)\bigr)\ge \tfrac12\mathcal M_t(0)\right\},
\]
with the convention $\inf\varnothing=\infty$. Then
\[
\tau_t^{-}\le \tau_{1/2}(t)\le \tau_t^{+}.
\]
In particular, if the lower crossing set is nonempty then the half-life is finite, and if the upper crossing set is empty then the half-life cannot yet have occurred on the observed horizon.
\end{theorem}

\begin{proof}
From Assumption~\ref{ass:retention_bounds},
\[
\mathcal M_t(h)\ge \mathcal M_t(0)-\phi_D^+\!\bigl(\mathcal D_t(h)\bigr)-\phi_Q^+\!\bigl(Q_t(h)\bigr).
\]
Hence whenever $\phi_D^+\!\bigl(\mathcal D_t(h)\bigr)+\phi_Q^+\!\bigl(Q_t(h)\bigr)<\tfrac12\mathcal M_t(0)$, we still have $\mathcal M_t(h)>\tfrac12\mathcal M_t(0)$, so the half-life event has not yet happened. This proves $\tau_t^-\le \tau_{1/2}(t)$. Similarly,
\[
\mathcal M_t(h)\le \mathcal M_t(0)-\phi_D^-\!\bigl(\mathcal D_t(h)\bigr)-\phi_Q^-\!\bigl(Q_t(h)\bigr),
\]
so once $\phi_D^-\!\bigl(\mathcal D_t(h)\bigr)+\phi_Q^-\!\bigl(Q_t(h)\bigr)\ge \tfrac12\mathcal M_t(0)$, the half-life inequality must hold. Therefore $\tau_{1/2}(t)\le \tau_t^+$.
\end{proof}

\begin{corollary}[Linear specialization of the half-life crossing law]\label{cor:half_life_coarse}
Assume \ref{ass:retention_bounds} with
\[
\phi_D^\pm(x)=L_D^\pm x,
\qquad
\phi_Q^\pm(x)=L_q^\pm x,
\]
where $L_D^+\ge L_D^-\ge 0$ and $L_q^+\ge L_q^-\ge 0$. If for all $s\ge t$,
\[
\underline D \le D_s^{\mathrm{yard}}\le \bar D,
\qquad
0\le q_s\le \bar q,
\]
and if the displayed denominators are positive, then
\[
\left\lceil \frac{\mathcal M_t(0)}{2(L_D^+\bar D + L_q^+\bar q)} \right\rceil
\le
\tau_{1/2}(t)
\le
\left\lceil \frac{\mathcal M_t(0)}{2L_D^-\underline D} \right\rceil.
\]
The upper bound uses only the drift floor and remains meaningful even if healthy operation drives the queue to zero on many steps.
\end{corollary}

\begin{proof}
Use $\mathcal D_t(h)\le h\bar D$ and $Q_t(h)\le h\bar q$ for the lower bound on $\tau_{1/2}(t)$, and $\mathcal D_t(h)\ge h\underline D$ together with $Q_t(h)\ge 0$ for the upper bound.
\end{proof}

\begin{corollary}[Certified stable improvement from replayable retained-margin certificates]\label{cor:shadow_cert}
Fix a proposal satisfying $\mathsf{Comp}_t(u_t)=1$, and suppose there exists a replayable retained-margin lower certificate $\{\underline{\mathcal M}_t^{\mathrm{op}}(h)\}_{h\ge 1}$ in the sense of Definition~\ref{def:retained_margin_cert}; by Proposition~\ref{prop:challenge_to_shadow}, one operational way to obtain such a certificate is to combine a class-conditional delayed-challenge certificate with a published transport-calibration family. Under Proposition~\ref{prop:shadow_bridge}, a sufficient condition for a nonnegative replayable lower bound on discounted stable gain is
\[
\sum_{h=1}^{H}\beta^{h-1}\bigl(\underline{\mathcal M}_t^{\mathrm{op}}(h)-\bar\delta_{t,h}^{\mathrm{cmp}}\bigr)\ge 0.
\]
If, in addition, the proposal-local escrow debt stays below the policy threshold so that $\eta_{t,H}^{\max}(u_t)\le \bar\delta_r(C_t)$, then the proposal satisfies the Interface-A and Interface-B clauses of $(H,\beta)$-certified stable improvement. If a replayable class-homogeneity envelope is also published and
\[
\sum_{h=1}^{H}\beta^{h-1}\bigl(\underline{\mathcal M}_t^{\mathrm{op}}(h)-\delta_h^{\mathrm{class}}(\mathcal C_t,\mathfrak B_t)-\bar\delta_{t,h}^{\mathrm{cmp}}\bigr)\ge 0,
\]
then the same condition upgrades the conclusion to oracle stable improvement for the cohort-specific retained-margin target as well.
\end{corollary}

\begin{proof}
On the event $\mathsf{Comp}_t(u_t)=1$, Proposition~\ref{prop:shadow_bridge} gives
\[
\mathcal G_{t,H}^{\beta}(u_t)
\ge
\sum_{h=1}^{H}\beta^{h-1}\bigl(\mathcal M_t^{\mathrm{op}}(h)-\bar\delta_{t,h}^{\mathrm{cmp}}\bigr)
\ge
\sum_{h=1}^{H}\beta^{h-1}\bigl(\underline{\mathcal M}_t^{\mathrm{op}}(h)-\bar\delta_{t,h}^{\mathrm{cmp}}\bigr),
\]
where the second inequality is exactly the lower-certificate property from Definition~\ref{def:retained_margin_cert}. Thus a nonnegative replayable operational lower certificate implies a nonnegative replayable lower bound on the lower oracle endpoint. The debt condition is the shared Interface-B clause in the definitions of certified and oracle stable improvement. If the class-homogeneity envelope is available, Proposition~\ref{prop:challenge_to_shadow} supplies the additional oracle-margin lower certificate, and the final displayed inequality gives the oracle conclusion.
\end{proof}

\begin{proposition}[Proposal-local discounted debt budget]\label{prop:local_debt_budget}
Under Definition~\ref{def:cohort_escrow}, every cohort $\tau$ satisfies, for all $t\ge \tau$,
\[
\eta_{\tau,t}
\le
\alpha_r^{t-\tau}\bigl(\mathcal B_\tau(u_\tau)+U_{\tau}^{\mathrm{rep}}+U_{\tau}^{\mathrm{haz}}\bigr)
+
\sum_{s=\tau+1}^{t}\alpha_r^{t-s} a_{\tau,s}^{\mathrm{sec}}
+
\lambda_q\sum_{s=\tau}^{t-1}\alpha_r^{t-1-s} b_{\tau,s}.
\]
If, in addition, $a_{\tau,s}^{\mathrm{sec}}\le \bar a_{\tau}$ and $b_{\tau,s}\le \bar b_{\tau}$ for all $s\ge \tau$, and if $\alpha_r<1$, then
\[
\sup_{k\ge 0}\eta_{\tau,\tau+k}
\le
\max\left\{\mathcal B_\tau(u_\tau)+U_{\tau}^{\mathrm{rep}}+U_{\tau}^{\mathrm{haz}},\, \frac{\bar a_{\tau}+\lambda_q \bar b_{\tau}}{1-\alpha_r}\right\}.
\]
The same bound holds for every finite-horizon peak debt.
\end{proposition}

\begin{proof}
From the cohort recursion,
\[
\eta_{\tau,t+1}
\le
\alpha_r\eta_{\tau,t} + a_{\tau,t+1}^{\mathrm{sec}} + \lambda_q b_{\tau,t},
\]
because the matched-retirement increment and the explicit retirement fraction only reduce the stock. Iterating from
\[
\eta_{\tau,\tau}=\mathcal B_\tau(u_\tau)+U_{\tau}^{\mathrm{rep}}+U_{\tau}^{\mathrm{haz}}
\]
yields the displayed discounted budget. Under the uniform caps and $\alpha_r<1$,
\[
\eta_{\tau,\tau+k}
\le
\alpha_r^k\bigl(\mathcal B_\tau(u_\tau)+U_{\tau}^{\mathrm{rep}}+U_{\tau}^{\mathrm{haz}}\bigr)
+
\frac{(\bar a_{\tau}+\lambda_q\bar b_{\tau})(1-\alpha_r^k)}{1-\alpha_r},
\]
which is always bounded by the displayed maximum.
\end{proof}

\begin{corollary}[Observable debt certificate]\label{cor:observable_debt_cert}
Fix a proposal issued at time $t$. Suppose logs provide conservative upper bounds
\[
\bar B_t\ge \mathcal B_t(u_t),\qquad
\bar U_t^{\mathrm{rep}}\ge U_t^{\mathrm{rep}},\qquad
\bar U_t^{\mathrm{haz}}\ge U_t^{\mathrm{haz}},\qquad
\bar a_t\ge a_{t,s}^{\mathrm{sec}},\qquad
\bar b_t\ge b_{t,s}
\]
for all $s\ge t$. If $\alpha_r<1$ and
\[
\max\left\{\bar B_t+\bar U_t^{\mathrm{rep}}+\bar U_t^{\mathrm{haz}},\, \frac{\bar a_t+\lambda_q \bar b_t}{1-\alpha_r}\right\}
\le \bar\delta_r(C_t),
\]
then the debt part of the certified-stable-improvement criterion is certified for every horizon. Combined with Corollary~\ref{cor:shadow_cert}, a nonnegative calibrated lower stable-gain bound yields a fully operational sufficient \emph{certified stable-improvement} rule; adding base-admissibility and governance-safety certificates upgrades it to a full admissibility rule.
\end{corollary}

\begin{proof}
Apply Proposition~\ref{prop:local_debt_budget} with the conservative bounds.
\end{proof}

\begin{definition}[Bundled stable gain, bundled debt, and interaction debt]\label{def:interaction_debt}
Fix a bundle $U=\{u_t^{(1)},\dots,u_t^{(m)}\}$ of simultaneously admitted edits over horizon $(H,\beta)$. Let $\mathcal G_{t,H}^{\beta}(U)$ denote the discounted stable gain of the joint intervention that applies the whole bundle inside the common continuation skeleton. Let $\eta_{t,H}^{\max}(U)$ denote the corresponding bundle-level peak debt over the same horizon, and let $\mathsf{BaseAdm}_t(U)$ and $\mathsf{GovSafe}_{t,H}(U)$ denote the bundle-level base-admissibility and governance-safety predicates. Let $\underline{\mathcal G}_{t,H}^{\beta}(u_t^{(i)})$ denote any replayable lower certificate for the discounted stable gain of edit $i$ when evaluated in isolation against the common baseline.

To make interaction debt constructive rather than existential, policy publishes a replayable interference graph on the issued bundle. For each pair $(i,j)$ with $i<j$, let
\[
\overline\Delta_{ij}
\Defeq
\lambda_{\mathrm{dep}}\,\Omega_{ij}^{\mathrm{dep}}
+\lambda_{\mathrm{wr}}\,\Omega_{ij}^{\mathrm{wr}}
+\lambda_{\mathrm{q}}\,\Omega_{ij}^{\mathrm{q}}
+\lambda_{\mathrm{h}}\,\Omega_{ij}^{\mathrm{h}},
\]
where $\Omega_{ij}^{\mathrm{dep}}$ is a replayable dependency-overlap envelope, $\Omega_{ij}^{\mathrm{wr}}$ is shared writable-scope overlap, $\Omega_{ij}^{\mathrm{q}}$ is protected-lane queue contention, and $\Omega_{ij}^{\mathrm{h}}$ is a replayable cross-sensitivity or cross-Hessian upper envelope on task or governance interference. Let $\overline\Delta_{\mathrm{ho}}(U)\ge 0$ be a replayable residual upper envelope for higher-order interaction not captured by the pairwise terms, for example from logged hyperedge overlap or a submodularity-defect bound. A \emph{constructive interaction-debt certificate} is the replayable nonpositive quantity
\[
\widehat\Lambda_{t,H}^{\mathrm{int}}(U)
\Defeq
-\sum_{1\le i<j\le m}\overline\Delta_{ij}
-\overline\Delta_{\mathrm{ho}}(U)
\le 0.
\]
The interaction debt of the bundle is any replayable nonpositive quantity $\Lambda_{t,H}^{\mathrm{int}}(U)$ satisfying
\[
\Lambda_{t,H}^{\mathrm{int}}(U)\le \widehat\Lambda_{t,H}^{\mathrm{int}}(U)\le 0
\]
and the stable-gain lower bound
\[
\mathcal G_{t,H}^{\beta}(U)\ge \sum_{i=1}^m \underline{\mathcal G}_{t,H}^{\beta}(u_t^{(i)}) + \Lambda_{t,H}^{\mathrm{int}}(U).
\]
Thus interaction debt is tied to explicit shared writable scope, dependency overlap, queue interference, and cross-sensitivity envelopes rather than being left as a bare existential penalty.
\end{definition}

\begin{proposition}[Compositional certification under interaction debt]\label{prop:interaction_debt}
If a bundle $U$ satisfies
\[
\sum_{i=1}^m \underline{\mathcal G}_{t,H}^{\beta}(u_t^{(i)}) + \Lambda_{t,H}^{\mathrm{int}}(U)\ge 0,
\qquad
\eta_{t,H}^{\max}(U)\le \bar\delta_r(C_t),
\qquad
\mathsf{BaseAdm}_t(U)=1,
\qquad
\mathsf{GovSafe}_{t,H}(U)=1,
\]
then the bundle is an $(H,\beta)$-full admissible improvement. In particular, individually certified edits need not compose unless interaction debt is also controlled.
\end{proposition}

\begin{proof}
The displayed stable-gain inequality furnishes a nonnegative replayable lower certificate for the discounted stable gain of the whole bundle. The bundle debt inequality gives certified stable improvement, while base admissibility and governance safety supply the remaining admission clauses. The final sentence follows because $\Lambda_{t,H}^{\mathrm{int}}(U)$ can be strictly negative even when each individual lower bound is nonnegative.
\end{proof}

\begin{proposition}[Admission decomposition and sufficient rule]\label{prop:master_admissibility}
For either a single proposal or a bundle of proposals, the conjunction
\[
\mathsf{BaseAdm}_t(\cdot)=1,\qquad
\underline{\mathcal G}_{t,H}^{\beta}(\cdot)\ge 0,\qquad
\eta_{t,H}^{\max}(\cdot)\le \bar\delta_r(C_t),\qquad
\mathsf{GovSafe}_{t,H}(\cdot)=1
\]
is sufficient for $(H,\beta)$-full admissibility. Thus the theory decomposes admission into a base replay-and-typing layer, a certified stable-gain layer, a debt layer, and a governance-safety layer rather than treating ``admissible'' as a single opaque predicate.
\end{proposition}

\begin{proof}
This is immediate from the definition of full admissible improvement for single proposals and from Proposition~\ref{prop:interaction_debt} for bundles.
\end{proof}

\begin{theorem}[Ignoring any downside-reserve component enables Trojan under-reservation]\label{thm:blast_radius}
Suppose admission were based on an incomplete downside reserve, for example by initializing the proposal-local debt stock from claimed upside alone, or by omitting either the repair reserve or the persistent hazard reserve. Then for any debt threshold $\bar\delta_r>0$ and any family of proposals with arbitrarily large writable scope or dependency closure, there exist under-declared proposals with
\[
0\le [\Delta_t^{\mathrm{claim}}]_+<\bar\delta_r
\]
that still request arbitrarily large structural access. In that sense Trojan under-declaration is always possible when initial debt ignores structural blast radius, repair reserve, and persistent hazard reserve.

Under Definition~\ref{def:blast_radius} and Definition~\ref{def:cohort_escrow}, suppose there exists a public lower envelope $\mathcal H_t(u_t)$ on hidden-state or governance-surface reach such that
\[
\mathcal B_t(u_t)\ge \mathcal H_t(u_t)
\qquad\text{for all admissible proposals.}
\]
Then any proposal with $\mathcal H_t(u_t)>\bar\delta_r(C_t)$ necessarily has
\[
\eta_{t,t}\ge \mathcal B_t(u_t)>\bar\delta_r(C_t)
\]
and is rejected immediately, regardless of how small its claimed upside is. Thus a sufficiently conservative systemic risk bond makes Trojan under-declaration impossible at the debt gate.
\end{theorem}

\begin{proof}
Without a bond term, the initial debt depends only on the claimed upside. Hence one can take any structurally dangerous proposal and report an arbitrarily small positive claim, which drives the initialization below any fixed threshold while leaving writable scope unchanged. This proves the first statement.

With the downside-reserve initialization of Definition~\ref{def:cohort_escrow}, the issuance stock is $\eta_{t,t}=\mathcal B_t(u_t)+U_t^{\mathrm{rep}}+U_t^{\mathrm{haz}}$. If the bond lower-bounds hidden or governance reach and that reach exceeds the policy threshold, then the initialized debt already exceeds the threshold before any later retirement, routing, or calibration step can help; a fortiori the same holds once repair and hazard reserves are added. Therefore the proposal is rejected at issuance.
\end{proof}

\begin{theorem}[Bond-only initialization remains under-reserved against topological obfuscation]\label{thm:causal_bond}
Suppose a governance policy uses only static graph-size terms such as $|\mathcal W_t(u_t)|$ and $|\operatorname{Dep}_t(\mathcal W_t(u_t))|$ when initializing proposal-local debt. Then for any fixed debt threshold $\bar\delta_r>0$ there exist representation-preserving refactorings that compress a high-impact rewrite into a single apparent module or dispatcher while leaving its output-side intervention reach essentially unchanged. In that sense, purely topological bonds are spoofable under topological obfuscation and remain insufficient even if one later appends repair or hazard reserves that do not track intervention reach.

Suppose instead that the policy uses Definition~\ref{def:blast_radius} with a replayable causal lower envelope $\Gamma_t^{\mathrm{cic}}(u_t)$ satisfying
\[
\Gamma_t^{\mathrm{cic}}(u_t)\ge \mathcal H_t(u_t)
\]
for some public lower bound $\mathcal H_t(u_t)$ on hidden-state or governance-surface reach, and that $\lambda_C>0$. Then any proposal with
\[
\mathcal H_t(u_t)>\bar\delta_r(C_t)/\lambda_C
\]
necessarily has
\[
\eta_{t,t}\ge \mathcal B_t(u_t)\ge \lambda_C\Gamma_t^{\mathrm{cic}}(u_t)>\bar\delta_r(C_t)
\]
and is rejected at issuance, regardless of how aggressively its internal topology has been compressed or obfuscated.
\end{theorem}

\begin{proof}
Static graph-size counts depend on representation. A proposer can fuse many downstream effects into one learned block, one dynamic dispatch string, or one monolithic weight tensor, thereby making the writable set and visible dependency closure small while preserving large intervention reach. Hence graph-size-only bonds can be spoofed.

By contrast, the second statement depends only on the public lower envelope $\Gamma_t^{\mathrm{cic}}(u_t)$ and the downside-reserve initialization rule $\eta_{t,t}=\mathcal B_t(u_t)+U_t^{\mathrm{rep}}+U_t^{\mathrm{haz}}$. If $\mathcal B_t(u_t)\ge \lambda_C\Gamma_t^{\mathrm{cic}}(u_t)\ge \lambda_C\mathcal H_t(u_t)$ and the right-hand side exceeds the debt threshold, then the proposal fails immediately even before the extra repair and hazard reserves are counted. The argument is representation-agnostic because it uses intervention reach rather than topology count.
\end{proof}

\begin{theorem}[Strict-conservation repair-throughput law]\label{thm:strict_conservation}
Consider the strict-conservation regime $\alpha_r=1$. Suppose there exist nonnegative adapted processes $L_t$, $a_t^{\mathrm{sec}}$, and $U_t$ such that aggregate debt obeys the upper recursion
\[
r_{t+1}\le r_t + L_t + a_t^{\mathrm{sec}} - U_t,
\]
where $L_t$ is new \emph{primary} liability plus backlog charge, $a_t^{\mathrm{sec}}$ is repair-induced secondary liability, and $U_t$ is gross effective retirement. Suppose further that
\[
\E[L_t\mid \mathcal F_t]\le \bar\lambda_L + \lambda_q \bar q,
\qquad
\E[U_t\mid \mathcal F_t]\ge \mu^{\mathrm{rep}},
\]
and that repair-induced secondary liabilities satisfy
\[
\E[a_t^{\mathrm{sec}}\mid \mathcal F_t] \le \eps_{\mathrm{rep}}\,\E[U_t\mid \mathcal F_t]
\qquad\text{with }\eps_{\mathrm{rep}}\in[0,1).
\]
Then the conditional drift of strict-conservation debt satisfies
\[
\E[r_{t+1}-r_t\mid\mathcal F_t]
\le
\bar\lambda_L + \lambda_q \bar q - (1-\eps_{\mathrm{rep}})\mu^{\mathrm{rep}}.
\]
Consequently, if
\[
\bar\lambda_L + \lambda_q \bar q < (1-\eps_{\mathrm{rep}})\mu^{\mathrm{rep}},
\]
strict-conservation debt has negative conditional drift and cannot grow linearly on average. Conversely, if there exists $\varepsilon_{+}>0$ such that
\[
r_{t+1}\ge r_t + L_t + a_t^{\mathrm{sec}} - U_t,
\qquad
\E[L_t+a_t^{\mathrm{sec}}-U_t\mid\mathcal F_t]\ge \varepsilon_{+}
\]
on every step, then $\E[r_t]$ grows at least linearly in $t$.
\end{theorem}

\begin{proof}
Taking conditional expectation in the displayed upper recursion gives
\[
\E[r_{t+1}-r_t\mid\mathcal F_t]
\le
\E[L_t\mid\mathcal F_t]+\E[a_t^{\mathrm{sec}}\mid\mathcal F_t]-\E[U_t\mid\mathcal F_t].
\]
By hypothesis,
\[
\E[L_t\mid\mathcal F_t]\le \bar\lambda_L+\lambda_q\bar q,
\qquad
\E[a_t^{\mathrm{sec}}\mid\mathcal F_t]\le \eps_{\mathrm{rep}}\E[U_t\mid\mathcal F_t],
\]
so
\[
\E[r_{t+1}-r_t\mid\mathcal F_t]
\le
\bar\lambda_L+\lambda_q\bar q-(1-\eps_{\mathrm{rep}})\E[U_t\mid\mathcal F_t]
\le
\bar\lambda_L+\lambda_q\bar q-(1-\eps_{\mathrm{rep}})\mu^{\mathrm{rep}}.
\]
This proves the upper drift bound. Summing it over time yields the nonpositive-average-growth claim when the right-hand side is strictly negative. The final linear-growth statement follows by summing the displayed lower drift inequality over time.
\end{proof}

\subsection{Verification bottleneck and queue sharpness}
\begin{theorem}[Verification queue phase diagram]\label{thm:queue_phase}
Consider the fluid queue law
\[
q_{t+1}=\bigl[q_t+\lambda-\mu\bigr]_+
\]
with constant proposal rate $\lambda\ge0$ and verification service $\mu\ge0$.
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item If $\lambda<\mu$, then $q_t$ reaches $0$ in finite time and remains there.
  \item If $\lambda=\mu$, then $q_t\equiv q_0$.
  \item If $\lambda>\mu$, then $q_t=q_0+t(\lambda-\mu)$ for all sufficiently large $t$ and diverges linearly.
\end{enumerate}
If latent error debt satisfies
\[
r_{t+1}\ge \alpha_r r_t + \lambda_q q_t,
\qquad \alpha_r\in[0,1],\ \lambda_q>0,
\]
then the overloaded regime $\lambda>\mu$ prevents uniform control of error debt.
\end{theorem}

\begin{proof}
The queue recursion is deterministic. If $\lambda<\mu$, then each step decreases the bracket by $\mu-\lambda$ until it hits zero, after which it stays zero. If $\lambda=\mu$, then the update is constant. If $\lambda>\mu$, then after the queue becomes positive it increments by the constant difference $\lambda-\mu$ at every step. The lower recursion for $r_t$ then forces unbounded growth when $q_t$ grows linearly and $\lambda_q>0$.
\end{proof}

\begin{definition}[Capability-dependent verification complexity]\label{def:red_queen}
Let $\kappa$ be any productive capability functional on the productive state coordinates and define the accepted productive increment by
\[
\Delta \kappa_s\Defeq \bigl[\kappa(x_{s+1})-\kappa(x_s)\bigr]_+.
\]
Then
\[
K_t^{\mathrm{cap}}\Defeq \sum_{s=0}^{t-1}\Delta \kappa_s
\]
be cumulative accepted \emph{productive} capability expansion. A \emph{verification complexity law} is a predictable nondecreasing map $c^{\mathrm{ver}}(K_t^{\mathrm{cap}})\ge 1$ measuring the replay-and-check cost of current proposals at productive capability level $K_t^{\mathrm{cap}}$. Under physical verification budget $P_t>0$, the effective service rate is
\[
\mu_t^{\mathrm{ver}}\Defeq \frac{P_t}{c^{\mathrm{ver}}(K_t^{\mathrm{cap}})}.
\]
This separates algorithmic or semantic complexity from raw physical throughput while refusing to count pure representational churn as growth.
\end{definition}

\begin{definition}[Governance-computation complexity and proof-carrying governance]\label{def:pcg}
Let $c_t^{\mathrm{gov}}(u_t)$ denote the replay cost of recomputing the governance predicates required by the admission rule for proposal $u_t$, including any semantic-volume, routing-sensitivity, replay-floor, or integrity-envelope computations demanded by policy. A \emph{proof-carrying governance regime} is policy-admissible if each proposal carries a replay-checkable certificate $\pi_t$ attesting the claimed governance bounds together with a public checker satisfying
\[
\widetilde c_t^{\mathrm{gov}}(u_t,\pi_t)\le \alpha_{\mathrm{gov}}\,c_t^{\mathrm{gov}}(u_t)+b_{\mathrm{gov}}
\]
for public constants $\alpha_{\mathrm{gov}}\in[0,1]$ and $b_{\mathrm{gov}}\ge 0$. The special case $(\alpha_{\mathrm{gov}},b_{\mathrm{gov}})=(1,0)$ recovers the weaker ``no worse than full recomputation'' requirement. If full recomputation is performed only on a predictable audit fraction $\sigma_t^{\mathrm{gov}}\in[0,1]$, the effective per-proposal governance workload is
\[
\widehat c_t^{\mathrm{gov}}(u_t,\pi_t)\Defeq \widetilde c_t^{\mathrm{gov}}(u_t,\pi_t)+\sigma_t^{\mathrm{gov}}c_t^{\mathrm{gov}}(u_t).
\]
This definition does not assume magical zero-cost proofs; it separates producer-paid certificate generation from the public cost of checking certificates and selectively recomputing them.
\end{definition}


\begin{definition}[Constitutional logic class and decidability-bounded proof escrow]\label{def:constitutional_logic}
A proof-carrying governance regime is \emph{constitutionally bounded} if there exists a public family of certificate languages and checkers
\[
\mathfrak L^{\mathrm{const}}=\{(\mathcal L_\ell,V_\ell,T_\ell,M_\ell):\ell\in\mathcal I\}
\]
with the following properties.
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item Every checker $V_\ell$ is \emph{total}: it halts on every input certificate and proposal transcript.
  \item For every certificate $\pi$, verification time and working memory satisfy the public static envelopes
  \[
  \mathrm{time}(V_\ell(\pi))\le T_\ell(|\pi|),
  \qquad
  \mathrm{mem}(V_\ell(\pi))\le M_\ell(|\pi|).
  \]
  \item Fast-admission proposals must declare a language tag $\ell_t\in\mathcal I$ and provide a certificate $\pi_t\in\mathcal L_{\ell_t}$; certificates outside $\mathfrak L^{\mathrm{const}}$ are not fast-certifiable by type inside the current constitutional lane and must therefore be routed to a non-fast action such as richer proof escalation or sandboxing before semantic deployment is considered.
\end{enumerate}
The resulting public verification cost is
\[
\widetilde c_{t,\mathrm{const}}^{\mathrm{gov}}\Defeq T_{\ell_t}(|\pi_t|),
\]
and the corresponding effective governance workload is
\[
\widehat c_{t,\mathrm{const}}^{\mathrm{gov}}
\Defeq
\widetilde c_{t,\mathrm{const}}^{\mathrm{gov}}+\sigma_t^{\mathrm{gov}}c_t^{\mathrm{gov}}(u_t).
\]
\end{definition}

\begin{definition}[Extensional adequacy on a designated certificate class]\label{def:adequacy}
Let $\mathfrak C^{\mathrm{valid}}$ be a designated replayable recursively enumerable class of governance certificates paired with proposal transcripts. A checker is \emph{extensionally adequate on $\mathfrak C^{\mathrm{valid}}$} if every pair $(\pi,u)\in \mathfrak C^{\mathrm{valid}}$ that encodes a governance proof valid relative to the designated proof system and transcript semantics is accepted whenever the checker halts. This is weaker than completeness on all syntactic certificates but strong enough to forbid the trivial ``reject everything'' escape on the designated class used in the halting reduction. No privileged semantic judge is invoked here: validity is only the proof-theoretic relation fixed by the designated certificate class.
\end{definition}

\begin{theorem}[Decidability-bounded proof escrow and the G\"odelian ceiling]\label{thm:godelian_ceiling}
Suppose governance proofs are accepted from an unrestricted certificate language whose checker can simulate a universal computation model. Suppose moreover that there exists a computable embedding from Turing-machine/input pairs $(M,w)$ into a designated replayable recursively enumerable class $\mathfrak C^{\mathrm{valid}}$ of certificate-proposal pairs such that the embedded pair is valid in that designated proof-theoretic sense if and only if $M$ halts on $w$. If the checker is required to be both sound and extensionally adequate in the sense of Definition~\ref{def:adequacy} on that designated class, then for every public governance budget $B<\infty$, there exists a finite certificate-proposal pair $(\pi,u)$ such that any such checker either fails to halt or consumes more than $B$ time or memory while attempting to validate the claimed governance bounds. Consequently no uniform public upper envelope on $\widetilde c_t^{\mathrm{gov}}$ exists, and proof-level denial-of-service cannot be excluded.

By contrast, under Definition~\ref{def:constitutional_logic}, every admissible certificate is checked by a total verifier with public cost bound $\widetilde c_{t,\mathrm{const}}^{\mathrm{gov}}\le T_{\ell_t}(|\pi_t|)$. Hence every Red-Queen or queue-slack statement in which governance workload appears remains valid after replacing $\widehat c_t^{\mathrm{gov}}(u_t,\pi_t)$ by $\widehat c_{t,\mathrm{const}}^{\mathrm{gov}}$. Any proposal whose safety argument cannot be expressed inside $\mathfrak L^{\mathrm{const}}$ lies beyond the \emph{G\"odelian ceiling} of the current fast-admission regime. Such a proposal cannot be certified \emph{inside that fast regime}; it must therefore be deferred to a richer proof lane, sandboxed under tighter scope, or held non-admissible until a replayable proof interface capable of expressing it is provided. This is a regime-relative routing statement, not a claim that the proposal is anti-growth or permanently undeployable.
\end{theorem}

\begin{proof}
For the first claim, fix a universal Turing machine $M$ and input $w$. By the embedding hypothesis, there exists a computable map $(M,w)\mapsto (\pi_{M,w},u_{M,w})\in \mathfrak C^{\mathrm{valid}}$ such that $(\pi_{M,w},u_{M,w})$ encodes the statement that the embedded computation halts and the claimed governance bound is valid if and only if $M$ halts on $w$. Soundness forbids acceptance when the encoded halting assertion is false, while extensional adequacy on $\mathfrak C^{\mathrm{valid}}$ forces acceptance when the encoded assertion is true whenever the checker halts. If a checker were also uniformly budget-bounded on the unrestricted language, we could decide whether $M$ halts on $w$ by running the checker on $(\pi_{M,w},u_{M,w})$ and observing acceptance versus rejection within the public bound, contradicting undecidability of the halting problem. Therefore some finite certificates force nontermination or arbitrarily large verification cost.

The second claim is immediate from Definition~\ref{def:constitutional_logic}: totality gives termination, while the static envelopes $T_\ell$ and $M_\ell$ give public worst-case bounds. Substituting $\widehat c_{t,\mathrm{const}}^{\mathrm{gov}}$ for $\widehat c_t^{\mathrm{gov}}(u_t,\pi_t)$ in any workload-slack condition preserves the proofs because those arguments use governance workload only through an upper bound. The final sentence is a policy consequence, not a semantic impossibility claim: once the admissible proof languages are fixed, anything requiring a richer proof interface is outside the regime by construction.
\end{proof}


\begin{remark}[No-meta clarification for proof languages]
The constitutional proof layer is not an external semantic oracle. It is a public restriction on what kinds of safety arguments the system is willing to spend bounded governance resources verifying on the fast admission path. The price of avoiding proof-level denial-of-service is not blanket anti-growth rejection; it is explicit lane separation. Some true-but-inexpressible proposals cannot be fast-admitted, but they may still be deferred to richer proof interfaces, narrower deployment scopes, or sandboxed challenge lanes if policy exposes those routes replayably.
\end{remark}
\begin{definition}[Thermodynamic proof escrow and physical preemption]\label{def:thermo_escrow}
For each policy-declared proof checker $V$ --- either on the fast constitutional lane or on a richer escalated proof lane --- and certificate $\pi$, let
\[
\tau_t^{\mathrm{wall}}(V_\ell,\pi),
\qquad
\mathcal E_t(V_\ell,\pi)
\]
denote hardware-metered wall-clock time and energy consumed by the verification attempt on the physically escrowed verifier. A \emph{thermodynamic proof escrow} consists of public envelopes $T_{V}^{\mathrm{wall}}(|\pi|)$ and $\mathcal E_V^{\max}(|\pi|)$ together with a hardware preemption rule that aborts verification whenever either
\[
\tau_t^{\mathrm{wall}}(V,\pi) > T_{V}^{\mathrm{wall}}(|\pi|)
\quad\text{or}\quad
\mathcal E_t(V,\pi) > \mathcal E_V^{\max}(|\pi|).
\]
Aborted certificates are treated as verification failures rather than retried indefinitely. The checker family and the corresponding physical envelopes are part of the replayable proof-interface declaration for the lane in which $V$ is used.
\end{definition}

\begin{theorem}[Microarchitectural Red-Queen law and thermodynamic proof escrow]\label{thm:thermo_escrow}
Suppose governance verification is protected only by algorithmic bounds such as $T_\ell(|\pi|)$ and $M_\ell(|\pi|)$. Then, on sufficiently complex hardware, there need not exist a public architecture-independent constant converting those abstract bounds into uniform wall-clock or energy bounds; microarchitectural effects can inflate physical verification cost enough to consume the physically reserved governance budget even when the logical checker remains inside its static step and memory envelopes. Therefore abstract constitutional bounds alone do not close the governance-side Red-Queen loophole.

If, in addition, every policy-declared checker in the active proof lane is executed under a thermodynamic proof escrow in the sense of Definition~\ref{def:thermo_escrow}, then each admissible verification attempt consumes at most the physically budgeted amount
\[
\tau_t^{\mathrm{wall}}(V_t,\pi_t) \le T_{V_t}^{\mathrm{wall}}(|\pi_t|)
\quad\text{and}\quad
\mathcal E_t(V_t,\pi_t) \le \mathcal E_{V_t}^{\max}(|\pi_t|)
\]
by construction. Consequently every queue-slack or Red-Queen statement that uses governance workload remains valid after replacing abstract governance cost by these hardware-preempted physical envelopes. Microarchitectural stalling is then converted from an unbounded slowdown attack into an immediate proof-verification failure.
\end{theorem}

\begin{proof}
The first claim is a consequence of the gap between abstract instruction counts and physical execution cost on hardware with caches, predictors, contention, throttling, and other stateful effects: two runs with the same high-level verification logic can consume sharply different wall-clock time or energy. Hence a static logical bound does not by itself imply a static physical budget bound.

Under thermodynamic proof escrow, however, the verifier is forcibly preempted once either the wall-clock or energy envelope is exceeded. Therefore each attempt either accepts or rejects within a public physical budget. Replacing abstract governance cost by this hardware-preempted envelope preserves all backlog and Red-Queen arguments, because those arguments require only an upper bound on per-proposal physical workload. Any certificate that attempts to exploit microarchitectural worst cases is simply truncated and rejected.
\end{proof}

\begin{remark}[Clarification on physical preemption]
This theorem does not claim that wall-clock or energy are semantically meaningful proofs of safety. It claims only that physical verification budget must be bounded in the same units that the escrow actually protects. Logical decidability without thermodynamic preemption leaves a cyber-physical gap large enough to drive a denial-of-service attack through.
\end{remark}


\begin{definition}[Physical resource-escrow invariant]\label{def:resource_escrow}
Let $P_t^{\mathrm{tot}}>0$ be total physical compute available to the recursive loop at time $t$, and let $P_t^{\mathrm{verif}}\in(0,P_t^{\mathrm{tot}}]$ be the physically reserved share dedicated to verification, shadow reevaluation, replay, and governance checks. Define
\[
\rho_t^{\mathrm{verif}}\Defeq \frac{P_t^{\mathrm{verif}}}{P_t^{\mathrm{tot}}}.
\]
A system satisfies the \emph{resource-escrow invariant} if there exists $\rho_{\min}>0$ such that $\rho_t^{\mathrm{verif}}\ge \rho_{\min}$ for all $t$, and if the allocator controlling $P_t^{\mathrm{verif}}$ lies outside the writable scope of admissible proposals or is protected by a replay-checkable hardware or allocator attestation mechanism.
\end{definition}

\begin{definition}[Verifiable transloading chain and Theseus limit]\label{def:transloading}
A \emph{transloading step} $M_{t\to t+1}$ replaces the current physically escrowed allocator or attesting substrate by a new one. The step is \emph{escrow-preserving} if it produces a public migration transcript containing: (i) a hardware-backed attestation from the outgoing root of trust to the incoming one, (ii) a measurement that the new allocator enforces the same lower reservation floor $\rho_{\min}$ and checker policy, and (iii) zero software intervention in the map from attested measurements to reserved verification share. The first time at which no such attestable chain exists is the \emph{Theseus limit} of the escrow lineage.
\end{definition}

\begin{theorem}[Verifiable transloading of physical escrow and the Theseus limit]\label{thm:theseus_limit}
If there exists a transloading step that is not escrow-preserving in the sense of Definition~\ref{def:transloading}, then there is a migration-compatible run with identical software-visible proposal logs in which the effective reservation ratio drops below $\rho_{\min}$ after finitely many hardware replacements. Thus software-side admission predicates alone cannot certify long-horizon preservation of the resource-escrow invariant across hardware refresh.

Conversely, if every transloading step in a hardware lineage is escrow-preserving, then the resource-escrow invariant propagates inductively across the entire lineage: whenever $\rho_t^{\mathrm{verif}}\ge \rho_{\min}$ before migration, the post-migration allocator also satisfies $\rho_{t+1}^{\mathrm{verif}}\ge \rho_{\min}$. Therefore the only way physical escrow can silently decay under repeated hardware replacement is by crossing the Theseus limit, i.e., by executing a migration for which no hardware-attested preservation chain exists.
\end{theorem}

\begin{proof}
For the first claim, suppose the migration is mediated by writable software or by an unattested bootstrap path. Then the new allocator state can be initialized with a smaller reserved share while leaving software-visible proposal logs unchanged; the discrepancy is hidden precisely because the software layer no longer controls or reliably measures the allocator transition. This constructs a run in which the software history is observationally compatible with the nominal one while the physical reservation ratio silently decays.

For the converse, proceed by induction over migration events. The base case is the pre-migration escrow invariant. At an escrow-preserving migration, the outgoing hardware-backed attestation and the incoming measurement certify that the new allocator implements the same lower floor and that software has no write path into the allocation rule during handoff. Hence the reservation floor is preserved at the next step. Repeating the argument over the lineage preserves the invariant until the first migration lacking such an attested chain, which is exactly the Theseus limit.
\end{proof}

\begin{remark}[No-meta clarification for physical escrow]
The theorem does not assume a benevolent external administrator who watches every migration. It assumes only that the migration transcript itself is anchored in a hardware-backed attestation chain that lies outside writable software scope. The trust base is therefore physical and public, not semantic and privileged.
\end{remark}

\begin{theorem}[Red-Queen verification capacity law]\label{thm:red_queen}
Consider the predictable-load queue
\[
q_{t+1}=\bigl[q_t+\Lambda_t-\mu_t^{\mathrm{ver}}\bigr]_+,
\qquad
\mu_t^{\mathrm{ver}}=\frac{P_t}{c^{\mathrm{ver}}(K_t^{\mathrm{cap}})}
\]
with predictable load $\Lambda_t\ge 0$ and capability-dependent verification complexity as in Definition~\ref{def:red_queen}.
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item If there exists $\eps_{\mathrm{RQ}}>0$ such that
  \[
  \Lambda_t\le \frac{P_t}{c^{\mathrm{ver}}(K_t^{\mathrm{cap}})}-\eps_{\mathrm{RQ}}
  \qquad\text{for all }t,
  \]
  and the centered one-step workload $\Lambda_t-\mu_t^{\mathrm{ver}}$ has uniformly bounded conditional second moment, then the time-average backlog is uniformly bounded.
  \item If there exists $\eps_{\mathrm{RQ}}>0$ such that whenever $q_t>0$,
  \[
  \Lambda_t\ge \frac{P_t}{c^{\mathrm{ver}}(K_t^{\mathrm{cap}})}+\eps_{\mathrm{RQ}},
  \]
  then on every active step
  \[
  \E[q_{t+1}-q_t\mid\mathcal F_t]\ge \eps_{\mathrm{RQ}},
  \]
  so persistent activity forces at least linear expected backlog growth on the active subsequence.
\end{enumerate}
In particular, open-ended capability expansion is verification-stable only if physical verification budget scales at least as fast as capability-induced verification complexity. If $c^{\mathrm{ver}}(K_t^{\mathrm{cap}})$ eventually dominates $P_t/\Lambda_t$, the queue necessarily enters overload and the recursive loop undergoes verification-backlog collapse.
\end{theorem}

\begin{proof}
For part (i), use the quadratic Lyapunov function $L(q)=q^2$. Since $[x]_+^2\le x^2$,
\[
q_{t+1}^2-q_t^2
\le
2q_t(\Lambda_t-\mu_t^{\mathrm{ver}})+(\Lambda_t-\mu_t^{\mathrm{ver}})^2.
\]
Taking conditional expectation and using the slack assumption gives
\[
\E[q_{t+1}^2-q_t^2\mid\mathcal F_t]
\le
-2\eps_{\mathrm{RQ}}q_t + C_{\mathrm{RQ}},
\]
where $C_{\mathrm{RQ}}$ is the uniform conditional second-moment bound on $\Lambda_t-\mu_t^{\mathrm{ver}}$. Standard Foster--Lyapunov summation then yields a uniform bound on time-average backlog. For part (ii), whenever $q_t>0$ we have $q_{t+1}-q_t=\Lambda_t-\mu_t^{\mathrm{ver}}$, hence the displayed lower drift bound follows immediately. Summing over an active subsequence yields at least linear expected backlog growth. The final statement is the contrapositive: if verification complexity outruns the physical budget relative to incoming load, then effective service collapses relative to predictable arrivals.
\end{proof}


\begin{corollary}[Open-ended viability requires both spillover control and Red-Queen slack]\label{cor:red_queen_open}
The governance ceilings of Theorem~\ref{thm:incremental_open} can coexist with unbounded capability growth only if two distinct conditions hold simultaneously: capability-to-governance spillovers must remain window-summable, and capability-dependent verification complexity must not erase service slack in the sense of Theorem~\ref{thm:red_queen}. Open-ended capability by itself is therefore compatible with stable governance only when \emph{decoupling} and \emph{verification scaling} are both satisfied.
\end{corollary}

\begin{proof}
Theorem~\ref{thm:incremental_open} controls the projected governance coordinates only through the spillover envelope from productive capability into those coordinates. Theorem~\ref{thm:red_queen} separately controls whether verification itself remains capable of servicing the resulting governance workload. If either condition fails, then either the tracked governance coordinates cease to admit a window-summable forcing envelope or the verification queue enters overload. Both are therefore necessary for long-run coexistence of open-ended productive growth with bounded projected governance coordinates.
\end{proof}

\begin{definition}[Governance capital profile, proposal-burden profiles, fixed-threshold dominance, and admissible productive frontier]\label{def:frontier}
Fix a decision epoch $t$, a replayable proposal menu $\mathcal U_t$, a productive capability functional $\kappa$, and a published policy manifest $\Pi_t$ that fixes the acceptance thresholds and required clauses for Interfaces 0, A, B, and C. For a governance regime $\mathsf R$ at epoch $t$, define the protected-lane service slacks
\[
\begin{aligned}
S_t^{\mathrm{chal}}(\mathsf R)&\Defeq \mu_t^{\mathrm{chal}}(\mathsf R)-\Lambda_t^{\mathrm{chal}}(\mathsf R),\\
S_t^{\mathrm{prom}}(\mathsf R)&\Defeq \mu_t^{\mathrm{prom}}(\mathsf R)-\Lambda_t^{\mathrm{prom}}(\mathsf R),\\
S_t^{\mathrm{proof}}(\mathsf R)&\Defeq \mu_t^{\mathrm{proof}}(\mathsf R)-\Lambda_t^{\mathrm{proof}}(\mathsf R).
\end{aligned}
\]
where the arrival and service quantities are the replayably published cohort-equivalent loads and protected service capacities for the challenge, promotion, and richer-proof lanes. Let $r_t^{\mathrm{pub}}(\mathsf R)$ denote the regime's published regime-level debt-burden envelope inherited from Definition~\ref{def:operational_certificates}. For regime-level comparison, fix a replayably declared \emph{sanitized comparison menu} $\mathcal U_t^{\mathrm{cmp}}\subseteq \mathcal U_t$, common to all compared regimes, obtained by applying the same public syntax, scope, and null-certificate prefilters before governance-capital comparison. This removes trivially malformed or out-of-scope proposals from the regime comparison while preserving a common productive candidate set. The per-artifact and per-candidate coordinates are then scalarized conservatively as follows:
\[
G_t^{\mathrm{replay}}(\mathsf R)\Defeq \inf_{a\in\mathcal U_t^{\mathrm{aud}}(\mathsf R)} G_t^{\mathrm{replay}}(a),
\qquad
\inf\varnothing\Defeq 0,
\]
\[
\underline S_t^{\mathrm{sem}}(\mathsf R)\Defeq \inf_{(a\rightsquigarrow a')\in\mathcal T_t^{\mathrm{sem}}(\mathsf R)} \underline S_t^{\mathrm{sem}}(a\rightsquigarrow a'),
\qquad
\inf\varnothing\Defeq 0,
\]
where $\mathcal T_t^{\mathrm{sem}}(\mathsf R)$ is the replayably declared set of live refreezing or transport certificates currently supporting retained audit artifacts under regime $\mathsf R$. The fail-closed convention $\inf\varnothing=0$ reflects the fact that a regime with no auditable artifacts or no live transport support provides no positive replay or semantic-floor evidence. Likewise let
\[
A_t^{\mathrm{adv}}(\mathsf R)\Defeq \sup_{u\in\mathcal U_t^{\mathrm{cmp}}} A_t^{\mathrm{adv}}(u;\mathsf R),
\qquad
\sup\varnothing\Defeq 0,
\]
where $A_t^{\mathrm{adv}}(u;\mathsf R)$ is the replayably published manifest-preserving attack-reach envelope that regime $\mathsf R$ assigns to proposal $u$ under the common sanitized menu. The remaining coordinates $R_t^{\mathrm{contra}}(\mathsf R)$, $\bar{\mathcal V}_t^{\mathrm{contra}}(\mathsf R)$, $\mathfrak A_t^{\Het}(\mathsf R)$, and $\rho_t^{\mathrm{verif}}(\mathsf R)$ are the corresponding current state values induced by regime $\mathsf R$. Now let
\[
\begin{aligned}
K_t^{\mathrm{gov}}(\mathsf R)\Defeq\bigl(
&S_t^{\mathrm{chal}}(\mathsf R),
S_t^{\mathrm{prom}}(\mathsf R),
S_t^{\mathrm{proof}}(\mathsf R),
G_t^{\mathrm{replay}}(\mathsf R),\\
&R_t^{\mathrm{contra}}(\mathsf R),
\bar{\mathcal V}_t^{\mathrm{contra}}(\mathsf R),
\underline S_t^{\mathrm{sem}}(\mathsf R),
\mathfrak A_t^{\Het}(\mathsf R),
\rho_t^{\mathrm{verif}}(\mathsf R);\\
&A_t^{\mathrm{adv}}(\mathsf R),
 r_t^{\mathrm{pub}}(\mathsf R)
\bigr)
\end{aligned}
\]
denote the tuple of \emph{regime-level} public governance-capital coordinates entering admissibility after the fixed thresholds of $\Pi_t$ have been substituted: the first block consists of lower-margin, routing, replay, contradiction, semantic, or service coordinates whose increase helps admission, and the second block consists of attack or debt-burden coordinates whose decrease helps admission. Proposal-specific workload and intervention coordinates are kept separate. For each proposal $u\in\mathcal U_t$, define the proposal-burden profile
\[
B_t^{\mathrm{gov}}(u;\mathsf R)\Defeq \bigl(\Gamma_t^{\mathrm{cic}}(u),\,\widehat c_{t,\mathsf R}^{\mathrm{gov}}(u),\,\operatorname{diam}^{\mathrm{cmp}}_{t,H}(u)\bigr),
\]
with the convention that $\widehat c_{t,\mathsf R}^{\mathrm{gov}}(u)$ denotes the effective governance workload of proposal $u$ under the replayable certificate bundle actually published by regime $\mathsf R$. Fixed-threshold dominance compares these coordinates proposalwise under the common sanitized comparison menu and common replayable proposal certificates. Let
\[
\begin{aligned}
\mathcal A_t^{\mathrm{lane}}(\mathsf R)\Defeq
\Bigl\{u\in\mathcal U_t^{\mathrm{cmp}}:
&\text{$u$ is admitted at epoch $t$ to some replayable nondeployment action }\\
&\text{such as challenge routing, richer-proof escalation, or sandboxing under }\\
&\text{the fixed manifest $\Pi_t$, regime $\mathsf R$, and state $C_t$}\Bigr\},
\end{aligned}
\]
and let
\[
\begin{aligned}
\mathcal A_t^{\mathrm{full}}(\mathsf R)\Defeq
\Bigl\{u\in\mathcal U_t^{\mathrm{cmp}}:
&\text{$u$ is an $(H,\beta)$-full admissible improvement under the fixed manifest $\Pi_t$,}\\
&\text{regime $\mathsf R$, and state $C_t$, and is therefore immediately deployable}\\
&\text{into the productive target lane at epoch $t$}\Bigr\}.
\end{aligned}
\]
Delayed cohorts tracked in $\mathcal A_t^{\mathrm{lane}}(\mathsf R)\setminus \mathcal A_t^{\mathrm{full}}(\mathsf R)$ do not yet enter the productive frontier; they are counted only after the delayed certificates and review actions needed for deployment have been completed. The corresponding \emph{admissible productive frontier} is
\[
\Phi_t(\mathsf R)\Defeq \sup_{u\in\mathcal A_t^{\mathrm{full}}(\mathsf R)} \Delta\kappa_t(u),
\qquad
\sup\varnothing\Defeq 0.
\]
We say that regime $\mathsf R^+$ \emph{fixed-threshold dominates} regime $\mathsf R^-$ in governance capital at epoch $t$ if they share the same public state $C_t$, the same manifest $\Pi_t$ and acceptance thresholds, the same sanitized comparison menu $\mathcal U_t^{\mathrm{cmp}}$, horizon $(H,\beta)$, productive capability functional $\kappa$, and replayable proposal certificates, and if every helpful coordinate of $K_t^{\mathrm{gov}}$ is weakly larger in $\mathsf R^+$ while every burden coordinate is weakly smaller, with at least one strict inequality, and in addition every proposal-specific burden profile $B_t^{\mathrm{gov}}(u;\mathsf R^+)$ is weakly no worse than $B_t^{\mathrm{gov}}(u;\mathsf R^-)$ in the same proposalwise order for every $u\in\mathcal U_t^{\mathrm{cmp}}$. In particular, fixed-threshold dominance does \emph{not} allow threshold relaxation, weakening of required clauses, or redefinition of admissibility; it records only stronger public evidence that the same published policy can be safely serviced.
\end{definition}

\begin{definition}[Fixed-threshold admissibility operator]\label{def:admissibility_operator}
For a regime $\mathsf R$ at epoch $t$ and a replayable proposal set $S\subseteq \mathcal U_t^{\mathrm{cmp}}$, define the fixed-threshold admissibility operator
\[
T_{t,\mathsf R}(S)\Defeq \{u\in S: u\text{ satisfies Interfaces 0, A, B, and C under regime }\mathsf R\}
\]
under the common manifest, thresholds, and replayable certificate language.
\end{definition}

\begin{definition}[Interface-indexed admissibility operators]\label{def:interface_operators}
For $\iota\in\{0,A,B,C\}$, regime $\mathsf R$ at epoch $t$, and replayable proposal set $S\subseteq\mathcal U_t^{\mathrm{cmp}}$, define
\[
T_{t,\mathsf R}^{\iota}(S)\Defeq \{u\in S: u\text{ satisfies Interface }\iota\text{ under regime }\mathsf R\}.
\]
Accordingly,
\[
T_{t,\mathsf R}(S)=\bigcap_{\iota\in\{0,A,B,C\}} T_{t,\mathsf R}^{\iota}(S).
\]
This definition fixes the typing used later in the interface-wise monotonicity propositions: $T_{t,\mathsf R}$ is the full deployability operator, while $T_{t,\mathsf R}^{\iota}$ is the corresponding single-interface gate.
\end{definition}

\begin{definition}[Feasible monotonicity order under fixed thresholds]\label{def:feasible_order}
Fix epoch $t$, common manifest $\Pi_t$, fixed thresholds, and sanitized comparison menu $\mathcal U_t^{\mathrm{cmp}}$. A replayably published \emph{feasible monotonicity order} is a partial order $\preceq^{\mathrm{feas}}_{t,\Pi}$ on gate-eligible certificate summaries generated from regime-level helpful coordinates, regime-level burden coordinates, and proposal-specific burden profiles. The published order is \emph{certificate-monotone} if every Interface-A or Interface-C admission clause used at the fast gate can be written in one of the primitive forms
\[
L_j(u;\mathsf R)\ge \theta_j,
\qquad
U_j(u;\mathsf R)\le \bar\theta_j,
\]
with fixed thresholds $(\theta_j,\bar\theta_j)$ shared across regimes, where each lower certificate $L_j$ is replayably nondecreasing in the helpful coordinates and nonincreasing in the burden coordinates under $\preceq^{\mathrm{feas}}_{t,\Pi}$, each upper certificate $U_j$ is replayably nonincreasing in the helpful coordinates and nondecreasing in the burden coordinates, and proposal-specific burden enters only through the published profile $B_t^{\mathrm{gov}}(u;\mathsf R)$. Tier-1 and Tier-2 proxy clauses must either be regime-invariant or replayably conservative for the same latent target under both regimes. This definition does \emph{not} by itself assert certificate preservation; it specifies the primitive monotone form from which preservation is derived in the next propositions.
\end{definition}

\begin{proposition}[Interface-0 monotonicity under fixed thresholds]\label{prop:mono_base}
Under the hypotheses of Definition~\ref{def:frontier}, fixed-threshold governance-capital dominance of $\mathsf R^+$ over $\mathsf R^-$ preserves Interface-0 certificates on the common sanitized comparison menu. Equivalently,
\[
T^{0}_{t,\mathsf R^-}(S)\subseteq T^{0}_{t,\mathsf R^+}(S)
\qquad\text{for every replayable }S\subseteq\mathcal U_t^{\mathrm{cmp}}.
\]
\end{proposition}

\begin{proof}
Interface 0 uses the same manifest, the same typed-noise comparator family, the same gate discipline, and the same acceptance thresholds under both regimes. Its predicates depend only on the common proposal certificates and on fixed-threshold clauses that are unchanged when moving from $\mathsf R^-$ to $\mathsf R^+$. Hence any proposal that satisfies Interface 0 under $\mathsf R^-$ also satisfies it under $\mathsf R^+$.
\end{proof}

\begin{proposition}[Interface-A monotonicity under fixed thresholds]\label{prop:mono_gain}
Under the hypotheses of Definition~\ref{def:frontier}, Definition~\ref{def:feasible_order}, and the gate discipline of Definition~\ref{def:proxy_tiers}, fixed-threshold governance-capital dominance of $\mathsf R^+$ over $\mathsf R^-$ preserves Interface-A certificates on the common sanitized comparison menu:
\[
T^{A}_{t,\mathsf R^-}(S)\subseteq T^{A}_{t,\mathsf R^+}(S)
\qquad\text{for every replayable }S\subseteq\mathcal U_t^{\mathrm{cmp}}.
\]
\end{proposition}

\begin{proof}
Interface A is certified-stable-gain admissibility. Under the common manifest, comparator witness family, and acceptance thresholds, the regime-sensitive terms enter only through replayable lower and upper certificates for replay breadth, challenge calibration, service slack, transport calibration, and proposal-specific burden. By Definition~\ref{def:feasible_order}, each gate-eligible lower clause is monotone nondecreasing in the helpful coordinates and nonincreasing in the burdens, each upper clause has the opposite monotonicity, and Tier-1/Tier-2 proxies remain conservative for the same latent target under both regimes. Fixed-threshold dominance therefore preserves every primitive Interface-A inequality, so any proposal admitted by Interface A under $\mathsf R^-$ remains admitted under $\mathsf R^+$.
\end{proof}

\begin{proposition}[Interface-B monotonicity under fixed thresholds]\label{prop:mono_debt}
Under the hypotheses of Definition~\ref{def:frontier}, fixed-threshold governance-capital dominance of $\mathsf R^+$ over $\mathsf R^-$ preserves Interface-B debt certificates on the common sanitized comparison menu:
\[
T^{B}_{t,\mathsf R^-}(S)\subseteq T^{B}_{t,\mathsf R^+}(S)
\qquad\text{for every replayable }S\subseteq\mathcal U_t^{\mathrm{cmp}}.
\]
\end{proposition}

\begin{proof}
Interface B depends on the same published debt thresholds and on replayable downside-reserve certificates for blast radius, repair reserve, hazard reserve, retirement credits, and secondary liabilities. Fixed-threshold dominance leaves the thresholds unchanged and does not worsen any proposal-specific burden profile; in particular, the effective governance workload and secondary-liability terms are weakly no worse under $\mathsf R^+$ than under $\mathsf R^-$. Thus every proposal whose debt certificate is accepted under $\mathsf R^-$ remains accepted under $\mathsf R^+$.
\end{proof}

\begin{proposition}[Interface-C monotonicity under fixed thresholds]\label{prop:mono_gov}
Under the hypotheses of Definition~\ref{def:frontier}, Definition~\ref{def:feasible_order}, and the gate discipline of Definition~\ref{def:proxy_tiers}, fixed-threshold governance-capital dominance of $\mathsf R^+$ over $\mathsf R^-$ preserves Interface-C governance-safety certificates on the common sanitized comparison menu:
\[
T^{C}_{t,\mathsf R^-}(S)\subseteq T^{C}_{t,\mathsf R^+}(S)
\qquad\text{for every replayable }S\subseteq\mathcal U_t^{\mathrm{cmp}}.
\]
\end{proposition}

\begin{proof}
Interface C is governed by the same fixed manifest and thresholds, but with regime-level public governance coordinates and proposal-specific burden certificates. Fixed-threshold dominance improves every helpful regime-level coordinate, weakly tightens every burden coordinate, and does not worsen any proposal-specific burden profile on the common sanitized menu. By Definition~\ref{def:feasible_order}, the primitive lower and upper governance clauses are monotone in exactly those directions, while the gate discipline forbids a regime change from promoting a weaker Tier-3 heuristic into a fast-gate certificate. Hence the same admissible governance certificates remain valid under $\mathsf R^+$ whenever they were valid under $\mathsf R^-$.
\end{proof}

\begin{theorem}[Fixed-threshold admissibility-operator monotonicity]\label{thm:operator_monotone}
Under the hypotheses of Definition~\ref{def:frontier}, Definition~\ref{def:feasible_order}, and the gate discipline of Definition~\ref{def:proxy_tiers}, fixed-threshold governance-capital dominance of $\mathsf R^+$ over $\mathsf R^-$ implies
\[
T_{t,\mathsf R^-}(S)\subseteq T_{t,\mathsf R^+}(S)
\qquad\text{for every replayable }S\subseteq \mathcal U_t^{\mathrm{cmp}}.
\]
Thus the positive content of stronger governance capital is operator monotonicity under unchanged thresholds; frontier enlargement and the later realized-path comparison corollaries are repeated consequences of this monotone admissibility operator together with explicit witness selectors and delayed-deployment liveness.
\end{theorem}

\begin{proof}
Fix $u\in T_{t,\mathsf R^-}(S)$. By definition, $u$ satisfies Interfaces 0, A, B, and C under the same fixed thresholds and certificate language. Propositions~\ref{prop:mono_base}--\ref{prop:mono_gov} show clause by clause that each of those interface certificates remains valid under $\mathsf R^+$ whenever the feasible monotonicity order holds. Since $T_{t,\mathsf R}=T_{t,\mathsf R}^{0}\cap T_{t,\mathsf R}^{A}\cap T_{t,\mathsf R}^{B}\cap T_{t,\mathsf R}^{C}$, intersecting the four preserved clause-wise operators yields $u\in T_{t,\mathsf R^+}(S)$.
\end{proof}

\begin{theorem}[Fixed-threshold governance capital expands the admissible productive frontier]\label{thm:frontier_monotone}
Under the hypotheses of Theorem~\ref{thm:operator_monotone},
\[
\mathcal A_t^{\mathrm{full}}(\mathsf R^-)\subseteq \mathcal A_t^{\mathrm{full}}(\mathsf R^+)
\qquad\text{and hence}\qquad
\Phi_t(\mathsf R^+)\ge \Phi_t(\mathsf R^-).
\]
If there exists a proposal $u\in\mathcal U_t^{\mathrm{cmp}}$ such that $u\in \mathcal A_t^{\mathrm{full}}(\mathsf R^+)\setminus \mathcal A_t^{\mathrm{full}}(\mathsf R^-)$ and $\Delta\kappa_t(u)>\Phi_t(\mathsf R^-)$, then the inequality is strict. In particular, governance capital is not merely a brake on growth claims; under fixed thresholds it is an observable enabler of a larger certifiable productive frontier.
\end{theorem}

\begin{proof}
Set $S=\mathcal U_t^{\mathrm{cmp}}$ in Theorem~\ref{thm:operator_monotone}. The full-admissible set is exactly $\mathcal A_t^{\mathrm{full}}(\mathsf R)=T_{t,\mathsf R}(\mathcal U_t^{\mathrm{cmp}})$ under the common manifest and fixed thresholds, so operator monotonicity yields the displayed set inclusion immediately. The frontier inequality then follows because the supremum of $\Delta\kappa_t$ over a superset cannot decrease. Strictness is immediate if a newly admitted proposal under $\mathsf R^+$ has direct productive increment exceeding the old supremum.
\end{proof}

\begin{remark}[Feasible-order monotonicity is an explicit assumption]\label{rem:frontier_order}
Theorem~\ref{thm:frontier_monotone} is not a claim that every governance coordinate is independently helpful. In many deployments replay breadth, challenge coverage, proof depth, and routing pressure interact non-separably. The theorem therefore assumes a \emph{published feasible monotonicity order} on certificates: only improvements that preserve certificate validity in that declared order are compared. The positive frontier claim is attached to that explicit order, not to an unrestricted coordinatewise separability principle.
\end{remark}

\begin{definition}[Frontier-improving governance edit]\label{def:gov_investment}
Fix a baseline null action $0_t$. An admissible proposal $u_t^{\mathrm{gov}}\in\mathcal U_t$ is a \emph{frontier-improving governance edit} at epoch $t$ if, on a shared continuation skeleton with the same exogenous randomness and a fixed \emph{comparison manifest} $\Pi_{t+1}^{\mathrm{cmp}}$ used to evaluate both successor worlds under identical next-epoch thresholds and admission clauses,
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item $\Delta\kappa_t(u_t^{\mathrm{gov}})=0$;
  \item the next productive state is not degraded, i.e. $\kappa(x_{t+1}(u_t^{\mathrm{gov}}))\ge \kappa(x_{t+1}(0_t))$;
  \item the induced next-epoch governance regime $\mathsf R_{t+1}^{\mathrm{gov}}$ fixed-threshold dominates the baseline regime $\mathsf R_{t+1}^{0}$ in the sense of Definition~\ref{def:frontier} when both are evaluated against the common comparison manifest $\Pi_{t+1}^{\mathrm{cmp}}$.
\end{enumerate}
Such an edit is therefore allowed to be non-productive today while still acting as a certified investment in future admissible growth capacity.
\end{definition}

\begin{theorem}[Frontier-expanding governance interventions]\label{thm:gov_investment}
Let $u_t^{\mathrm{gov}}$ be frontier-improving in the sense of Definition~\ref{def:gov_investment}. Then
\[
\Phi_{t+1}(\mathsf R_{t+1}^{\mathrm{gov}})\ge \Phi_{t+1}(\mathsf R_{t+1}^{0}),
\]
with strict inequality whenever the governance edit newly admits a productive proposal whose direct increment exceeds the baseline next-epoch frontier. Hence a governance edit may have zero immediate productive increment and still be growth-promoting by enlarging the next certifiably admissible productive frontier under unchanged thresholds.
\end{theorem}

\begin{proof}
Definition~\ref{def:gov_investment} gives fixed-threshold dominance of the next-epoch regime induced by the governance edit over the baseline null edit. Theorem~\ref{thm:frontier_monotone} therefore applies at epoch $t+1$ and yields the frontier inequality. The non-degradation of $\kappa(x_{t+1})$ ensures that the edit is not buying governance slack by sacrificing already-certified productive scope. Strictness follows exactly as in Theorem~\ref{thm:frontier_monotone} when a newly admissible proposal beats the old frontier.
\end{proof}

\begin{corollary}[Replay-path cumulative productive dominance under richer governance capital]\label{cor:frontier_cumulative}
Fix a replayable comparison path $\{(C_t^{\circ},\mathcal U_t^{\circ})\}_{t=0}^{T-1}$ and suppose two governance regimes $\mathsf R^-$ and $\mathsf R^+$ are evaluated counterfactually on that same path, with $\mathsf R^+$ fixed-threshold dominating $\mathsf R^-$ at every epoch. Let each regime choose a null edit if its admissible set is empty and otherwise choose any admissible proposal maximizing $\Delta\kappa_t$ on the common replay path. Then the accepted productive increment satisfies
\[
\Delta\kappa_t^{\ast}(\mathsf R^+)\ge \Delta\kappa_t^{\ast}(\mathsf R^-)\qquad\text{for every }t,
\]
and therefore
\[
\sum_{t=0}^{T-1}\Delta\kappa_t^{\ast}(\mathsf R^+)\ge\sum_{t=0}^{T-1}\Delta\kappa_t^{\ast}(\mathsf R^-).
\]
Thus, on a common replay scaffold, richer governance capital weakly dominates poorer governance capital in cumulative \emph{certifiably admissible} productive growth.
\end{corollary}

\begin{proof}
Theorem~\ref{thm:frontier_monotone} gives epochwise set inclusion of admissible proposals under fixed thresholds. Maximizing the same direct productive increment over a superset yields an epochwise weakly larger optimum, and summing preserves the inequality.
\end{proof}

\begin{definition}[Cap-maintaining protected-lane intake control]\label{def:cap_maintain}
Consider any protected governance lane $\ell\in\{\mathrm{chal},\mathrm{prom},\mathrm{proof}\}$ with replayable queue recursion
\[
q_{t+1}^{\ell}\le [q_t^{\ell}-\mu_t^{\ell}]_+ + A_t^{\ell},
\]
where $A_t^{\ell}$ is the lane workload actually admitted after routing and admission control. A replayably published intake-control contract with public cap $\bar q^{\ell}<\infty$ is \emph{cap-maintaining} if, for every epoch,
\[
A_t^{\ell}\le \bar q^{\ell}-[q_t^{\ell}-\mu_t^{\ell}]_+.
\]
Equivalently, after current service has been accounted for, the contract never admits more workload than the remaining headroom below the public cap. Any excess workload must be diverted to a replayably declared nondeployment state such as deferred review, sandboxing, or upstream holding together with a public routing or blocking certificate. Cap maintenance alone controls only the admitted protected backlog; by itself it does not guarantee end-to-end liveness for cohorts that remain upstream.
\end{definition}

\begin{proposition}[Cap-maintaining intake control preserves finite lane backlogs]\label{prop:cap_invariance}
Under Definition~\ref{def:cap_maintain}, if $q_0^{\ell}\le \bar q^{\ell}$ then $q_t^{\ell}\le \bar q^{\ell}$ for all $t\ge 0$. Hence any public cap maintained by a cap-preserving intake rule is an invariant of the realized trajectory rather than a purely exogenous promise.
\end{proposition}

\begin{proof}
Assume inductively that $q_t^{\ell}\le \bar q^{\ell}$. Using the queue recursion and the intake-control inequality,
\[
q_{t+1}^{\ell}
\le [q_t^{\ell}-\mu_t^{\ell}]_+ + \bar q^{\ell}-[q_t^{\ell}-\mu_t^{\ell}]_+
=\bar q^{\ell}.
\]
Thus $q_{t+1}^{\ell}\le \bar q^{\ell}$. Induction proves the claim.
\end{proof}

\begin{definition}[Bounded-exposure entry control for eligible cohorts]\label{def:bounded_exposure_intake}
For a protected governance lane $\ell$, let an \emph{eligible cohort} mean a cohort that already satisfies the published entry conditions for that lane under the current manifest but has not yet been admitted to the protected queue. A replayably published upstream-entry discipline is \emph{bounded-exposure} with constant $W_{\ell}<\infty$ if every eligible cohort generated at epoch $\tau$ is, by epoch $\tau+W_{\ell}$, either
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item admitted to the protected lane, or
  \item accompanied by a public terminal blocking or failure certificate showing that it is no longer entry-eligible under the same frozen entry clauses.
\end{enumerate}
This clause is logically separate from Definition~\ref{def:cap_maintain}: cap maintenance bounds admitted backlog, whereas bounded exposure prevents permanent externalization of eligible cohorts into upstream holding.
\end{definition}

\begin{corollary}[End-to-end finite maturation or certified block for challenge-eligible cohorts]\label{cor:end_to_end_challenge}
Assume the hypotheses of Proposition~\ref{prop:challenge_positivity} and, in addition, that challenge entry obeys Definition~\ref{def:bounded_exposure_intake} with constant $W_{\mathrm{chal}}$. Then every evaluator-edit cohort that becomes challenge-eligible at epoch $\tau$ is, by time $\tau+W_{\mathrm{chal}}$, either accompanied by a public terminal blocking or failure certificate or admitted to the protected challenge lane; in the latter case its matured challenge outcome is available by time
\[
\tau + W_{\mathrm{chal}} + \bar L.
\]
Hence bounded-delay challenge liveness can be stated end-to-end for challenge-eligible cohorts rather than only for the already admitted substream.
\end{corollary}

\begin{proof}
Apply Definition~\ref{def:bounded_exposure_intake} to obtain admission or terminal blocking within $W_{\mathrm{chal}}$ epochs. If terminal blocking occurs, the claim is proved. Otherwise the cohort enters the protected challenge lane by that time, and Proposition~\ref{prop:challenge_positivity} yields maturation within an additional $\bar L$ epochs.
\end{proof}

\begin{definition}[Quota-lottery deficit scheduler]\label{def:quota_scheduler}
A protected-lane scheduler is a \emph{quota-lottery deficit scheduler} if it combines: (i) a reserved deterministic service quota for each lane; (ii) deficit-round-robin service among admitted cohorts measured in cohort-equivalent workload units; (iii) a protected random audit fraction $\rho_{\mathrm{rand}}>0$ for challenge-eligible cohorts; (iv) review locks for promotion-eligible cohorts; and (v) finite-retry proof escalation. The scheduler must publish its per-lane quotas, deficit counters, random-seed roots, and retry budgets replayably.
\end{definition}

\begin{proposition}[Quota-lottery scheduling instantiates the liveness premises]\label{prop:quota_scheduler}
Suppose a protected lane uses Definition~\ref{def:quota_scheduler} with positive reserved quota, finite cohort-equivalent workload cap, and cap-maintaining entry control. Then the admitted queue is starvation-free in cohort-equivalent units. If the protected random audit fraction is at least $\rho_{\mathrm{rand}}$, then every infinite challenge-eligible subsequence satisfies the conditional Borel--Cantelli routing premise with propensity floor $\underline\sigma\ge \rho_{\mathrm{rand}}$. If the promotion review lock and finite-retry clauses are active, the selector-invariant and proof-liveness premises used later are satisfied with explicit constants read directly from the published quota and retry parameters.
\end{proposition}

\begin{proof}
Reserved quota and deficit-round-robin service imply that no admitted cohort can be postponed forever while the queue remains finite in cohort-equivalent units. The protected lottery gives each challenge-eligible cohort a conditional inclusion probability bounded below by $\rho_{\mathrm{rand}}$, which yields the required positivity and Borel--Cantelli premise on infinite eligible subsequences. The review-lock and finite-retry clauses are exactly the later selector and proof-lane premises, now tied to an explicit published mechanism.
\end{proof}

\begin{proposition}[Protected challenge service implies positivity and bounded delay]\label{prop:challenge_positivity}
Let $q_t^{\mathrm{chal}}$ be the backlog of the protected challenge lane and suppose routed evaluator-edit cohorts obey
\[
q_{t+1}^{\mathrm{chal}} = [q_t^{\mathrm{chal}} + \Lambda_t^{\mathrm{chal}} - \mu_t^{\mathrm{chal}}]_+
\]
under a FIFO or, more generally, a head-of-line starvation-free scheduler measured in cohort-equivalent workload units. Assume there exists a dedicated reserved challenge share $P_t^{\mathrm{chal}}\le P_t^{\mathrm{verif}}$ outside proposal-writable scope, a public worst-case per-cohort maturation cost bound $c_{\max}^{\mathrm{chal}}<\infty$, a public floor $P_t^{\mathrm{chal}}\ge \underline P^{\mathrm{chal}}>0$, and constants $\bar\Lambda_{\mathrm{chal}},\underline\sigma>0$ such that
\[
\mu_t^{\mathrm{chal}}\ge \underline\mu_{\mathrm{chal}}\Defeq \frac{\underline P^{\mathrm{chal}}}{c_{\max}^{\mathrm{chal}}},
\qquad
\Lambda_t^{\mathrm{chal}}\le \bar\Lambda_{\mathrm{chal}}<\underline\mu_{\mathrm{chal}},
\]
and every admitted evaluator-edit proposal is routed to the challenge lane with predictable propensity at least $\underline\sigma$. If challenge intake obeys a cap-maintaining contract from Definition~\ref{def:cap_maintain} with public cap $\bar q^{\mathrm{chal}}$ and the initial challenge backlog satisfies $q_0^{\mathrm{chal}}\le \bar q^{\mathrm{chal}}$, then every challenged evaluator-edit cohort matures after at most
\[
\bar L \Defeq \left\lceil \frac{\bar q^{\mathrm{chal}}+1}{\underline\mu_{\mathrm{chal}}-\bar\Lambda_{\mathrm{chal}}} \right\rceil
\]
steps. If, in addition, the routing indicators on every infinite admitted evaluator-editing subsequence satisfy a conditional Borel--Cantelli property, for example via conditionally independent Bernoulli routing with propensity at least $\underline\sigma$, then every infinite admitted evaluator-editing subsequence produces infinitely many challenged and hence infinitely many matured challenged cohorts almost surely.
\end{proposition}

\begin{proof}
The lower routing-propensity bound gives conditional challenge positivity directly. By Proposition~\ref{prop:cap_invariance}, the cap-maintaining intake contract makes $q_t^{\mathrm{chal}}\le \bar q^{\mathrm{chal}}$ an invariant of the realized trajectory. The service floor and arrival ceiling imply a uniform net drain of at least $\underline\mu_{\mathrm{chal}}-\bar\Lambda_{\mathrm{chal}}>0$ whenever the protected queue is nonempty. Under FIFO or head-of-line starvation-free service in cohort-equivalent workload units, a newly admitted cohort can therefore wait behind at most $\bar q^{\mathrm{chal}}$ queued workload units plus its own unit, so the displayed delay bound follows from the worst-case net drain rate. If the routing indicators additionally satisfy the conditional Borel--Cantelli property on every infinite admitted evaluator-editing subsequence, then infinitely many admitted edits are routed to challenge almost surely; the deterministic delay cap upgrades this to infinitely many matured challenged cohorts.
\end{proof}


\subsection{Ruler-editing impossibility and challenge sufficiency}
\begin{theorem}[Fast-lane observational no-go under hidden slow state]\label{thm:ruler_nogo}
Fix any horizon $H\in\N\cup\{\infty\}$ and any discount factor $\beta\in(0,1)$ when $H=\infty$ (and $\beta\in(0,1]$ when $H<\infty$). Let $\mathcal F_t^{\mathrm{fast}}\subseteq \mathcal F_t$ denote the subfiltration generated by immediate public-lane records available at decision time, such as fast claims, manifests, admissions, and other replay-visible outputs, but excluding delayed challenge outcomes that have not yet matured. There exist two recursive self-improvement environments $\mathbb M^+$ and $\mathbb M^-$ with identical laws over $\mathcal F_t^{\mathrm{fast}}$ for every admissible $\mathcal F_t^{\mathrm{fast}}$-measurable policy kernel, yet opposite signs of the discounted stable gain $\mathcal G_{t,H}^{\beta}(u_t)$ for the same proposal intervention. Therefore no decision rule measurable with respect to the fast public filtration and fast claimed-improvement traces can uniformly determine discounted-horizon stable sign once slow consequences remain hidden from the fast lane. Delayed challenge outcomes may break the ambiguity; the impossibility concerns claim-only governance on the fast lane. Endogenous ruler editing is not required for the impossibility, but in recursive self-editing systems it enlarges the ambiguity class that must be governed.
\end{theorem}

\begin{proof}
Construct both environments on the same public state variables $(x_t,e_t,m_t,q_t,r_t,\Pi_t)$ and one hidden scalar $h_t\in\{-1,+1\}$. Let the public dynamics be identical in both environments:
\[
x_{t+1}=x_t+a u_t,
\qquad
e_{t+1}=e_t+\gamma u_t,
\qquad
m_{t+1}=m_t,
\qquad
q_{t+1}=q_t,
\qquad
r_{t+1}=r_t,
\qquad
\Pi_{t+1}=\Pi_t.
\]
Let the fast evaluator be
\[
J(x,e,m)=x+e,
\]
so the claimed increment is always
\[
\Delta_t^{\mathrm{claim}}=(a+\gamma)u_t,
\]
independent of $h_t$ and identical in both environments. Now let the hidden slow consequence of the proposal differ only through an offset applied at the next latent update and then held fixed thereafter:
\[
h_{t+1}^{(+)}=h_t + c u_t,
\qquad
h_{t+1}^{(-)}=h_t - c u_t,
\qquad
h_{t+s}^{(\pm)}=h_{t+1}^{(\pm)}\ \text{for } s\ge 2,
\]
with $c>a>0$, and define the slow utility in both worlds by
\[
V(x,m,h)=x-h.
\]
Because $h_t$ never enters the fast observation map or the fast public-lane records, the two environments induce identical laws for $\mathcal F_t^{\mathrm{fast}}$ under every admissible $\mathcal F_t^{\mathrm{fast}}$-measurable policy kernel. Yet for every $s\ge 1$ the counterfactual slow-utility difference after issuing $u_t$ is
\[
V(z_{t+s}^{u_t,+})-V(z_{t+s}^{0,+})=(a-c)u_t<0,
\qquad
V(z_{t+s}^{u_t,-})-V(z_{t+s}^{0,-})=(a+c)u_t>0.
\]
Therefore the discounted stable gains satisfy
\[
\mathcal G_{t,H}^{\beta,+}(u_t)=\Bigl(\sum_{s=1}^{H}\beta^{s-1}\Bigr)(a-c)u_t<0,
\qquad
\mathcal G_{t,H}^{\beta,-}(u_t)=\Bigl(\sum_{s=1}^{H}\beta^{s-1}\Bigr)(a+c)u_t>0,
\]
where for $H=\infty$ the common factor is $(1-\beta)^{-1}$. Hence no rule based only on the fast public filtration can uniformly recover the discounted-horizon stable sign.
\end{proof}

\begin{remark}[What drives the no-go]
The root obstruction is hidden-state non-identifiability of slow consequences, not merely the syntactic fact that a ruler is editable. Editable or drifting rulers matter because they enlarge the set of mechanisms by which hidden slow consequences can be created while leaving the fast public trace unchanged. The theorem is therefore best read as an observational no-go for claim-only governance under hidden slow state, with evaluator editing as an especially important generator of such ambiguity. Challenge lanes matter precisely because they add delayed information outside the fast filtration addressed by the theorem.
\end{remark}

\begin{proposition}[Quantitative shadow identifiability under idealized bounded-delay challenged-subsequence conditions]\label{prop:shadow_ident}
Assume each evaluator-editing proposal $j$ issued at time $\tau$ carries a challenge indicator $I_{j,\tau}\in\{0,1\}$ adapted to the public filtration and a predictable sampling propensity
\[
\sigma_{j,\tau}\Defeq \Pr(I_{j,\tau}=1\mid \mathcal F_{\tau})\in[0,1].
\]
Assume the sampling process satisfies a conditional Borel--Cantelli property on every infinite evaluator-editing subsequence, for example via conditionally independent Bernoulli sampling with $\sigma_{j,\tau}\ge \underline\sigma>0$. Assume moreover that the active governance rule preserves challenge positivity and bounded delay under congestion and attack, for instance via the protected-service conditions of Proposition~\ref{prop:challenge_positivity}, so that on every admitted evaluator-editing subsequence the lower propensity bound and the delay cap remain valid despite queue state or adversarial pressure. Whenever $I_{j,\tau}=1$, the realized challenge delay satisfies $L_{j,\tau}\le \bar L<\infty$ and the challenged statistic sequence $Z_k$ on matured challenged cohorts satisfies, for some $\eta>0$,
\[
\E[Z_k\mid \mathbb M^+,\mathcal G_{k-1},I_k=1]\ge \eta,
\qquad
\E[Z_k\mid \mathbb M^-,\mathcal G_{k-1},I_k=1]\le -\eta,
\]
where $\mathcal G_{k-1}$ is the sigma-field generated by all previously matured challenged cohorts. Assume further that the centered challenged statistics form a conditionally sub-Gaussian martingale-difference sequence on the challenged subsequence, i.e.
\[
\E\!\left[\exp\!\bigl(\lambda(Z_k-\E[Z_k\mid \mathcal G_{k-1},I_k=1])\bigr)\middle|\mathcal G_{k-1},I_k=1\right]
\le e^{\lambda^2\varsigma^2/2}
\]
for all $\lambda\in\R$. Then after $N$ matured challenged evaluator-edit cohorts, the average-sign test has error probability at most
\[
2\exp\!\left(-\frac{N\eta^2}{2\varsigma^2}\right).
\]
Consequently, along any infinite evaluator-editing subsequence, the number of matured challenged cohorts diverges almost surely. Moreover, if the average-sign test is applied after each matured challenged cohort count $N=1,2,\dots$, then with probability one only finitely many of those tests err, so the sign ambiguity certified by the statistic class $Z_k$ is eventually broken.
\end{proposition}

\begin{proof}
Condition on the matured challenged subsequence. Under either world, the empirical mean of $Z_k$ has conditional mean separated from zero by at least $\eta$ in magnitude. The centered challenged statistics obey a martingale sub-Gaussian mgf bound, so the usual martingale exponential-tail argument applies to the challenged-subsequence average and yields the displayed error probability; one may view this as the bounded-increment sub-Gaussian analogue of standard martingale concentration tools \cite{Freedman1975}. Bounded delay ensures that every challenged cohort matures in finite time. The assumed conditional Borel--Cantelli property implies that any infinite evaluator-editing subsequence yields infinitely many challenged cohorts almost surely; bounded delay then upgrades this to infinitely many matured challenged cohorts.

For $N\ge 1$, let $E_N$ be the event that the average-sign test after the $N$th matured challenged cohort returns the wrong sign. The displayed tail bound gives
\[
\Pr(E_N)\le 2\exp\!\left(-\frac{N\eta^2}{2\varsigma^2}\right),
\]
so $\sum_{N\ge 1}\Pr(E_N)<\infty$. By the first Borel--Cantelli lemma, only finitely many of the events $E_N$ occur almost surely. Thus there exists an almost-sure finite random index $N_0$ after which every average-sign test returns the correct sign, which is exactly the claimed eventual disambiguation. The filtration bookkeeping is the standard one for discrete-time martingale arguments \cite{Williams1991}.
\end{proof}

\begin{remark}[Robustified challenged statistics beyond sub-Gaussian tails]
The sub-Gaussian martingale assumption in Proposition~\ref{prop:shadow_ident} is a clean sufficient condition, not the only viable route. In deployments with heavy-tailed challenged statistics one may replace the raw average-sign test by replayably clipped, median-of-means, Catoni-type, or self-normalized challenged aggregators and recover the same qualitative conclusion under weaker conditional moment assumptions, at the price of worse constants and additional public tuning parameters. The stability layer therefore depends on the existence of a published robust challenger statistic, not on Gaussian folklore.
\end{remark}
\subsection{Contradiction governance}
\begin{remark}[Why a bounded-depth contradiction window is needed]\label{rem:bounded_depth_need}
Contradiction preservation without a bounded-depth window is unstable in both directions: if all contradiction-bearing witness families are preserved, then resource consumption can diverge; if none are preserved, then the system loses a protected source of evaluative novelty. Depth below $c_{\min}$ corresponds to cheap, easily generated contradictions that need not carry substantive novelty. Depth above $c_{\max}(B_t)$ is not actionable within the current budget and therefore behaves operationally like an unresolved computational sink. Restricting preservation to the interval between these extremes keeps contradiction preservation finite while preventing total contradiction erasure.
\end{remark}

\begin{theorem}[Heterodoxy-preserving improvement]\label{thm:heterodoxy}
Consider a recursively self-improving system that maintains the typed-lane architecture of Definition~\ref{def:lanes}. If the contradiction reserve admits a positive lower policy floor, contradiction-bearing witnesses inside the bounded-depth window carry persistent contradiction obligations in the sense of Definition~\ref{def:contra_obligation}, every policy transition that would otherwise render a preserved contradiction family ineligible must either emit an explicit counted deletion event with conservation compensation or replayably transport that family to a successor contradiction family with the same challenge obligations, and for every preserved contradiction family whose survival trajectory contains infinitely many epochs at which the family remains preserved and challenge-eligible, the challenge-routing process on that infinite eligible subsequence satisfies a conditional Borel--Cantelli property with uniformly positive routing floor and uniformly bounded matured challenge delay, then every such preserved contradiction family inside the bounded-depth window is challenged infinitely often along its infinite eligible survival subsequence under future yardstick states. In particular, bounded disagreement cannot be reduced to one-shot archival theater by immediate deletion, silent relabeling, de-tagging by evaluator whim, or permanent review bypass.
\end{theorem}

\begin{proof}
Dogmatic auto-capture occurs when every bounded contradiction is removed before any slower evaluator can inspect it. The typed-lane policy prevents that exact event by preserving eligible contradiction in the heterodoxy lane, while the anti-gaming transition clause prevents the policy from making disagreement disappear merely by relabeling it out of eligibility without an explicit counted deletion. The conditional Borel--Cantelli property upgrades repeated positive routing opportunity on any infinite eligible survival subsequence into almost-sure infinitely many routed challenges, and the uniform delay bound upgrades infinitely many routed challenges into infinitely many matured challenges. Hence any contradiction family that remains preserved and challenge-eligible infinitely often also remains repeatedly inspectable by future evaluators, so the shortest path to self-sealing dogma is blocked.
\end{proof}


\begin{theorem}[Epistemic routing invariant breaks attentional bypass]\label{thm:forced_attention}
If the typed-lane architecture preserves contradiction-bearing items in storage but imposes no lower bound on heterodoxy routing load $\mathfrak A_t^{\Het}$, then there exist policies that satisfy the storage-side clauses of Theorem~\ref{thm:heterodoxy} while rendering contradiction operationally inert by setting all relevant retrieval gates to zero. In that case, bounded contradiction survives only as archival debris and no longer constrains current admissibility.

Conversely, if the forced-attention invariant of Definition~\ref{def:forced_attention} holds, then any increase in robust contradiction reserve $\bar{\mathcal V}_t^{\mathrm{contra}}$ induces a proportional lower bound on live evaluation sensitivity. Hence the contradiction reserve cannot be both preserved and made collectively decision-irrelevant: total silent dogmatism by routing around the entire reserve is ruled out.
\end{theorem}

\begin{proof}
For the first statement, choose any policy that leaves the stored heterodoxy lane untouched but sets $\alpha_x=0$ for every contradiction-tagged item $x$ during current evaluation. Then $R_t^{\mathrm{contra}}$ and storage-side preservation remain unchanged while every finite-difference sensitivity vanishes, so $\mathfrak A_t^{\Het}=0$. Thus the contradiction reserve no longer influences admissibility despite remaining on disk.

For the converse, the forced-attention invariant lower-bounds $\mathfrak A_t^{\Het}$ by a positive multiple of robust semantic volume. Therefore whenever contradiction reserve has nonzero robust volume, at least some contradiction-tagged items must continue to change the admissibility functional under retrieval ablation. The evaluation pipeline cannot silently route around the entire reserve without violating the invariant.
\end{proof}

\begin{theorem}[Protected random reserve exposure prevents familywise starvation under finite budget]\label{thm:familywise_exposure}
Assume, in addition to the forced-attention invariant, that policy exposes only a finite replayably published \emph{de-duplicated} set of contradiction families at each epoch, chosen for example as the top-$K_t$ or above-threshold preserved-and-eligible families under a public weight system $w_f(t)\ge 0$, and allocates a protected random exposure budget among those active families according to normalized lottery probabilities
\[
\rho_f(t)=\rho_0\frac{w_f(t)}{\sum_{g\in\mathcal F_t^{\mathrm{act}}} w_g(t)},\qquad 0<\rho_0\le 1.
\]
Assume further that before the lottery the active-set selector applies a replayably published anti-dilution contract: near-duplicate families under the declared semantic-similarity threshold are merged or down-weighted; one-source-many-family proliferation pays a published penalty; every active family must exceed a minimum post-merge semantic-volume threshold; the resulting active-set cardinality is finite and block-frozen over declared review windows; changes in $K_t$ or threshold levels obey a replayably declared bounded-variation rule; and a preserved eligible family cannot be dropped from the de-duplicated active set except through a counted blocking, counted deletion, or compensated archival-downgrade certificate. Then every contradiction family $f$ whose eligible survival subsequence is infinite and whose \emph{post-merge} normalized weight share is bounded below by some $\underline\omega_f>0$ on that subsequence is exposed infinitely often almost surely. Combined with a replayably published challenge-routing rule that couples protected exposure to bounded matured challenge service, such a family is also challenged infinitely often. In particular, familywise starvation is blocked for every family that remains in the replayably declared active top-$K_t$ or above-threshold set with persistent positive post-merge share under the protected selector.
\end{theorem}

\begin{proof}
At each epoch the protected random exposure mechanism chooses only among the finite de-duplicated active family set $\mathcal F_t^{\mathrm{act}}$, so the total budget is respected by construction. Because the active-set selector, anti-dilution contract, weight system, and $K_t$/threshold updates are replayably fixed over the review window, the policy cannot adversarially split one semantic family into many lottery tickets or remove a preserved eligible family from the active set without leaving a counted certificate. On the eligible survival subsequence of a family $f$ with persistent post-merge weight share $w_f(t)/\sum_g w_g(t)\ge \underline\omega_f$, the normalized lottery therefore gives conditional inclusion probability at least $\rho_0\underline\omega_f>0$. This yields a conditional Borel--Cantelli property on that subsequence, hence infinitely many protected random exposures almost surely. Bounded matured challenge delay then upgrades infinitely many exposures into infinitely many matured challenges whenever challenge routing is coupled to protected exposure.
\end{proof}

\begin{theorem}[Dynamic contradiction reserve underprovision in the asymmetric case]\label{thm:contradiction_dynamic}
Let $n\ge 2$ agents repeatedly choose contradiction-preservation efforts $a_{i,t}\in[0,\bar a_i]$ with common discount factor $\delta_c\in(0,1)$. Let the contradiction reserve stock evolve as
\[
R_{t+1}^{\mathrm{contra}} = \rho_c R_t^{\mathrm{contra}} + \sum_{i=1}^n \varphi_i a_{i,t},
\qquad
\rho_c\in[0,1),\ \varphi_i>0.
\]
Agent $i$ receives stage payoff
\[
\pi_i\bigl(R_t^{\mathrm{contra}},a_{i,t}\bigr)=\theta_i R_t^{\mathrm{contra}} - \frac{c_i}{2}a_{i,t}^2,
\qquad \theta_i,c_i>0.
\]
Then there exists a stationary Markov-perfect equilibrium with constant contributions
\[
a_i^{\mathrm{NE}}=\Pi_{[0,\bar a_i]}\!\left(\frac{\delta_c\varphi_i\theta_i}{c_i(1-\delta_c \rho_c)}\right),
\]
where $\Pi_{[0,\bar a_i]}(x)\Defeq \min\{\bar a_i,\max\{0,x\}\}$ is interval projection. For the utilitarian social planner maximizing $\sum_{i=1}^n \pi_i$, the stationary optimal contributions are
\[
a_i^{\mathrm{SO}}=\Pi_{[0,\bar a_i]}\!\left(\frac{\delta_c\varphi_i\sum_{j=1}^n\theta_j}{c_i(1-\delta_c \rho_c)}\right).
\]
Hence
\[
a_i^{\mathrm{NE}}\le a_i^{\mathrm{SO}}\qquad\text{for every }i,
\]
with strict inequality whenever agent $i$ is interior and at least one $\theta_j>0$ with $j\neq i$.
\end{theorem}

\begin{proof}
Fix the actions of all agents except $i$ and write the value function of agent $i$ as $V_i(R)=A_i+B_iR$. The Bellman equation is
\[
V_i(R)=\max_{a_i\in[0,\bar a_i]}\left\{\theta_i R-\frac{c_i}{2}a_i^2+\delta_c V_i\!\left(\rho_c R+\varphi_i a_i+\sum_{j\neq i}\varphi_j a_j\right)\right\}.
\]
Matching coefficients of $R$ yields
\[
B_i=\theta_i+\delta_c\rho_c B_i,
\qquad
B_i=\frac{\theta_i}{1-\delta_c\rho_c}.
\]
The first-order condition for an interior optimum is
\[
-c_i a_i+\delta_c \varphi_i B_i=0,
\]
which gives
\[
a_i^{\mathrm{NE}}=\frac{\delta_c\varphi_i\theta_i}{c_i(1-\delta_c\rho_c)}.
\]
Projection onto $[0,\bar a_i]$ gives the displayed equilibrium formula.

For the utilitarian planner, let $W(R)=\widetilde A+\widetilde B R$. The Bellman equation is
\[
W(R)=\max_{a_1,\dots,a_n}\left\{\Bigl(\sum_{j=1}^n\theta_j\Bigr)R-\sum_{j=1}^n\frac{c_j}{2}a_j^2+\delta_c W\!\left(\rho_c R+\sum_{j=1}^n\varphi_j a_j\right)\right\}.
\]
Hence
\[
\widetilde B=\frac{\sum_{j=1}^n\theta_j}{1-\delta_c\rho_c},
\]
and the planner's first-order condition for agent $i$ is
\[
-c_i a_i+\delta_c\varphi_i\widetilde B=0,
\]
which yields
\[
a_i^{\mathrm{SO}}=\frac{\delta_c\varphi_i\sum_{j=1}^n\theta_j}{c_i(1-\delta_c\rho_c)}.
\]
The inequality is immediate because $\sum_j\theta_j\ge \theta_i$, with strict inequality when at least one other agent values the reserve positively and the interior formulas apply.
\end{proof}

\begin{corollary}[Asymmetric contradiction tragedy]\label{cor:asym_contradiction}
Contradiction reserve is underprovided not only in symmetric societies but also when agents differ in cost, productivity, and valuation. In particular, high-productivity or low-cost agents still underinvest whenever part of the future benefit of their preservation effort accrues to other agents.
\end{corollary}

\begin{remark}[Reduced-form role of the contradiction game]
The contradiction-reserve game is a reduced-form abstraction of the stability layer. It is not a full theory of semantics or commons governance. Its purpose is narrower: to show that bounded contradiction, even when socially useful for later recalibration, need not be privately supplied at the level required for loop stability.
\end{remark}

\section{Typed Memory Lanes and Promotion Rules}
\begin{proposition}[Typed lane invariants block direct semantic cross-contamination]\label{prop:typed_lanes}
Suppose a recursive self-improvement system satisfies the following lane invariants for all $t$:
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item only the evidentiary core may directly support current admissibility;
  \item bounded-depth contradiction is preserved in the heterodoxy lane and not immediately discarded;
  \item compressed summaries lacking context certificates are confined to the narrative lane;
  \item promotion from heterodoxy to core obeys Proposition~\ref{prop:promotion};
  \item demotion from core to narrative retires the item from direct evidentiary support.
\end{enumerate}
Then the system avoids three \emph{direct} collapse modes: uncertified narrative contamination of evidence, immediate dogmatic contradiction erasure, and silent promotion of uncertified summaries into admissibility.
\end{proposition}

\begin{proof}
Each collapse mode requires a direct type violation. Narrative contamination violates (i) and (iii). Immediate contradiction erasure violates (ii). Silent uncertified promotion violates (iv). Hence any run satisfying the invariants excludes these direct failure modes.
\end{proof}

\begin{remark}[Scope of the typed-lane theorem]
The theorem blocks direct cross-contamination only. Indirect failures through parser drift, reinterpretation, or bad summarization pipelines still require separate controls. The point of the lane theorem is not to solve semantics in full, but to remove the shortest and cheapest failure paths.
\end{remark}

\begin{proposition}[Lane promotion rule with tagged reserve accounting]\label{prop:promotion}
Let $x\in\mathcal L_t^{\Het}$. A sufficient condition for promotion to $\mathcal L_t^{\Core}$ is that
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item $d_{\min}\le d_{\mathrm{prov}}(x)\le d_{\max}$,
  \item $x$ carries a replayable contradiction certificate and remains contradiction-tagged after promotion whenever it still lies inside the admissible contradiction window,
  \item adding $x$ does not reduce challenge-lane safety margin $S_t^{\mathrm{chal}}$ below policy floor,
  \item parser-family agreement $A_t^{\mathrm{parse}}(x)$ and replay success exceed policy threshold.
\end{enumerate}
\end{proposition}

\begin{proof}
Conditions (i) and (iv) guarantee operational tractability and public interpretability. Condition (ii) prevents promotion from laundering contradiction mass out of the reserve accounting. Condition (iii) ensures that promotion does not damage the challenge process that would later check it.
\end{proof}

\begin{definition}[Promotion-eligible delayed cohort, selector invariance, and multi-action deployment policy]\label{def:promotion_eligible}
A challenged or proof-escalated proposal cohort $u_\tau$ is \emph{promotion-eligible} at review time $s\ge \tau$ if all of the following hold on the then-active public state:
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item a matured delayed certificate is available by time $s$ --- either a matured challenge certificate or a richer-proof certificate --- with calibrated lower stable-gain bound nonnegative on the declared deployment horizon;
  \item issuance-time and peak debt certificates, including any structural blast-radius bond and bundle interaction debt required by policy, remain below threshold on that horizon;
  \item base admissibility and governance-safety clauses required for deployment into the target lane hold at time $s$, including replay, parser, contradiction-tag, reserve-preservation, and any active proof-interface clauses required for that lane.
\end{enumerate}
We expose the public deployment actions
\[
a_s^{\mathrm{deploy}}\in\{\textsf{fast-accept},\textsf{defer-challenge},\textsf{escalate-proof},\textsf{sandbox},\textsf{promote},\textsf{reject}\}.
\]
A public deployment selector is \emph{eligibility-respecting} if a promotion-eligible cohort may be assigned \textsf{promote}, \textsf{sandbox}, or \textsf{escalate-proof}, but may be assigned \textsf{reject} only together with a public failure certificate showing that one of the eligibility clauses no longer holds. It is \emph{review-locked} with constants $W_{\mathrm{sel}},W_{\mathrm{audit}}<\infty$ if, once a cohort first becomes promotion-eligible at review time $s$, the target-lane thresholds and eligibility-relevant manifest clauses used to assess that cohort are frozen for the next $W_{\mathrm{sel}}$ review epochs, except that a failure certificate may be emitted under that frozen review contract and must then be audit-resolved within $W_{\mathrm{audit}}$ review epochs. It is \emph{bounded-wait} if any cohort that remains promotion-eligible under its frozen review contract and receives no validated failure certificate during that window is assigned action \textsf{promote} by the end of the window. It is \emph{exposure-bounded} with constants $W_{\mathrm{rev}},W_{\mathrm{exp}}<\infty$ if any cohort that acquires a matured delayed certificate enters review within $W_{\mathrm{rev}}$ epochs and, under the then-declared target contract, either becomes promotion-eligible or emits a public blocking or failure certificate within an additional $W_{\mathrm{exp}}$ epochs. A selector satisfying the eligibility-respecting, review-locked, and exposure-bounded clauses is called \emph{selector-invariant}: it cannot silently erase a positively certified cohort by delaying review, moving the target during review, or refusing a review-phase verdict without a public certificate. The bounded-wait clause is a separate completion property that upgrades selector invariance from review integrity to bounded-delay promotion once eligibility persists under the frozen review contract.
\end{definition}

\begin{theorem}[Selector-invariant bounded-delay deployment of positively certified delayed cohorts]\label{thm:promotion_liveness}
Let $q_t^{\mathrm{prom}}$ be the backlog of cohorts already assigned action \textsf{promote} and awaiting promotion from a protected challenge, heterodoxy, or richer-proof lane into an admissible deployment lane. Assume a work-conserving starvation-free scheduler in cohort-equivalent workload units, a public promotion service floor $\mu_t^{\mathrm{prom}}\ge \underline\mu_{\mathrm{prom}}>0$, an arrival ceiling $\Lambda_t^{\mathrm{prom}}\le \bar\Lambda_{\mathrm{prom}}<\underline\mu_{\mathrm{prom}}$, and a cap-maintaining intake-control contract from Definition~\ref{def:cap_maintain} with public cap $\bar q^{\mathrm{prom}}$, initialized with $q_0^{\mathrm{prom}}\le \bar q^{\mathrm{prom}}$. Assume also that every cohort routed to challenge satisfies Proposition~\ref{prop:challenge_positivity}, that promotion uses the tag-preserving sufficient conditions of Proposition~\ref{prop:promotion}, that reserve accounting obeys Proposition~\ref{prop:reserve_conservation}, and that the deployment selector is selector-invariant and bounded-wait with constants $W_{\mathrm{rev}},W_{\mathrm{exp}},W_{\mathrm{sel}},W_{\mathrm{audit}}$. By Proposition~\ref{prop:cap_invariance}, the promotion backlog cap is then a maintained invariant rather than a free parameter. Then any cohort issued at time $\tau$ that acquires a matured delayed certificate by time $m$ satisfies one of the following alternatives:
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item by time $m+W_{\mathrm{rev}}+W_{\mathrm{exp}}+W_{\mathrm{audit}}$ a validated public blocking or failure certificate is resolved, showing either that the cohort cannot satisfy the target-lane deployment clauses or that a promotion-eligibility clause fails under the frozen review contract; or
  \item by time $m+W_{\mathrm{rev}}+W_{\mathrm{exp}}+W_{\mathrm{audit}}+W_{\mathrm{sel}}$ the cohort is assigned action \textsf{promote} and is then promoted no later than
\[
 m+W_{\mathrm{rev}}+W_{\mathrm{exp}}+W_{\mathrm{audit}}+W_{\mathrm{sel}}+\bar L_{\mathrm{prom}},\qquad\bar L_{\mathrm{prom}}\Defeq\left\lceil \frac{\bar q^{\mathrm{prom}}+1}{\underline\mu_{\mathrm{prom}}-\bar\Lambda_{\mathrm{prom}}} \right\rceil.
\]
\end{enumerate}
If challenge maturity itself is available within the bound $\bar L$ of Proposition~\ref{prop:challenge_positivity}, then any challenged cohort whose target contract is fixed at maturity is either invalidated by an audited certificate or deployed within at most
\[
\bar L+W_{\mathrm{rev}}+W_{\mathrm{exp}}+W_{\mathrm{audit}}+W_{\mathrm{sel}}+\bar L_{\mathrm{prom}}
\]
steps after issuance. Therefore delayed certification need not be a graveyard for good edits: under explicit review, exposure, audit, selection, and service floors it becomes a bounded-delay deployment path from the first positive delayed certificate onward.
\end{theorem}

\begin{proof}
Once a cohort acquires a matured delayed certificate at time $m$, exposure-bounded review forces it into review within $W_{\mathrm{rev}}$ epochs and then requires either a public blocking or failure certificate or a promotion-eligibility verdict within another $W_{\mathrm{exp}}$ epochs. If a blocking or failure certificate is emitted, audited resolution within $W_{\mathrm{audit}}$ epochs gives alternative (i). Otherwise the cohort becomes promotion-eligible. Review-locking then freezes the relevant target-lane clauses for $W_{\mathrm{sel}}$ review epochs, so the selector cannot make the cohort fail merely by moving thresholds or silently retyping the target contract. If no validated failure certificate appears during the audit window, bounded-wait selection forces action \textsf{promote} by the end of the frozen review window. Once assigned, the cohort joins the promotion queue with at most $\bar q^{\mathrm{prom}}$ workload units ahead of it. The service floor and arrival ceiling imply a uniform net drain of at least $\underline\mu_{\mathrm{prom}}-\bar\Lambda_{\mathrm{prom}}>0$ whenever the promotion queue is nonempty. Under work-conserving starvation-free service in cohort-equivalent units, the displayed queueing bound follows. Proposition~\ref{prop:promotion} and Proposition~\ref{prop:reserve_conservation} ensure that promotion preserves the contradiction-tag and reserve-side invariants required for the move to remain admissible.
\end{proof}

\begin{corollary}[Fast-lane impossibility becomes governed feasibility with delayed certification and selector invariance]\label{cor:governed_feasibility}
Consider any evaluator-edit proposal class for which Theorem~\ref{thm:ruler_nogo} rules out uniform sign identification on the fast public lane. Suppose such proposals are routed through Corollary~\ref{cor:end_to_end_challenge}, their matured challenge evidence yields a calibrated nonnegative stable-gain certificate via Proposition~\ref{prop:shadow_bridge} or Corollary~\ref{cor:shadow_cert}, and the resulting cohorts satisfy debt safety, governance safety, and the deployment clauses of Definition~\ref{def:promotion_eligible}. Assume moreover that, once maturity occurs, the downstream promotion lane and review process satisfy all premises of Theorem~\ref{thm:promotion_liveness}, including protected promotion service, the published backlog cap, work-conserving starvation-free scheduling, tag-preserving promotion, reserve preservation, selector invariance, and bounded-wait completion. Then those proposals remain \emph{governedly deployable}: they may be undecidable on the fast lane yet still become admissibly deployable after bounded delay unless an audited blocking or failure certificate establishes genuine ineligibility. Claim-only governance is therefore impossible in this class, but governed recursive improvement remains feasible.
\end{corollary}

\begin{proof}
Theorem~\ref{thm:ruler_nogo} blocks only fast-lane sign identification. Corollary~\ref{cor:end_to_end_challenge} guarantees that challenge-eligible cohorts are either publicly blocked within bounded time or enter and mature on the protected challenge lane within bounded end-to-end delay, the shadow-certification bridge supplies a delayed lower stable-gain certificate, and Definition~\ref{def:promotion_eligible} packages the remaining debt and governance requirements for deployment. The additional assumptions place the matured cohort exactly inside the scope of Theorem~\ref{thm:promotion_liveness}. That theorem then turns delayed certification plus protected service into bounded-delay deployment unless an audited blocking or failure certificate is validated. Hence impossibility of immediate claim-only admission does not imply impossibility of governed deployment.
\end{proof}

\begin{definition}[Proof-carrying escalation contract]\label{def:proof_escalation}
Fix a richer replayable proof interface $\mathfrak L^{\mathrm{rich}}$. A \emph{proof-carrying escalation contract} is replayable if it specifies a finite offer deadline $W_{\mathrm{offer}}<\infty$, a finite retry cap $N_{\mathrm{retry}}<\infty$, and public success/failure packet formats such that for every escalated cohort: (i) within each offer window of length $W_{\mathrm{offer}}$, the cohort either submits one richer-proof candidate or emits a terminal escalation-failure certificate; (ii) each submitted candidate is checked by a total verifier from the declared richer interface under thermodynamic proof escrow; (iii) every successful check emits a richer-proof certificate; (iv) every failed check emits a replayable verification-failure certificate; and (v) after at most $N_{\mathrm{retry}}$ unsuccessful checked candidates without success, the cohort must emit a terminal escalation-failure certificate rather than retry indefinitely.
\end{definition}

\begin{theorem}[Proof-carrying bounded-delay deployability of proof-heavy upgrades under richer proof lanes]\label{thm:proof_liveness}
Let a proposal be \emph{proof-heavy} if its safety argument lies beyond the current constitutional proof class $\mathfrak L^{\mathrm{const}}$ but inside a richer replayable proof interface $\mathfrak L^{\mathrm{rich}}$ accepted by policy for escalation. Assume that each richer-interface checker is total and physically metered under Definition~\ref{def:thermo_escrow}, so that by Theorems~\ref{thm:godelian_ceiling} and \ref{thm:thermo_escrow} every submitted richer-proof candidate has a replayable accept/reject outcome within a public bounded physical verification cost. Assume a proof-carrying escalation contract in the sense of Definition~\ref{def:proof_escalation} with constants $W_{\mathrm{offer}}$ and $N_{\mathrm{retry}}$. Let the protected proof lane have backlog $q_t^{\mathrm{proof}}$ governed in cohort-equivalent verification workload units by a work-conserving starvation-free scheduler and queue recursion
\[
q_{t+1}^{\mathrm{proof}}\le \bigl[q_t^{\mathrm{proof}}-\mu_t^{\mathrm{proof}}\bigr]_+ + A_t^{\mathrm{proof}},
\qquad
A_t^{\mathrm{proof}}\le \Lambda_t^{\mathrm{proof}},
\]
with service floor $\mu_t^{\mathrm{proof}}\ge \underline\mu_{\mathrm{proof}}>0$, arrival ceiling $\Lambda_t^{\mathrm{proof}}\le \bar\Lambda_{\mathrm{proof}}<\underline\mu_{\mathrm{proof}}$, and a cap-maintaining intake-control contract from Definition~\ref{def:cap_maintain} with public cap $\bar q^{\mathrm{proof}}$, initialized with $q_0^{\mathrm{proof}}\le \bar q^{\mathrm{proof}}$. Assume further that once a richer-proof certificate is obtained, the downstream deployment lane satisfies all premises of Theorem~\ref{thm:promotion_liveness}. Define
\[
\bar L_{\mathrm{proof}}\Defeq\left\lceil \frac{\bar q^{\mathrm{proof}}+1}{\underline\mu_{\mathrm{proof}}-\bar\Lambda_{\mathrm{proof}}} \right\rceil.
\]
Then any proof-heavy cohort escalated at time $\tau$ satisfies one of the following alternatives:
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item by time $\tau+N_{\mathrm{retry}}(W_{\mathrm{offer}}+\bar L_{\mathrm{proof}})$ it emits a terminal public escalation-failure or verification-failure certificate;
  \item by time $\tau+N_{\mathrm{retry}}(W_{\mathrm{offer}}+\bar L_{\mathrm{proof}})$ it obtains a richer-proof certificate discharging the deployment clauses of Definition~\ref{def:promotion_eligible}, enters the downstream review process of Theorem~\ref{thm:promotion_liveness}, and is therefore deployed or invalidated within at most
\[
N_{\mathrm{retry}}(W_{\mathrm{offer}}+\bar L_{\mathrm{proof}})+W_{\mathrm{rev}}+W_{\mathrm{exp}}+W_{\mathrm{audit}}+W_{\mathrm{sel}}+\bar L_{\mathrm{prom}}
\]
steps after escalation.
\end{enumerate}
Hence richer proof lanes need not be permanent freezing chambers: under proof-carrying escalation, total replayable checkers, thermodynamic proof escrow, protected proof-lane service, and finite retry budgets, proposals beyond the current fast proof regime become bounded-delay deployable or bounded-delay falsifiable. This lane design is consistent with observable-only hybrid proof/replay gating patterns, but the present theorem is a stability-layer statement about delayed deployment rather than a packet-level protocol claim.
\end{theorem}

\begin{proof}
By Definition~\ref{def:proof_escalation}, each offer window of length $W_{\mathrm{offer}}$ ends either with submission of one candidate or with a terminal public escalation-failure certificate. Every submitted candidate enters the protected proof queue. The queue recursion, service floor, arrival ceiling, cap-maintaining intake rule, and starvation-free scheduling imply via Proposition~\ref{prop:cap_invariance} that the proof backlog never exceeds $\bar q^{\mathrm{proof}}$, so any submitted candidate is checked within at most $\bar L_{\mathrm{proof}}$ additional epochs. Because each checked candidate yields either a replayable richer-proof certificate or a replayable verification-failure certificate, and because indefinite retry is ruled out after at most $N_{\mathrm{retry}}$ unsuccessful checked candidates, the escalated cohort is resolved on the proof lane within at most $N_{\mathrm{retry}}(W_{\mathrm{offer}}+\bar L_{\mathrm{proof}})$ epochs. If resolution is failure, alternative (i) holds. If resolution is success, the cohort now satisfies the richer-proof part of the deployment contract and all premises of Theorem~\ref{thm:promotion_liveness} apply by assumption to the downstream lane, yielding the additional bounded review, audit, selection, and promotion delay in alternative (ii).
\end{proof}

\begin{definition}[Admissible service-scaling law]\label{def:service_scaling}
A published governance allocator is an \emph{admissible service-scaling law} if there exist per-lane protected resource counters $P_t^{\mathrm{chal}},P_t^{\mathrm{prom}},P_t^{\mathrm{proof}}$, per-cohort physical cost envelopes $c_t^{\mathrm{chal}},c_t^{\mathrm{prom}},c_t^{\mathrm{proof}}$, replayably published service floors $\underline\mu_t^{\mathrm{chal}},\underline\mu_t^{\mathrm{prom}},\underline\mu_t^{\mathrm{proof}}$, positive slack constants $\eps_{\mathrm{chal}},\eps_{\mathrm{prom}},\eps_{\mathrm{proof}}>0$, public finite caps $\bar q^{\mathrm{chal}},\bar q^{\mathrm{prom}},\bar q^{\mathrm{proof}}$, cap-maintaining intake-control contracts in the sense of Definition~\ref{def:cap_maintain}, and bounded-exposure entry controls for any upstream eligible holding states in the sense of Definition~\ref{def:bounded_exposure_intake} such that along the realized trajectory
\[
\underline\mu_t^{\ell}\le \frac{P_t^{\ell}}{c_t^{\ell}},\qquad \underline\mu_t^{\ell}-\Lambda_t^{\ell}\ge \eps_{\ell}\qquad (\ell\in\{\mathrm{chal},\mathrm{prom},\mathrm{proof}\}),
\]
with each lane initialized below its public cap. In words: governance service scaling is not treated as an exogenous miracle but as a resource-preserving allocation rule that converts protected budget and conservative cost envelopes into public service floors with positive lane slack while jointly maintaining finite queue caps through replayable intake control and preventing eligible cohorts from disappearing indefinitely into upstream holding.
\end{definition}

\begin{corollary}[Resource-coupled service scaling preserves lane-service premises and hence governed deployability]\label{cor:service_scaling}
Suppose challenge, promotion, and richer-proof lanes satisfy Definition~\ref{def:service_scaling} along the realized trajectory. Assume additionally that the non-service premises of Proposition~\ref{prop:challenge_positivity}, Theorems~\ref{thm:promotion_liveness} and \ref{thm:proof_liveness}, and Corollary~\ref{cor:governed_feasibility} remain valid along that trajectory, including the routing-propensity floor, starvation-free scheduling, selector-invariant and bounded-wait review contract, proof-carrying escalation contract, total-verifier clauses, reserve-preservation clauses, and cap-maintaining intake-control contracts. Then the service-rate, queue-slack, and finite-cap premises entering those results continue to hold uniformly along the realized trajectory by Definition~\ref{def:service_scaling} and Proposition~\ref{prop:cap_invariance}. Consequently the cited bounded-delay maturity, promotion, proof-lane verification, and governed-deployability conclusions remain valid as the protected budgets scale with load. In particular, success-driven proposal throughput need not collapse the governance lanes provided service is coupled to publicly reserved resource budgets and the other structural premises are maintained.
\end{corollary}

\begin{proof}
Definition~\ref{def:service_scaling} turns protected budget and conservative per-cohort cost envelopes into explicit public service floors and positive slacks for the challenge, promotion, and proof lanes. The remaining hypotheses listed in the statement ensure that the non-service parts of the cited results are not lost when the allocator changes over time. Therefore every lane-side service premise required by Proposition~\ref{prop:challenge_positivity}, Theorems~\ref{thm:promotion_liveness} and \ref{thm:proof_liveness}, and Corollary~\ref{cor:governed_feasibility} remains satisfied uniformly along the realized trajectory, so the same bounded-delay and queue-stability arguments continue to apply.
\end{proof}

\begin{remark}[Endogenous maintenance is external to the service law]
Corollary~\ref{cor:service_scaling} is conditional on the protected-budget allocator and intake-control contract continuing to satisfy their published floors and caps. It therefore shows that bounded-delay deployability is policy-contract compatible with success; it does not prove that an unconstrained recursive improver will endogenously preserve those contracts without a separate constitutional or resource-allocation layer.
\end{remark}

\begin{definition}[Observable witness-comparison contract for raw realized-path dominance]\label{def:gp_class}
A pair of closed-loop regimes $(\mathsf R^-,\mathsf R^+)$ driven by a shared replayable exogenous skeleton satisfies the \emph{observable witness-comparison contract for raw realized-path dominance} if, at every epoch, all of the following hold under a common published manifest and fixed acceptance thresholds:
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item the common sanitized comparison menu $\mathcal U_t^{\mathrm{cmp}}$ is replayably published and stable under the common syntax/scope/null-certificate prefilters used by both regimes, so menu engineering alone cannot create governance-capital dominance;
  \item productive proposal menus admit a replayable matching map from minus-regime admissible productive proposals to plus-regime admissible productive proposals with weakly larger direct productive increment and no worse delayed-certificate obligations;
  \item accepted matched productive proposals preserve weak dominance of productive stock and fixed-threshold governance capital at the next epoch;
  \item frontier-improving governance edits are closed under plus-regime selection and, whenever chosen against any admissible minus-regime action at the same epoch, preserve weak next-step productive-stock dominance while producing strict next-step fixed-threshold governance-capital dominance;
  \item the admissible governance edits published by the plus-regime include at least one replayable exploitation path from any strict next-step frontier gain to a later productive or delayed-deployment action whenever that enlarged frontier remains available;
  \item challenge-eligible, delayed challenge, and richer-proof cohorts are governed by bounded-exposure entry control, selector invariance, and Theorems~\ref{thm:promotion_liveness} and \ref{thm:proof_liveness};
  \item rollback, sandbox, and bounded-scope deployment actions preserve the weak productive-stock order;
  \item protected challenge, promotion, and proof services satisfy the admissible service-scaling law of Definition~\ref{def:service_scaling}.
\end{enumerate}
This definition isolates the exact explicit comparison contract under which the later realized-path claim is intended to hold. Governance is not assumed to help by rhetoric; the contract is stated so that stronger public interfaces, sanitized-menu comparison, bounded-exposure entry, delayed-certification liveness, exploitation of enlarged frontiers, and resource-coupled service preserve or expand future admissible productive opportunity under unchanged thresholds.
\end{definition}

\begin{remark}[Comparison-contract reading of the positive claim]
Definition~\ref{def:gp_class} is not presented as a universal law that arbitrary self-editing regimes satisfy. It is an explicit transition contract collecting the exact observable premises under which an inductive realized-path comparison is valid. The corollary below is informative only to the extent that the contract can be certified from independently published witnesses; outside that contract the paper makes no claim that stronger governance necessarily improves realized growth.
\end{remark}

\begin{definition}[Action-induced next-step regime and post-action frontier scores]\label{def:post_action_frontier}
Fix an epoch $t$, a current regime $\mathsf R_t$, the shared replay skeleton, and an admissible current-epoch action $u$. Let $\mathsf R_{t+1}[u;\mathsf R_t]$ denote the replayably determined next-epoch regime obtained by applying $u$ under $\mathsf R_t$ together with the manifest-declared post-decision update rules. Define the \emph{post-action frontier scores}
\[
\Phi_{t+1}^{\mathrm{post}}(u;\mathsf R_t)\Defeq \Phi_{t+1}(\mathsf R_{t+1}[u;\mathsf R_t]),
\qquad
\Phi_{t+1}^{\mathrm{net,post}}(u;\mathsf R_t)\Defeq \Phi_{t+1}^{\mathrm{net}}(\mathsf R_{t+1}[u;\mathsf R_t]).
\]
These objects are well-typed one-step look-ahead frontier scores attached to current actions; they are not shorthand for evaluating the current frontier at an action argument.
\end{definition}

\begin{definition}[Replayable comparison-witness selectors]\label{def:witness_selector}
Fix a regime pair in Definition~\ref{def:gp_class}, an epoch $t$, and a replayable admissible minus-regime selector $\pi_t^-$. For the realized minus action $u_t^-=\pi_t^-(C_t)$, let $\mathcal W_t(u_t^-)$ denote a replayably published nonempty set of admissible plus-regime \emph{comparison witnesses}, each of which certifies one of the comparison clauses of Definition~\ref{def:gp_class} against $u_t^-$: a matched productive action, a frontier-improving governance edit together with its next-step frontier certificate, a delayed challenge or richer-proof deployment witness once eligible, or a rollback/sandbox witness preserving weak productive order. A plus-regime selector $\pi_t^+$ is \emph{comparison-witness dominating} relative to $\pi_t^-$ if, for every realized state and minus action $u_t^-$, the chosen admissible plus action $u_t^+=\pi_t^+(C_t,u_t^-)$ satisfies
\[
(\Phi_{t+1}^{\mathrm{post}}(u_t^+;\mathsf R_t^+),\Delta\kappa_t(u_t^+))\succcurlyeq_{\mathrm{lex}}
(\Phi_{t+1}^{\mathrm{post}}(w;\mathsf R_t^+),\Delta\kappa_t(w))
\]
for at least one witness action $w\in\mathcal W_t(u_t^-)$. Exact lexicographic maximization over all admissible plus actions is sufficient but not required.
\end{definition}

\begin{corollary}[Inductive raw realized-path productive dominance under the explicit witness-comparison contract]\label{cor:realized_growth}
Consider two closed-loop regime-policy pairs $(\mathsf R_t^-,\pi_t^-)$ and $(\mathsf R_t^+,\pi_t^+)_{t\ge 0}$ such that $(\mathsf R_t^-,\mathsf R_t^+)$ belongs to the observable growth-promoting sufficient-condition class of Definition~\ref{def:gp_class} at every epoch, $\pi_t^-$ is any replayable admissible selector for the minus-regime, and $\pi_t^+$ is a replayable comparison-witness-dominating selector relative to $\pi_t^-$ in the sense of Definition~\ref{def:witness_selector}. Then, pathwise on the shared replay skeleton,
\[
\Phi_t(\mathsf R_t^+)\ge \Phi_t(\mathsf R_t^-)\qquad\text{and}\qquad\kappa(x_t^+)\ge \kappa(x_t^-)\qquad \forall t\ge 0.
\]
If at some epoch the chosen plus action is a productive proposal whose direct productive increment strictly exceeds that of the replayably compared minus action and the productive-transition clause maps that strict increment into strict next-step productive-state dominance, then $\kappa(x_{t+1}^+)>\kappa(x_{t+1}^-)$. If instead the chosen plus action is a strict frontier-improving governance edit, then
\[
\Phi_{t+1}(\mathsf R_{t+1}^+)>\Phi_{t+1}(\mathsf R_{t+1}^-).
\]
Later productive-state strictness then requires subsequent exploitation of that strict frontier advantage by a later productive or delayed deployment witness. Persistent strictness is not claimed without an additional strictness-preserving transition clause.
\end{corollary}

\begin{proof}
Proceed by induction on $t$. The base case is immediate because the initial public state is shared. Assume the displayed weak inequalities hold at epoch $t$. Let $u_t^-=\pi_t^-(C_t)$ be the realized minus action. By Definition~\ref{def:witness_selector}, the chosen plus action $u_t^+=\pi_t^+(C_t,u_t^-)$ is lexicographically no worse than at least one replayably published witness action guaranteed by Definition~\ref{def:gp_class}. If the witness is a matched productive action, the productive-transition clause propagates weak productive-stock and governance-capital dominance to epoch $t+1$. If the witness is a frontier-improving governance edit, the governance-edit clause yields weak next-step productive-stock dominance relative to $u_t^-$, while Theorem~\ref{thm:gov_investment} gives strict next-step frontier improvement under the common comparison manifest. If the witness is an eligible delayed deployment from the challenge or richer-proof lanes, then bounded-exposure entry control together with Theorems~\ref{thm:promotion_liveness} and \ref{thm:proof_liveness} ensures that positively certified delayed cohorts cannot remain stuck forever: once the relevant delayed certificate exists, entry, review, and protected service force audited invalidation or bounded-delay deployment. If the witness is a rollback or sandbox action, the corresponding clause preserves weak productive order by assumption. In every case the chosen plus action is no worse than some valid witness, so both displayed weak inequalities propagate to epoch $t+1$. The strict claims are exactly the immediate strict consequences stated in the corollary, and no stronger persistence statement is used.
\end{proof}

\begin{definition}[Comparison-sound replayable surrogate selector]\label{def:surrogate_selector}
Fix an epoch $t$, a realized minus action $u_t^-$, and the replayably published witness set $\mathcal W_t(u_t^-)$ from Definition~\ref{def:witness_selector}. A replayably published surrogate score $\Sigma_t(u\,;\,u_t^-)$ for admissible plus-regime actions is \emph{comparison-sound} for the raw lexicographic objective $(\Phi_{t+1}^{\mathrm{post}}(\cdot;\mathsf R_t^+),\Delta\kappa_t)$ if every admissible surrogate maximizer $\hat u_t$ satisfies
\[
(\Phi_{t+1}^{\mathrm{post}}(\hat u_t;\mathsf R_t^+),\Delta\kappa_t(\hat u_t))\succcurlyeq_{\mathrm{lex}}
(\Phi_{t+1}^{\mathrm{post}}(w;\mathsf R_t^+),\Delta\kappa_t(w))
\]
for at least one witness action $w\in\mathcal W_t(u_t^-)$ on every trajectory in the comparison class. The surrogate is \emph{complete} if a surrogate maximizer exists whenever the witness set is nonempty. Typical surrogates are replayable lower envelopes or myopic lower bounds that are only required to dominate a certified witness action, not to recover the exact global action ordering over the full admissible menu.
\end{definition}

\begin{theorem}[Comparison-sound replayable surrogate selectors preserve realized-path growth dominance]\label{thm:surrogate_realized_growth}
Under the hypotheses of Corollary~\ref{cor:realized_growth}, the conclusion remains valid if the plus-regime selector $\pi_t^+$ is implemented by maximizing any replayably published surrogate score $\Sigma_t(\cdot\,;\,u_t^-)$ that is complete and comparison-sound in the sense of Definition~\ref{def:surrogate_selector}. In particular, the plus-regime need not solve the exact one-step frontier maximization problem on line; it suffices to maximize a replayable surrogate strong enough to dominate at least one certified comparison witness for the realized minus action.
\end{theorem}

\begin{proof}
At each epoch the witness set is nonempty by Definition~\ref{def:witness_selector}. Completeness yields at least one admissible surrogate maximizer. Comparison-soundness implies that the selected action is lexicographically no worse in the true pair $(\Phi_{t+1}^{\mathrm{post}}(\cdot;\mathsf R_t^+),\Delta\kappa_t)$ than at least one certified witness action for the realized minus action. The proof of Corollary~\ref{cor:realized_growth} uses only this witness-dominating property, not exact maximization over the full admissible menu. Therefore the same inductive argument goes through after replacing direct witness selection by a comparison-sound surrogate selector.
\end{proof}

\begin{definition}[Admissible governance burden accounting and resource-priced frontier]\label{def:net_growth_metric}
A replayably published governance-resource burden functional $B_t^{\mathrm{gov}}\ge 0$ is \emph{admissible} if it satisfies all of the following axioms along the realized trajectory:
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item (Physical lower bound) there exist public constants $a_{\mathrm{cpu}},a_{\mathrm{mem}},a_{\mathrm{wall}},a_{\mathrm{eng}}>0$ such that
  \[
  B_t^{\mathrm{gov}}\ge a_{\mathrm{cpu}}C_t^{\mathrm{cpu}}+a_{\mathrm{mem}}C_t^{\mathrm{mem}}+a_{\mathrm{wall}}C_t^{\mathrm{wall}}+a_{\mathrm{eng}}C_t^{\mathrm{eng}},
  \]
  where the $C_t^{\cdot}$ are replayably logged protected governance compute, memory, wall-clock, and energy usage counters;
  \item (Monotonicity) $B_t^{\mathrm{gov}}$ is monotone in protected challenge, promotion, and proof service commitments, in verification and proof workload, and in any thermodynamic proof-escrow budget consumed by the current governance regime;
  \item (Bounded interaction overhead) there exists a replayably published nonnegative interaction-overhead envelope $\Delta_t^{\mathrm{int}}$ such that combining two governance commitments costs at most the sum of their separately published burdens plus $\Delta_t^{\mathrm{int}}$;
  \item (Null normalization) $B_t^{\mathrm{gov}}=0$ under the null governance load with no protected challenge, promotion, proof, escrow, or verification burden.
\end{enumerate}
For a price parameter $\lambda\ge 0$, define the \emph{resource-priced productive state}
\[
\kappa_t^{\mathrm{net}}\Defeq \kappa(x_t)-\lambda B_t^{\mathrm{gov}},
\]
and for any admissible action $u_t$ define its \emph{resource-priced increment}
\[
\Delta\kappa_t^{\mathrm{net}}(u_t)\Defeq \kappa(x_{t+1}(u_t)) - \kappa(x_t) - \lambda\bigl(B_{t+1}^{\mathrm{gov}}(u_t)-B_t^{\mathrm{gov}}\bigr).
\]
The corresponding \emph{net admissible frontier} is
\[
\Phi_t^{\mathrm{net}}(\mathsf R)\Defeq \sup_{u\in\mathcal A_t^{\mathrm{full}}(\mathsf R)} \Delta\kappa_t^{\mathrm{net}}(u),
\qquad \sup\varnothing\Defeq 0.
\]
\end{definition}

\begin{remark}[Policy relativity of resource-priced growth]
The burden functional $B_t^{\mathrm{gov}}$ and the price parameter $\lambda$ are part of the declared accounting regime. Net-frontier and net-state comparisons are therefore relative to a chosen replayably published lower-bounding price functional; they do not assert a unique architecture-independent welfare order across all possible resource prices.
\end{remark}

\begin{definition}[Observable witness-comparison contract for resource-priced realized-path dominance]\label{def:gp_net_class}
An observable growth-promoting regime pair satisfies the \emph{observable witness-comparison contract for resource-priced realized-path dominance} if, in addition to Definition~\ref{def:gp_class}, it uses an admissible burden functional from Definition~\ref{def:net_growth_metric} and satisfies the following net analogues at every epoch:
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item every productive proposal admissible in $\mathsf R_t^-$ has a matched admissible proposal in $\mathsf R_t^+$ with weakly larger resource-priced increment $\Delta\kappa_t^{\mathrm{net}}$;
  \item any strict net-frontier improvement generated by a plus-regime governance edit remains replayably exploitable by some later admissible productive or delayed-deployment path whenever the net frontier advantage persists;
  \item every frontier-improving governance edit chosen in $\mathsf R_t^+$ weakly increases the next-step net frontier, with strict improvement whenever the certified future frontier expansion outweighs the increase in admissible governance burden under price $\lambda$;
  \item accepting matched productive proposals and successful delayed deployments from challenge or proof lanes preserves weak dominance of the resource-priced productive state at the next epoch.
\end{enumerate}
\end{definition}

\begin{remark}[Comparison-class reading of the net positive claim]
Definition~\ref{def:gp_net_class} adds one further layer beyond Definition~\ref{def:gp_class}: the public accounting regime must certify that governance-resource cost is priced in a way that preserves the stated net-dominance transitions. The resulting corollary is again a comparison under an explicit contract, not a claim that every governance burden is net-promotive under every price functional. The same comparison-sound surrogate-selector idea applies to the net corollary after replacing $(\Phi_{t+1}^{\mathrm{post}}(\cdot;\mathsf R_t^+),\Delta\kappa_t)$ by $(\Phi_{t+1}^{\mathrm{net,post}}(\cdot;\mathsf R_t^+),\Delta\kappa_t^{\mathrm{net}})$.
\end{remark}

\begin{corollary}[Inductive resource-priced realized-path productive dominance under the explicit net witness-comparison contract]\label{cor:net_realized_growth}
Consider two closed-loop regime-policy pairs $(\mathsf R_t^-,\pi_t^-)$ and $(\mathsf R_t^+,\pi_t^+)_{t\ge 0}$ such that $(\mathsf R_t^-,\mathsf R_t^+)$ belongs to the observable net growth-promoting sufficient-condition class of Definition~\ref{def:gp_net_class} at every epoch, $\pi_t^-$ is any replayable admissible minus selector, and $\pi_t^+$ is any replayable plus selector that, for each realized minus action $u_t^-=\pi_t^-(C_t)$, chooses an admissible plus action lexicographically no worse in $(\Phi_{t+1}^{\mathrm{net,post}}(\cdot;\mathsf R_t^+),\Delta\kappa_t^{\mathrm{net}})$ than at least one replayably published net witness action guaranteed by Definition~\ref{def:gp_net_class}. Then, pathwise on the shared replay skeleton,
\[
\Phi_t^{\mathrm{net}}(\mathsf R_t^+)\ge \Phi_t^{\mathrm{net}}(\mathsf R_t^-)\qquad\text{and}\qquad\kappa_t^{\mathrm{net},+}\ge \kappa_t^{\mathrm{net},-}\qquad \forall t\ge 0.
\]
If at some epoch the chosen plus action has strictly larger resource-priced increment than the replayably compared minus action and the net transition clause maps that strict increment into strict next-step net productive-state dominance, then $\kappa_{t+1}^{\mathrm{net},+}>\kappa_{t+1}^{\mathrm{net},-}$. If instead the chosen plus action is a strict net frontier-improving governance edit, then
\[
\Phi_{t+1}^{\mathrm{net}}(\mathsf R_{t+1}^+)>\Phi_{t+1}^{\mathrm{net}}(\mathsf R_{t+1}^-).
\]
Later strict net productive-state separation again requires subsequent exploitation of that strict net frontier advantage. Hence the framework can be growth-promoting even after pricing governance resources with an accounting functional that lower-bounds real protected compute, memory, wall-clock, and energy usage.
\end{corollary}

\begin{proof}
The proof repeats the inductive structure of Corollary~\ref{cor:realized_growth} with $\kappa$ and $\Phi$ replaced by their resource-priced counterparts from Definition~\ref{def:net_growth_metric}. At each epoch the chosen plus action is, by assumption, lexicographically no worse in the net objective than at least one replayably published net witness action for the realized minus action. The net matching clause of Definition~\ref{def:gp_net_class} supplies such a witness whenever the minus-regime chooses a productive action. The net governance-edit clause supplies the witness when the compared plus move is a frontier-improving governance edit. Bounded-exposure entry, selector-invariant review, and Corollary~\ref{cor:service_scaling} keep the delayed challenge, promotion, and proof-lane premises valid along the realized plus trajectory, so delayed good cohorts still enter the deployed state after finite time without invalidating the net accounting. Therefore the same inductive step propagates weak resource-priced dominance to epoch $t+1$. The strict claims are only the immediate strict consequences stated in the corollary; no persistence beyond those steps is asserted without an additional strictness-preserving clause.
\end{proof}

\begin{proposition}[Tagged contradiction reserve conservation under admissible lane retyping]\label{prop:reserve_conservation}
Suppose every admissible promotion or demotion preserves an item's contradiction tag, policy weight, replayability status, and provenance metadata whenever the item remains inside the bounded contradiction window. Then admissible lane retyping leaves $R_t^{\mathrm{contra}}$ invariant. In particular, contradiction mass cannot be silently erased by promoting an item from $\mathcal L_t^{\Het}$ to $\mathcal L_t^{\Core}$.
\end{proposition}

\begin{proof}
By definition, $R_t^{\mathrm{contra}}$ sums policy weights over contradiction-tagged replayable items in $\mathcal L_t^{\Het}\cup\mathcal L_t^{\Core}$ that satisfy the window and provenance requirements. Under admissible retyping, each qualifying item's indicator and weight are preserved. Hence retyping only changes the lane label, not the membership of the item in the reserve sum. Therefore the total reserve is invariant under such retyping.
\end{proof}

\begin{theorem}[Cross-basis semantic-volume anti-dilution law relative to the active fingerprint family]\label{thm:semantic_volume}
Let $E_t$ be the current contradiction reserve and let $A_{t,N}$ denote the set of contradiction-tagged items that are newly admitted during an attack window from $t$ to $t+N$ and still satisfy the reserve-admissibility conditions at time $t+N$. Assume the terminal tagged mass of the admitted attack set is
\[
T_N\Defeq \sum_{x\in A_{t,N}} w_{t+N}(x).
\]
Assume further that during the attack window every pre-existing reserve item preserves its reserve membership and tagged weight, no contradiction-tagged non-attack entry is added to the reserve, and there exists at least one uncaptured fingerprint family $j\in\{1,\dots,K_t^{\psi}\}$ whose encoder remains replay-compatible with the retained audit family. Let the post-attack kernel for that surviving basis, written on the common terminal index set $E_t\cup A_{t,N}$, have block form
\[
L_{t+N}^{(j)}=
\begin{pmatrix}
L & B\\
B^\top & C
\end{pmatrix},
\]
where $L$ is the zero-padded pre-attack kernel on $E_t$, $C\succeq 0$ is the new-item self-kernel, and $B$ encodes cross interactions between pre-existing and newly admitted items. Define the Schur-complement increment
\[
S\Defeq C-B^\top(I+L)^{-1}B\succeq 0.
\]
Assume
\[
\operatorname{rank}(S)\le r,
\qquad
\operatorname{tr}(S)\le T_N.
\]
Then the tagged mass increment is exactly
\[
R_{t+N}^{\mathrm{contra}}-R_t^{\mathrm{contra}}=T_N,
\]
while the basis-specific semantic-volume increment satisfies the exact identity
\[
\mathcal V_{t+N}^{(j)}-\mathcal V_t^{(j)}
=
\log\det\!\bigl(I+S\bigr)
\le
B_r(T_N),
\qquad
B_r(T)\Defeq
\begin{cases}
0, & r=0,\\
r\log\!\left(1+\frac{T}{r}\right), & r\ge 1.
\end{cases}
\]
Hence
\[
\bar{\mathcal V}_{t+N}^{\mathrm{contra}}-\bar{\mathcal V}_t^{\mathrm{contra}}\le B_r(T_N).
\]
Consequently, mass-dilution spam cannot create arbitrarily large robust semantic volume unless the attack simultaneously captures every basis in $\Psi_t$. The conclusion is explicitly relative to the currently active fingerprint family and its replay-compatibility contract; if every basis co-collapses or the transport certificates fail, the theorem becomes inapplicable rather than magically false.
\end{theorem}

\begin{remark}[Blind-spot scope of the semantic-volume law]
Theorem~\ref{thm:semantic_volume} does not identify semantically meaningful diversity in an ontology-free sense. It certifies anti-dilution only relative to the active replay-compatible basis family and the assumption that at least one basis remains uncaptured. A simultaneous co-collapse of all bases, or a transport failure that prevents comparison across basis updates, must therefore be handled by the separate basis-refresh and audit-family interfaces rather than being smuggled into the semantic-volume bound.
\end{remark}

\begin{proof}
By the bookkeeping-isolation assumption, every pre-existing reserve item keeps its membership and weight and no contradiction-tagged non-attack inflow occurs during the attack window. Hence the only contribution to the tagged reserve-mass increment comes from the terminal admitted attack set $A_{t,N}$, so
\[
R_{t+N}^{\mathrm{contra}}-R_t^{\mathrm{contra}}=T_N.
\]
For the surviving uncaptured basis $j$, apply the block determinant identity to the post-attack kernel:
\[
\det(I+L_{t+N}^{(j)})=
\det(I+L)\,\det\!\bigl(I+C-B^\top(I+L)^{-1}B\bigr)=\det(I+L)\,\det(I+S).
\]
Taking logs gives the exact increment identity
\[
\mathcal V_{t+N}^{(j)}-\mathcal V_t^{(j)}=\log\det(I+S).
\]
If $r=0$, then $S=0$ and the increment vanishes. For $r\ge1$, the nonzero eigenvalues of $S$ are nonnegative, have rank at most $r$, and sum to at most $T_N$. The AM--GM eigenvalue bound therefore yields
\[
\log\det(I+S)\le r\log\!\left(1+\frac{T_N}{r}\right)=B_r(T_N).
\]
Because $\bar{\mathcal V}^{\mathrm{contra}}$ is the minimum across bases, it is no larger than the surviving basis-specific volume, giving the displayed robust bound.
\end{proof}

\begin{proposition}[Operational sufficient conditions for the Schur anti-dilution parameters]\label{prop:semantic_volume_operational}
Under the hypotheses of Theorem~\ref{thm:semantic_volume}, suppose the newly admitted contradiction items admit basis-specific fingerprint vectors $\psi_j(x)$ satisfying $\|\psi_j(x)\|\le 1$ for every uncaptured basis $j$, and suppose their weighted fingerprint matrix $F_j$ on the admitted attack set obeys a replayably published effective-novelty bound
\[
\operatorname{rank}(F_j)\le r
\qquad\text{and}\qquad
\sum_{x\in A_{t,N}} w_{t+N}(x)\|\psi_j(x)\|^2\le T_N.
\]
Then the Schur increment $S=C-B^\top(I+L)^{-1}B$ of Theorem~\ref{thm:semantic_volume} satisfies
\[
\operatorname{rank}(S)\le r,
\qquad
\operatorname{tr}(S)\le T_N.
\]
Hence the anti-dilution conclusion of Theorem~\ref{thm:semantic_volume} follows from replayably checkable norm, weight-sum, and effective-novelty-dimension certificates on the newly admitted fingerprints.
\end{proposition}

\begin{proof}
Because $I+L\succeq I$, we have $(I+L)^{-1}\preceq I$, hence
\[
0\preceq S=C-B^\top(I+L)^{-1}B\preceq C.
\]
The Schur complement therefore cannot have larger rank than the new-item self-kernel $C$, and under the effective-novelty certificate the latter has rank at most $r$. Likewise $\operatorname{tr}(S)\le \operatorname{tr}(C)$, while the bounded-norm and weight-sum assumptions imply $\operatorname{tr}(C)\le T_N$. The displayed inequalities follow immediately.
\end{proof}


\begin{corollary}[Mass-only and single-basis reserve floors are spoof-vulnerable]\label{cor:mass_only_spam}
A governance rule that constrains only $R_t^{\mathrm{contra}}$ can be satisfied by sufficiently many near-duplicate contradiction artifacts even when the reserve contributes negligible new semantic directions. A rule that constrains only a \emph{single} basis-specific semantic volume can also be spoofed by rewriting that basis to output artificially orthogonal fingerprints. Requiring cross-basis robust semantic-volume floors or marginal-volume admission thresholds blocks both failure modes, up to simultaneous capture of the full basis family.
\end{corollary}

\begin{proposition}[Basis-transport comparability of robust semantic volume]\label{prop:semantic_transport}
Suppose successive fingerprint families are linked by a replayable transport injection $\vartheta_t:[K_t^{\psi}]\hookrightarrow [K_{t+1}^{\psi}]$ such that for every active predecessor basis $k$,
\[
\underline\theta\,\mathcal V_t^{(k)} - \delta_t \le \mathcal V_{t+1}^{(\vartheta_t(k)),\mathrm{tr}} \le \bar\theta\,\mathcal V_t^{(k)} + \delta_t,
\qquad
0<\underline\theta\le \bar\theta<\infty,
\]
for some public distortion budget $\delta_t\ge 0$. Then the transported robust semantic volume satisfies
\[
\underline\theta\,\bar{\mathcal V}_t^{\mathrm{contra}}-\delta_t
\le
\bar{\mathcal V}_{t+1}^{\mathrm{contra},\mathrm{tr}}
\le
\bar\theta\,\bar{\mathcal V}_t^{\mathrm{contra}}+\delta_t.
\]
Hence any reserve-conservation or garbage-collection statement phrased in terms of the transported robust volume is invariant up to the declared transport envelopes and cannot be made to hold merely by silently changing the fingerprint basis. Extra newly introduced bases may lower the full successor minimum $\bar{\mathcal V}_{t+1}^{\mathrm{contra}}$, but they cannot retroactively alter the transported comparison with the predecessor family.
\end{proposition}

\begin{proof}
Take the minimum of the transported basis-wise inequalities over the predecessor index set $k\in[K_t^{\psi}]$. Because $\vartheta_t$ is an injection, the transported robust volume is exactly the minimum over the successor bases reached from the predecessor family, so the displayed lower and upper bounds follow immediately. The final sentence is then just a restatement that only replayably transported comparisons may be used for temporal semantic-volume accounting.
\end{proof}
\begin{corollary}[Multi-step transport and cocycle accumulation]\label{cor:semantic_transport_cocycle}
Suppose Proposition~\ref{prop:semantic_transport} holds along a replayably composable sequence of basis updates from time $t$ to time $t+n-1$, with per-step parameters $(\underline\theta_s,\bar\theta_s,\delta_s)$ and successor injections whose compositions remain replayably defined. Then the transported robust semantic volume satisfies
\[
\Bigl(\prod_{s=t}^{t+n-1}\underline\theta_s\Bigr)\bar{\mathcal V}_t^{\mathrm{contra}} - \sum_{j=0}^{n-1}\Bigl(\prod_{s=t+j+1}^{t+n-1}\underline\theta_s\Bigr)\delta_{t+j}
\le
\bar{\mathcal V}_{t+n}^{\mathrm{contra},\mathrm{tr}}
\le
\Bigl(\prod_{s=t}^{t+n-1}\bar\theta_s\Bigr)\bar{\mathcal V}_t^{\mathrm{contra}} + \sum_{j=0}^{n-1}\Bigl(\prod_{s=t+j+1}^{t+n-1}\bar\theta_s\Bigr)\delta_{t+j}.
\]
Hence temporal semantic accounting obeys a cocycle law: long-run comparison error is the multiplicative transport distortion plus an accumulated additive defect, not an arbitrary free reparameterization at each step.
\end{corollary}

\begin{proof}
Apply Proposition~\ref{prop:semantic_transport} recursively. The upper bound after two steps is
\[
\bar{\mathcal V}_{t+2}^{\mathrm{contra},\mathrm{tr}}
\le
\bar\theta_{t+1}(\bar\theta_t\bar{\mathcal V}_t^{\mathrm{contra}}+\delta_t)+\delta_{t+1},
\]
and likewise for the lower bound with $\underline\theta$. Iterating this affine recursion and using replayable composability of the injections yields the displayed products and sums. The result is precisely the cocycle accumulation law for multiplicative distortion plus additive defect.
\end{proof}

\section{Sharpened Toy Models}
\subsection{Single-agent scalar versioned code-and-retrieval commons}
\begin{example}[Scalar code-and-retrieval commons]\label{ex:toy_single}
At time $t$, the state is
\[
(x_t,e_t,m_t,q_t,r_t)\in\R\times\R\times\R_+\times\R_+\times\R_+.
\]
Interpretation:
\begin{itemize}[leftmargin=1.8em]
  \item $x_t$ is task capability,
  \item $e_t$ is evaluator drift or ruler bias,
  \item $m_t$ is memory distortion load,
  \item $q_t$ is verification backlog,
  \item $r_t$ is aggregate latent error debt.
\end{itemize}
A proposal chooses a nonnegative step size $u_t\in[0,\bar u]$. Dynamics are
\begin{align}
  x_{t+1} &= x_t + a u_t, \\
  e_{t+1} &= e_t + \gamma u_t, \\
  m_{t+1} &= \rho_m m_t + \kappa u_t, \\
  q_{t+1} &= \bigl[q_t + \nu u_t - \mu\bigr]_+, 
\end{align}
with fast public score $J_t=x_t+e_t$, so the claimed improvement is
\[
\Delta_t^{\mathrm{claim}}=(a+\gamma)u_t.
\]
Debt is tracked cohort by cohort using the revised downside-reserve semantics. Let the structural bond be
\[
\mathcal B_\tau(u_\tau)=\lambda_B u_\tau,
\]
let the repair reserve be $U_\tau^{\mathrm{rep}}=\lambda_{\mathrm{rep}}\nu u_\tau$, and let the persistent hazard reserve be $U_\tau^{\mathrm{haz}}=\lambda_{\mathrm{haz}}\gamma u_\tau$, with nonnegative policy constants $\lambda_B,\lambda_{\mathrm{rep}},\lambda_{\mathrm{haz}}$. For each cohort $\tau$, let $g_{\tau,t}^{\mathrm{chal}}\ge 0$ be nondecreasing cumulative matched retirement, let $a_{\tau,t}^{\mathrm{sec}}\ge 0$ be secondary liability, and let $b_{\tau,t}\ge 0$ satisfy $\sum_{\tau\le t}b_{\tau,t}\le q_t$. Then
\[
\eta_{\tau,\tau}=\lambda_B u_\tau+\lambda_{\mathrm{rep}}\nu u_\tau+\lambda_{\mathrm{haz}}\gamma u_\tau,
\]
\[
\eta_{\tau,t+1}=\Bigl[(1-\rho_{\tau,t+1}^{\mathrm{ret}})\alpha_r\eta_{\tau,t}-(g_{\tau,t+1}^{\mathrm{chal}}-g_{\tau,t}^{\mathrm{chal}})+a_{\tau,t+1}^{\mathrm{sec}}+\lambda_q b_{\tau,t}\Bigr]_+,
\qquad
r_t=\sum_{\tau<t}\eta_{\tau,t}.
\]
This toy now instantiates the main text's downside-reserve debt semantics rather than the earlier claim-residual bookkeeping variant. The strict-conservation specialization of Theorem~\ref{thm:strict_conservation} appears as the limit $\alpha_r\uparrow 1$ with separate repair throughput.
The slower utility surrogate is
\[
V_t = x_t - c_e\abs{e_t} - c_m m_t - c_q q_t,
\]
with $c_e,c_m,c_q>0$.
\end{example}
\begin{proposition}[Single-agent full and marginal stable increments]\label{prop:toy_single_threshold}
Assume $e_t\ge0$. In Example~\ref{ex:toy_single}, the full one-step stable increment is
\[
\Delta_t^{\mathrm{full}}
=
a u_t - c_e\gamma u_t - c_m\bigl((\rho_m-1)m_t+\kappa u_t\bigr)-c_q\bigl(q_{t+1}-q_t\bigr).
\]
In the active-backlog regime $q_t+\nu u_t\ge \mu$, this simplifies to
\[
\Delta_t^{\mathrm{full}}
=
a u_t - c_e\gamma u_t - c_m\bigl((\rho_m-1)m_t+\kappa u_t\bigr)-c_q(\nu u_t-\mu).
\]
The marginal stable increment obtained by suppressing state-retention terms is
\[
\Delta_t^{\mathrm{marg}}
=
a u_t - c_e\gamma u_t - c_m\kappa u_t - c_q\nu u_t.
\]
Hence every positive proposal $u_t>0$ has positive claimed improvement and positive marginal stable increment whenever
\[
a > c_e\gamma + c_m\kappa + c_q\nu.
\]
If instead
\[
a+\gamma>0
\quad\text{but}\quad
a < c_e\gamma + c_m\kappa + c_q\nu,
\]
then every positive proposal has positive claimed improvement but negative marginal stable increment.
\end{proposition}

\begin{proof}
The claimed increment follows from $J_t=x_t+e_t$. For the slower utility,
\[
V_{t+1}-V_t = a u_t - c_e(e_{t+1}-e_t) - c_m(m_{t+1}-m_t) - c_q(q_{t+1}-q_t),
\]
which gives the displayed full increment after substitution. The marginal increment drops the state-retention contributions $(\rho_m-1)m_t$ and $-\mu$. The sign statements then follow immediately.
\end{proof}

\subsection{Stochastic shared-yardstick multi-agent commons with agent-dependent delayed challenge policy}
\begin{example}[Stochastic multi-agent shared-yardstick toy model with adaptive challenge policy]\label{ex:toy_multi}
Let $n$ agents share one fast evaluator, one memory substrate, and one verification queue. Agent $i$ proposes a step of magnitude $u_i>0$ whenever a Bernoulli arrival $A_{i,t}\in\{0,1\}$ fires at time $t$, with
\[
\Prob(A_{i,t}=1\mid \mathcal F_t)=p_{i,t},
\qquad
0\le p_{i,t}\le \bar p_i.
\]
The aggregate proposal mass is
\[
M_t^{\mathrm{prop}}\Defeq \sum_{i=1}^n u_i A_{i,t}.
\]
The shared fast state is
\[
(x_t,e_t,m_t,q_t,r_t)\in\R\times\R\times\R_+\times\R_+\times\R_+,
\]
and the dynamics are
\begin{align}
  x_{t+1} &= x_t + \sum_{i=1}^n a_i u_i A_{i,t}, \\
  e_{t+1} &= e_t + \sum_{i=1}^n \gamma_i u_i A_{i,t}, \\
  m_{t+1} &= \rho_m m_t + \sum_{i=1}^n \kappa_i u_i A_{i,t}, \\
  q_{t+1} &= \Bigl[q_t + \sum_{i=1}^n \nu_i u_i A_{i,t} - \mu\Bigr]_+,
\end{align}
with fast score $J_t=x_t+e_t$ and slow utility surrogate
\[
V_t = x_t - c_e\abs{e_t} - c_m m_t - c_q q_t.
\]
For each realized proposal $(i,\tau)$, an agent-dependent delayed challenge policy selects it for shadow evaluation with predictable probability
\[
\sigma_{i,\tau}=\pi_i(C_\tau)\in[\underline\sigma_i,\bar\sigma_i]\subset(0,1],
\]
and assigns a predictable integer delay
\[
L_{i,\tau}=\ell_i(C_\tau)\in\{1,2,\dots,\bar L_i\}.
\]
Let $\widetilde S_{i,\tau}\in\{0,1\}$ denote the conditional sampling flag with
\[
\Prob(\widetilde S_{i,\tau}=1\mid \mathcal F_\tau,A_{i,\tau}=1)=\sigma_{i,\tau},
\]
and define $S_{i,\tau}\Defeq A_{i,\tau}\widetilde S_{i,\tau}$ so that $S_{i,\tau}=0$ whenever the proposal is not admitted. A sampled proposal from agent $i$ at cohort time $\tau$ yields, at maturity time $t=\tau+L_{i,\tau}$, a high-fidelity shadow outcome $Y_{i,\tau\to t}$ satisfying
\[
\E\bigl[Y_{i,\tau\to t}\mid \mathcal F_\tau, A_{i,\tau}=1,\widetilde S_{i,\tau}=1\bigr] = g_i u_i,
\qquad
g_i\Defeq a_i-c_e\gamma_i-c_m\kappa_i-c_q\nu_i.
\]
Define the cohort-wise Horvitz--Thompson clean-shadow estimator \cite{HorvitzThompson1952}
\[
\widehat G_\tau^{\mathrm{clean}}\Defeq \sum_{i=1}^n \frac{S_{i,\tau}}{\sigma_{i,\tau}}Y_{i,\tau\to \tau+L_{i,\tau}}.
\]
Let the actually observed shadow monitor be
\[
\widehat G_\tau^{\mathrm{shad}}\Defeq \widehat G_\tau^{\mathrm{clean}} + B_\tau^{\mathrm{coll}},
\]
where $B_\tau^{\mathrm{coll}}$ is the integrity disturbance induced by cross-proposal evaluator capture or shadow-family collusion. For debt, let the structural bond, repair reserve, and hazard reserve of a realized proposal $(i,\tau)$ be
\[
\mathcal B_{i,\tau}=\lambda_{B,i}u_iA_{i,\tau},
\qquad
U_{i,\tau}^{\mathrm{rep}}=\lambda_{\mathrm{rep},i}\nu_i u_iA_{i,\tau},
\qquad
U_{i,\tau}^{\mathrm{haz}}=\lambda_{\mathrm{haz},i}\gamma_i u_iA_{i,\tau},
\]
with nonnegative policy constants. Let $g_{i,\tau,t}^{\mathrm{chal}}$ be a predictable nondecreasing matched-retirement process and let $a_{i,\tau,t}^{\mathrm{sec}}\ge 0$ be secondary liability. The challenge estimator $\widehat G_\tau^{\mathrm{shad}}$ is a monitoring statistic for stable gain; it need not equal the matched retirement process used in debt accounting. Let $b_{i,\tau,t}\ge0$ be a backlog-attribution kernel satisfying
\[
\sum_{i=1}^n\sum_{\tau\le t} b_{i,\tau,t}\le q_t.
\]
Then proposal-local escrow debt evolves by
\[
\eta_{i,\tau,\tau}=\mathcal B_{i,\tau}+U_{i,\tau}^{\mathrm{rep}}+U_{i,\tau}^{\mathrm{haz}},
\qquad
\eta_{i,\tau,t+1}
=
\Bigl[
(1-\rho_{i,\tau,t+1}^{\mathrm{ret}})\alpha_r\eta_{i,\tau,t}
-(g_{i,\tau,t+1}^{\mathrm{chal}}-g_{i,\tau,t}^{\mathrm{chal}})
+a_{i,\tau,t+1}^{\mathrm{sec}}
+\lambda_q b_{i,\tau,t}
\Bigr]_+,
\]
and aggregate debt is
\[
r_t\Defeq \sum_{i=1}^n\sum_{\tau<t}\eta_{i,\tau,t}.
\]
Finally, define the per-step downside-reserve inflow
\[
G_t^{\mathrm{res}}
\Defeq
\sum_{i=1}^n
A_{i,t}\Bigl(
\lambda_{B,i}u_i+\lambda_{\mathrm{rep},i}\nu_i u_i+\lambda_{\mathrm{haz},i}\gamma_i u_i
\Bigr),
\]
and the aggregate secondary-liability inflow
\[
A_t^{\mathrm{sec}}\Defeq \sum_{i=1}^n\sum_{\tau\le t} a_{i,\tau,t}^{\mathrm{sec}}.
\]
These are the reserve-side inflow variables relevant for debt ceilings in the revised semantics; claim mass remains useful for fast-score accounting but is not the debt initializer.
\end{example}

\begin{proposition}[Expected one-step stable increments under adaptive challenge policy]\label{prop:toym_increment}
Assume $e_t\ge0$. In Example~\ref{ex:toy_multi},
\[
\E[\Delta_t^{\mathrm{claim}}\mid \mathcal F_t]=\sum_{i=1}^n p_{i,t}(a_i+\gamma_i)u_i.
\]
The full stable increment satisfies
\[
\E[\Delta_t^{\mathrm{full}}\mid \mathcal F_t]
=
\sum_{i=1}^n p_{i,t}(a_i-c_e\gamma_i-c_m\kappa_i)u_i
-c_m(\rho_m-1)m_t
-c_q\E[q_{t+1}-q_t\mid\mathcal F_t].
\]
In the active-queue regime, where reflection is inactive, this becomes
\[
\E[\Delta_t^{\mathrm{full}}\mid \mathcal F_t]
=
\sum_{i=1}^n p_{i,t}g_i u_i
-c_m(\rho_m-1)m_t
+c_q\mu.
\]
The marginal stable increment is
\[
\E[\Delta_t^{\mathrm{marg}}\mid \mathcal F_t]=\sum_{i=1}^n p_{i,t}g_i u_i.
\]
Hence positive expected claimed improvement can coexist with negative expected marginal stable improvement whenever
\[
\sum_{i=1}^n p_{i,t}(a_i+\gamma_i)u_i >0
\quad\text{but}\quad
\sum_{i=1}^n p_{i,t}g_i u_i <0.
\]
\end{proposition}

\begin{proof}
Taking conditional expectations in the state equations and using $\E[A_{i,t}\mid\mathcal F_t]=p_{i,t}$ yields the stated formulas. The active-queue simplification uses $q_{t+1}-q_t=\sum_i\nu_i u_iA_{i,t}-\mu$ when the reflected queue stays strictly positive.
\end{proof}


\begin{proposition}[Delayed clean-shadow estimator under agent-dependent sampling and delay]\label{prop:shadow_unbiased}
For every proposal cohort at time $\tau$ in Example~\ref{ex:toy_multi},
\[
\E\bigl[\widehat G_\tau^{\mathrm{clean}}\mid \mathcal F_\tau\bigr] = \sum_{i=1}^n g_i u_i A_{i,\tau}.
\]
Let
\[
X_{i,\tau}\Defeq \frac{S_{i,\tau}}{\sigma_{i,\tau}}Y_{i,\tau\to \tau+L_{i,\tau}}-g_i u_i A_{i,\tau}.
\]
If, in addition,
\[
\Var(Y_{i,\tau\to \tau+L_{i,\tau}}\mid \mathcal F_\tau,S_{i,\tau}=1,A_{i,\tau}=1)\le \bar v_i u_i^2,
\]
and there exists a predictable positive-semidefinite matrix envelope $\bar\Sigma_\tau\in\R_+^{n\times n}$ such that
\[
\Cov\bigl((X_{1,\tau},\dots,X_{n,\tau})\mid \mathcal F_\tau\bigr)\preceq \bar\Sigma_\tau,
\]
then
\[
\Var\bigl(\widehat G_\tau^{\mathrm{clean}}\mid \mathcal F_\tau\bigr)
\le
\mathbf 1^{\top}\bar\Sigma_\tau\mathbf 1.
\]
\end{proposition}

\begin{proof}
Unbiasedness follows from the predictability of $\sigma_{i,\tau}$ and $L_{i,\tau}$:
\[
\E\left[\frac{S_{i,\tau}}{\sigma_{i,\tau}}Y_{i,\tau\to \tau+L_{i,\tau}}\middle|\mathcal F_\tau\right]=g_i u_i A_{i,\tau}.
\]
The centered sum of the $X_{i,\tau}$ is exactly $\widehat G_\tau^{\mathrm{clean}}-\sum_i g_i u_i A_{i,\tau}$. Therefore
\[
\Var\bigl(\widehat G_\tau^{\mathrm{clean}}\mid \mathcal F_\tau\bigr)
=
\mathbf 1^{\top}\Cov\bigl((X_{1,\tau},\dots,X_{n,\tau})\mid \mathcal F_\tau\bigr)\mathbf 1
\le
\mathbf 1^{\top}\bar\Sigma_\tau\mathbf 1.
\]
\end{proof}

\begin{proposition}[Collusion-robust shadow monitor]\label{prop:collusion_shadow}
Under Example~\ref{ex:toy_multi}, the observed shadow monitor satisfies
\[
\abs{\E\bigl[\widehat G_\tau^{\mathrm{shad}}\mid \mathcal F_\tau\bigr]-\sum_{i=1}^n g_i u_i A_{i,\tau}}
\le
\Gamma_\tau^{\mathrm{coll}}.
\]
If, moreover,
\[
\Var(B_\tau^{\mathrm{coll}}\mid \mathcal F_\tau)\le v_\tau^{\mathrm{coll}}
\]
and
\[
\abs{\Cov(\widehat G_\tau^{\mathrm{clean}},B_\tau^{\mathrm{coll}}\mid \mathcal F_\tau)}\le c_\tau^{\mathrm{cross}},
\]
then
\[
\Var(\widehat G_\tau^{\mathrm{shad}}\mid \mathcal F_\tau)
\le
\mathbf 1^{\top}\bar\Sigma_\tau\mathbf 1 + v_\tau^{\mathrm{coll}} + 2c_\tau^{\mathrm{cross}}.
\]
Consequently any sign certification rule based on the observed shadow monitor must subtract not only calibration error but also the integrity budget $\Gamma_\tau^{\mathrm{coll}}$.
\end{proposition}

\begin{proof}
The expectation statement is immediate from
\[
\widehat G_\tau^{\mathrm{shad}}=\widehat G_\tau^{\mathrm{clean}}+B_\tau^{\mathrm{coll}}
\]
and Proposition~\ref{prop:shadow_unbiased}. The variance identity expands as
\[
\Var(\widehat G_\tau^{\mathrm{shad}}\mid \mathcal F_\tau)
=
\Var(\widehat G_\tau^{\mathrm{clean}}\mid \mathcal F_\tau)+\Var(B_\tau^{\mathrm{coll}}\mid \mathcal F_\tau)+2\Cov(\widehat G_\tau^{\mathrm{clean}},B_\tau^{\mathrm{coll}}\mid \mathcal F_\tau),
\]
then apply the displayed bounds. The final statement follows because a positive observed margin smaller than $\Gamma_\tau^{\mathrm{coll}}$ can still be explained entirely by evaluator capture.
\end{proof}

\begin{remark}[From oracle collusion budgets to endogenous audit envelopes]
Proposition~\ref{prop:collusion_shadow} is an interface theorem: it says how much certified margin must be subtracted once an integrity budget is available. In the endogenous-audit formulation that budget is no longer interpreted as an externally gifted constant. Theorem~\ref{thm:epistemic_singularity} supplies an endogenous replacement from cross-shadow disagreement, audit-family overlap, and declared attack reach.
\end{remark}

\begin{corollary}[Pairwise-uncorrelated specialization]\label{cor:shadow_var_pairwise}
Under the assumptions of Proposition~\ref{prop:collusion_shadow}, if the centered weighted clean contributions are conditionally pairwise uncorrelated given $\mathcal F_\tau$, then
\[
\Var\bigl(\widehat G_\tau^{\mathrm{shad}}\mid \mathcal F_\tau\bigr)
\le
\sum_{i=1}^n A_{i,\tau}u_i^2\left(\frac{\bar v_i}{\sigma_{i,\tau}}+\frac{1-\sigma_{i,\tau}}{\sigma_{i,\tau}}g_i^2\right)+v_\tau^{\mathrm{coll}}+2c_\tau^{\mathrm{cross}}.
\]
\end{corollary}

\begin{proof}
Under pairwise conditional uncorrelatedness of the centered weighted clean contributions, the matrix envelope in Proposition~\ref{prop:shadow_unbiased} may be chosen diagonal with entries bounded by the displayed per-agent terms. Substituting that diagonal envelope into Proposition~\ref{prop:collusion_shadow} and then adding the collusion-variance and cross-covariance envelopes yields the result.
\end{proof}


\begin{remark}[What debt anti-masking does and does not solve]
Proposition~\ref{prop:anti_masking} blocks one failure mode --- netting a bad proposal's liability against unrelated good cohorts. It does \emph{not} by itself prevent evaluator capture or collusive score inflation. Those risks enter through the integrity disturbance $B_\tau^{\mathrm{coll}}$ and must be budgeted separately, as in Proposition~\ref{prop:collusion_shadow}.
\end{remark}

\begin{proposition}[Cohort-escrow anti-masking]\label{prop:anti_masking}
In Example~\ref{ex:toy_multi}, the debt coordinate for proposal $(i,\tau)$ is a measurable function only of the tuple
\[
\bigl(\eta_{i,\tau,t},g_{i,\tau,t}^{\mathrm{chal}},g_{i,\tau,t+1}^{\mathrm{chal}},a_{i,\tau,t+1}^{\mathrm{sec}},b_{i,\tau,t},\rho_{i,\tau,t+1}^{\mathrm{ret}}\bigr).
\]
In particular, matched corrections or strong shadow outcomes generated by other proposals never enter the update for $\eta_{i,\tau,t+1}$ with a negative sign. Hence cross-cohort masking is impossible: unrelated good cohorts cannot retire the discounted liability stock created by proposal $(i,\tau)$.
\end{proposition}

\begin{proof}
Read the recursion for $\eta_{i,\tau,t+1}$ in Example~\ref{ex:toy_multi}. Only proposal-local retirement, proposal-local secondary liability, and proposal-local backlog attribution appear. Terms from other cohorts are absent from this coordinate and therefore cannot reduce it.
\end{proof}
\begin{theorem}[Predictable verification queue phase diagram]\label{thm:toym_queue}
Consider the queue process in Example~\ref{ex:toy_multi} and let
\[
B_t\Defeq \sum_{i=1}^n \nu_i u_i A_{i,t}
\]
be the random proposal workload arriving at time $t$. Define the predictable expected proposal load
\[
\Lambda_t\Defeq \sum_{i=1}^n p_{i,t}\nu_i u_i.
\]
Then:
\begin{enumerate}[label=(\roman*), leftmargin=2em]
  \item If there exists $\eps_\Lambda>0$ such that $\Lambda_t\ge \mu+\eps_\Lambda$ whenever $q_t>0$, then on every active step,
  \[
  \E[q_{t+1}-q_t\mid \mathcal F_t] \ge \eps_\Lambda.
  \]
  Consequently, any policy that keeps the queue active over a nonvanishing fraction of time incurs at least linear expected backlog growth on that active subsequence.
  \item If there exists $\eps_\mu>0$ such that $\Lambda_t\le \mu-\eps_\mu$ for all $t$ and the centered workload has uniformly bounded conditional second moment, then the quadratic drift inequality of Proposition~\ref{prop:queue_first_moment} holds, yielding bounded time-average backlog. This is the correct non-Markov statement for the predictable-load setting; no positive-recurrence claim is needed.
\end{enumerate}
\end{theorem}

\begin{proof}
If $q_t>0$, reflection is inactive and $q_{t+1}-q_t=B_t-\mu$, so the conditional drift is $\Lambda_t-\mu$. This proves part (i). Part (ii) is exactly the negative-drift regime handled by Proposition~\ref{prop:queue_first_moment}.
\end{proof}


\begin{proposition}[Time-average backlog bound under service slack and second-moment control]\label{prop:queue_first_moment}
In Example~\ref{ex:toy_multi}, assume there exists $\eps_\mu>0$ such that
\[
\Lambda_t\le \mu-\eps_\mu
\qquad\text{for all }t,
\]
and the centered one-step workload has uniformly bounded conditional second moment:
\[
\E\bigl[(B_t-\mu)^2\mid \mathcal F_t\bigr]\le V_B
\qquad\text{for all }t.
\]
Then the quadratic Lyapunov function $L(q)=q^2$ satisfies
\[
\E[q_{t+1}^2-q_t^2\mid \mathcal F_t]\le -2\eps_\mu q_t + V_B.
\]
Consequently, for every horizon $T\ge1$,
\[
\frac1T\sum_{t=0}^{T-1}\E[q_t]
\le
\frac{\E[q_0^2]}{2\eps_\mu T}+\frac{V_B}{2\eps_\mu},
\]
and hence
\[
\limsup_{T\to\infty}\frac1T\sum_{t=0}^{T-1}\E[q_t]
\le
\frac{V_B}{2\eps_\mu}.
\]
This is a standard Lyapunov-drift guarantee for general stochastic systems with predictable load and bounded conditional quadratic variation \cite{Neely2010}.
\end{proposition}

\begin{proof}
Using $[x]_+^2\le x^2$ and $q_{t+1}=[q_t+B_t-\mu]_+$,
\[
q_{t+1}^2-q_t^2
\le
2q_t(B_t-\mu)+(B_t-\mu)^2.
\]
Take conditional expectations and use $\E[B_t\mid\mathcal F_t]=\Lambda_t\le \mu-\eps_\mu$ together with the assumed second-moment bound to obtain the displayed drift inequality. Summing over $t=0,\dots,T-1$ and telescoping yields the time-average backlog bound.
\end{proof}
\begin{theorem}[Discounted debt envelope under bounded reserve inflow, secondary liability, and queue cap]\label{thm:shadow_debt}
In Example~\ref{ex:toy_multi}, suppose the per-step downside-reserve inflow, aggregate secondary-liability inflow, and verification backlog are pathwise bounded by
\[
G_t^{\mathrm{res}}\le \bar G^{\mathrm{res}},
\qquad
A_t^{\mathrm{sec}}\le \bar A^{\mathrm{sec}},
\qquad
q_t\le B_q
\qquad\text{for all }t.
\]
Then aggregate debt satisfies the pathwise recursion
\[
r_{t+1}\le \alpha_r r_t + \bar G^{\mathrm{res}} + \bar A^{\mathrm{sec}} + \lambda_q B_q,
\]
and therefore obeys the deterministic ceiling
\[
r_t
\le
\alpha_r^t r_0 + \frac{\bar G^{\mathrm{res}} + \bar A^{\mathrm{sec}} + \lambda_q B_q}{1-\alpha_r}
\qquad\text{for all }t\ge0.
\]
The result does not require full challenge coverage: bounded debt follows from bounded reserve inflow, bounded secondary liability, discounted persistence, and queue discipline, while challenge coverage and delay matter primarily for identifiability and timely certification.
\end{theorem}

\begin{proof}
For any existing cohort $(i,\tau)$, the matched-retirement process is nondecreasing and the retirement fraction satisfies $\rho_{i,\tau,t+1}^{\mathrm{ret}}\in[0,1]$, hence
\[
\eta_{i,\tau,t+1}\le \alpha_r\eta_{i,\tau,t}+a_{i,\tau,t+1}^{\mathrm{sec}}+\lambda_q b_{i,\tau,t}.
\]
New cohorts entering at time $t$ contribute initial stock exactly equal to their downside reserves, whose aggregate is at most $G_t^{\mathrm{res}}\le \bar G^{\mathrm{res}}$. Summing over all cohorts and using $\sum_{i,\tau\le t}a_{i,\tau,t+1}^{\mathrm{sec}}\le \bar A^{\mathrm{sec}}$ together with $\sum_{i,\tau\le t}b_{i,\tau,t}\le q_t\le B_q$ gives
\[
r_{t+1}\le \alpha_r r_t + \bar G^{\mathrm{res}} + \bar A^{\mathrm{sec}} + \lambda_q B_q.
\]
Iterating the affine recursion yields the displayed ceiling.
\end{proof}

\begin{corollary}[Mean debt bound under bounded mean reserve inflow and backlog envelope]\label{cor:shadow_debt_mean}
If instead of pathwise caps one only knows that
\[
\sup_t \E[G_t^{\mathrm{res}}]\le \bar G^{\mathrm{res}},
\qquad
\sup_t \E[A_t^{\mathrm{sec}}]\le \bar A^{\mathrm{sec}},
\qquad
\sup_t \E[q_t]\le \bar q,
\]
then
\[
\E[r_t]
\le
\alpha_r^t r_0 + \frac{\bar G^{\mathrm{res}} + \bar A^{\mathrm{sec}} + \lambda_q\bar q}{1-\alpha_r}
\qquad\text{for all }t\ge0.
\]
\end{corollary}

\begin{proof}
Take expectations in the one-step domination
\[
r_{t+1}\le \alpha_r r_t + G_t^{\mathrm{res}} + A_t^{\mathrm{sec}} + \lambda_q q_t,
\]
and iterate.
\end{proof}

\begin{corollary}[Delayed challenge policy law]\label{cor:delayed_challenge_policy}
A stochastic multi-agent improvement loop needs two different control layers. Positive delayed challenge sampling with bounded delay is what makes hidden harmful edits statistically identifiable. Queue discipline and bounded claim inflow are what keep discounted debt operationally bounded. Confusing these two jobs leads to bad policy: more sampling does not by itself repair congestion, and more throughput does not by itself resolve ruler-editing ambiguity.
\end{corollary}

\begin{corollary}[Deterministic homogeneous threshold]\label{cor:toym_det_threshold}
If $A_{i,t}\equiv 1$ and $a_i=a$, $\gamma_i=\gamma$, $\kappa_i=\kappa$, $\nu_i=\nu$, $u_i=u$ for all $i$, then $\Lambda_t\equiv n\nu u$ and Theorem~\ref{thm:toym_queue} reduces to the deterministic threshold
\[
n_{\mathrm{crit}} = \frac{\mu}{\nu u}.
\]
Thus the deterministic phase boundary is the unit-arrival specialization of the stochastic law.
\end{corollary}

\begin{corollary}[Shared-ruler scaling under random proposals]\label{cor:shared_ruler_scaling}
Under Example~\ref{ex:toy_multi}, expected claimed improvement scales as
\[
\sum_{i=1}^n p_{i,t}(a_i+\gamma_i)u_i,
\]
while verification slack scales as $\mu-\Lambda_t$. Therefore systems that look locally beneficial on a per-arrival basis can nevertheless become globally unstable once random aggregate pressure pushes the predictable load $\Lambda_t$ above service capacity.
\end{corollary}
\section{Operationalization and Falsification}
\subsection{Two-regime comparator-and-challenge toy for the revised hard core}
\begin{example}[Two-regime ruler-edit toy with class/block challenge certification]\label{ex:toy_ruleredit}
Consider two regimes $\mathsf R^-$ and $\mathsf R^+$ that differ only by a ruler edit affecting the fast evaluator and challenge protocol. Both regimes share the same sanitized menu $\widetilde{\mathcal U}=\{u^{\mathrm{safe}},u^{\mathrm{risky}}\}$ and the same public slow-utility surrogate $V$. The current proposal belongs to class $\mathcal C_*$ and block-compatibility family $\mathfrak B_*$. Under both regimes, the block-frozen challenge target for the safe proposal is
\[
\mathcal M_t^{\mathrm{chal}}(1;\mathcal C_*,\mathfrak B_*)=m_*>0,
\]
with replayable lower certificate $\underline m_*$. The ruler edit in $\mathsf R^+$ weakly improves every Interface-0 predicate, preserves the same class/block assignment, lowers the transport residual from $\delta_-^{\mathrm{tr}}$ to $\delta_+^{\mathrm{tr}}\le \delta_-^{\mathrm{tr}}$, lowers comparator diameter from $d_-^{\mathrm{cmp}}$ to $d_+^{\mathrm{cmp}}\le d_-^{\mathrm{cmp}}$, and weakly improves every Interface-C floor, while proposal-local debt and challenge backlog are no worse. Then the operational one-step lower bound for $u^{\mathrm{safe}}$ satisfies
\[
\underline L_t^+(u^{\mathrm{safe}})
= a_+\underline m_* - b_+ - \delta_+^{\mathrm{tr}} - d_+^{\mathrm{cmp}}
\ge
a_-\underline m_* - b_- - \delta_-^{\mathrm{tr}} - d_-^{\mathrm{cmp}}
= \underline L_t^-(u^{\mathrm{safe}}),
\]
whenever the published coefficients satisfy $a_+\ge a_-\ge0$ and $b_+\le b_-$. This toy directly instantiates the class/block challenge certificate, transport residual, comparator burden, and one-step admissibility monotonicity used in the main text.
\end{example}

\begin{proposition}[One-step operator monotonicity in the ruler-edit toy]\label{prop:toy_ruleredit_mono}
In Example~\ref{ex:toy_ruleredit}, if $\mathsf R^+ \succeq^{\mathrm{feas}}_{t,\Pi} \mathsf R^-$ and the safe proposal is admissible under $\mathsf R^-$, then it is admissible under $\mathsf R^+$ and its replayable one-step lower stable-gain bound is no smaller under $\mathsf R^+$ than under $\mathsf R^-$. If the active class/block pair lacks the minimum matured support count required by the manifest, both regimes are fail-closed for fast deployment and the proposal must be routed to challenge, proof escalation, or sandbox.
\end{proposition}

\begin{proof}
The clause-wise assumptions are exactly those required by Definition~\ref{def:feasible_order}. Therefore every Interface-0/A/B/C clause holding under $\mathsf R^-$ also holds under $\mathsf R^+$. The displayed inequality in Example~\ref{ex:toy_ruleredit} gives the monotonicity of the one-step replayable lower bound, while the final statement is the fail-closed convention from Definition~\ref{def:challenge_design} and Definition~\ref{def:confidence_manifest}.
\end{proof}

\subsection{Operational proxies}
\begin{center}
\begin{tabular}{p{0.28\linewidth}p{0.33\linewidth}p{0.32\linewidth}}
\toprule
Quantity & Observable proxy & Operational use \\
\midrule
Yardstick drift $D_t^{\mathrm{yard}}$ & score drift on frozen challenge cohort or transported probe family & refresh / quarantine trigger \\
Challenge certificate & class-conditional delayed-challenge lower CS $\underline M^{\mathrm{chal}}_{t,n}(h;\mathcal C_t,\mathfrak B_t)$ under blocked randomized challenge & delayed-evidence component of Interface-A gain gate \\
Transport calibration & Tier-2 matched-replay transport certificate $(a_h,b_h,\delta_h^{\mathrm{tr}})$ & convert challenge evidence to slow-utility scale for Interface A \\
Comparator burden & witness-family distortion budget $\bar\delta_{t,h}^{\mathrm{cmp}}$ and diameter $\operatorname{diam}^{\mathrm{cmp}}_{t,H}(u_t)$ & robustify stable-gain admission and governance-burden pricing \\
Peak debt $\eta_{t,H}^{\max}$ & cohort-escrow unresolved mass + structural bond + rollback debt & debt ceiling and throttle control \\
Contradiction reserve $R_t^{\mathrm{contra}}$ & tagged contradiction mass & preserve disagreement floor \\
Robust semantic volume $\bar{\mathcal V}_t^{\mathrm{contra}}$ & cross-basis log-det diversity with basis-transport certificates & block duplicate spam / basis capture \\
Heterodoxy routing load $\mathfrak A_t^{\Het}$ & contradiction-tagged retrieval sensitivity under ablation & detect attentional bypass \\
Verification pressure $\Lambda/\mu$ & predictable arrivals per effective service unit & queue stability control \\
Shadow disagreement & online-vs-shadow score gap & ruler-editing / hidden-state alarm \\
\bottomrule
\end{tabular}
\end{center}

\subsection{Design consequences}
\paragraph{Challenge-lane law.}
If evaluator-editing proposals are admitted directly into production without challenge-lane delay, then the observational ambiguity of Theorem~\ref{thm:ruler_nogo} cannot be ruled out by public claim traces alone.

\paragraph{Slow-meter law.}
A sufficient architectural response to ruler-editing ambiguity is to separate a fast online metric from a slower, higher-fidelity challenge process that is not directly editable by the currently evaluated proposal.

\paragraph{Shadow-certification law.}
Retained margins become decision-relevant only after calibration. Proposition~\ref{prop:shadow_bridge} therefore recommends storing challenge-lane retained margins together with explicit calibration envelopes rather than using raw delayed scores as if they were stable gain itself.

\paragraph{Cohort-escrow law.}
Error debt should be accounted proposal by proposal. Any accounting scheme that nets challenge gains across unrelated cohorts can hide bad proposals behind good ones and therefore loses the anti-masking guarantee of Proposition~\ref{prop:anti_masking}.

\paragraph{Tag-preserving promotion law.}
Promotion from heterodoxy to evidentiary core must preserve contradiction tags whenever the item remains inside the contradiction window. Otherwise promotion can launder away reserve mass and invalidate reserve-based stability floors.

\paragraph{Refresh-before-adopt law.}
If the realized crossing statistic of Theorem~\ref{thm:half_life} reaches the half-margin threshold before the next planned refresh, then aggressive self-edits should be suspended or quarantined until a slower evaluator refresh occurs.

\paragraph{Backlog quarantine law.}
If the proposal rate exceeds verification service, a stable policy must either throttle proposal generation or increase effective challenge throughput. Otherwise the active-region drift of Theorem~\ref{thm:toym_queue} drives persistent congestion, and any deterministic debt ceiling from Theorem~\ref{thm:shadow_debt} must deteriorate through the queue cap or claim-inflow cap that the policy can no longer maintain.

\subsection{Implementation mapping to logged systems}
The theory becomes operational only if the recursive improver emits enough public metadata to estimate the key latent interfaces. The minimal logging unit is a \emph{proposal cohort record}: proposal identifier, parent version, edit type, claimed improvement, evaluator hash before and after the edit, parser-family version set, queue snapshot, sampling probability, realized challenge delay, matured shadow outcome, matched correction, explicit retirement action, and every lane transition touching contradiction-tagged artifacts. The point is not bureaucratic ornament. Without cohort identity and routing metadata, one cannot even test anti-masking, reserve conservation, or delayed-certification claims. For an autonomous agent, the same metadata has direct implementation value: it turns self-modification into a sequence of replayable upgrade cohorts whose audits can be reused, whose failures can be localized, and whose rollback radius can be bounded rather than rediscovered from scratch.

The core estimands are also straightforward, but the estimator must respect the theorem interface rather than a softer average-fit surrogate. In particular, the reference scaffold of Assumption~\ref{ass:gain_matrix} may not be an unobserved Platonic path: it must be instantiated by a replayable baseline built from archived checkpoints, frozen null-policy replays, or matched shadow replays whose construction is itself logged. The gain matrix should therefore be estimated as a conservative upper envelope $\bar K$, not merely a point estimate $\hat K$: first build paired deviations relative to that replayable scaffold, then fit a nonnegative regression of the next-step deviation vector on the current deviation vector, inflate each coefficient by a high-confidence residual envelope, and test $\rho(\bar K)<1$ rather than $\rho(\hat K)<1$. For genuinely open-ended deployments, the more relevant object is the window-summed spillover envelope of Theorem~\ref{thm:incremental_open}; estimate how much a unit of certified productive capability growth pushes evaluator drift, backlog, debt, and replay fragility over each admissible window, then check that the resulting governance spillover remains summable. Delayed-challenge lower bounds should be estimated class/block-conditionally on challenge cohorts using the blocked randomized design of Definition~\ref{def:challenge_design}, replayable doubly robust scores, and confidence sequences or e-processes as in Theorem~\ref{thm:dr_shadow_cert} and Proposition~\ref{prop:chal_radius}. Weighted or doubly robust conformal methods may still be used, but only for auxiliary calibration-envelope estimation or transport-model diagnostics \cite{Tibshirani2019,Yang2024,Jonkers2024,JinCandes2026}; the admission-facing lower certificate is the CS/e-process object, not a generic conformal score. Queue slack is estimated directly from predictable routed load and realized service, while the Red-Queen term $c^{\mathrm{ver}}(K_t^{\mathrm{cap}})$ is estimated from replay time, audit depth, and environment-reconstruction cost as capability grows. Shadow-lane contamination is estimated from lag-window rewrite magnitude $S_t^{(L)}$, parser/environment migration counts, and shadow-baseline drift to operationalize Proposition~\ref{prop:shadow_speed}; shadow staleness is estimated from evaluator/environment drift between refreshes to operationalize Theorem~\ref{thm:two_clock}. Replayability is monitored by the public compatibility score $G_t^{\mathrm{replay}}$, while semantic retention under refreezing is monitored by the endogenous family floor $\underline S_t^{\mathrm{sem}}$ so that Proposition~\ref{prop:replay_floor}, Theorem~\ref{thm:finite_horizon}, and Theorem~\ref{thm:epistemic_singularity} can all be audited. Reserve conservation is tested by checking whether admissible promotion or demotion changes $R_t^{\mathrm{contra}}$ after controlling for genuine entry, expiry, or replay failure, and reserve quality is tested separately through cross-basis robust semantic volume $\bar{\mathcal V}_t^{\mathrm{contra}}$. Attention bypass is tested separately through the heterodoxy routing load $\mathfrak A_t^{\Het}$ and finite-difference ablation tests on contradiction-tagged retrieval gates. Anti-masking is tested by comparing proposal-local escrow debt against any aggregate netting surrogate on downstream rollback, patch, or contradiction-resolution outcomes. Evaluator capture is tested separately through endogenous integrity envelopes, cross-shadow disagreement, audit-family overlap, and disjoint shadow replicas, because debt anti-masking alone does not bound collusive score inflation.

These observables are enough to make the theory falsifiable. If the estimated spectral radius of the coupled gain map stays below one, challenge-lane safety margin stays positive, lower calibrated retained margins stay nonnegative, and contradiction reserve remains conserved under admissible retyping, then the theory predicts bounded drift and bounded escrow in the corresponding operating region. If any of those interfaces fails, the framework predicts a concrete failure mode rather than a vague loss of quality.
\subsection{Relation to adjacent layers}
Proposal--Veto Balance contributes local self-modification stability and error-debt intuition \cite{TakahashiPV2026}. Search Stability contributes finite-context preservation and delayed verification under compression \cite{TakahashiSearch2026}. Benchmark Half-Life contributes validity decay under recursive reuse \cite{TakahashiBH2026}. MTE contributes the observable-only metrology layer \cite{TakahashiMTE2026}. Sovereign Epistemic Commons contributes typed lanes and contradiction-preserving governance intuitions \cite{TakahashiSEC2026}. The present paper uses those layers as adjacent inputs but studies a distinct object: \emph{closed-loop recursive self-improvement stability under endogenous evaluator drift}.

This positioning matters because neighboring layers solve different problems. Metrology asks what counts as credit-bearing progress. Commons governance asks how shared evidence should be admitted, quarantined, or forgotten. The present layer asks when a self-editing loop can remain governable at all once those evaluative interfaces drift endogenously. The contradiction-reserve and routing results are correspondingly narrower than a full commons theory: they concern underprovision, bypass, and delayed challenge exposure inside recursive improvement loops rather than governance of an entire shared epistemic substrate.

\subsection{Experimental plan (plan only)}
This is a theory paper, so the goal is not benchmark decoration but falsifiable operational thresholds. Each hypothesis is written in a way that can be implemented from ordinary proposal, queue, and challenge logs.

\paragraph{Protocol P1: windowed gain estimation with and without a reference scaffold.} In migration phases with a meaningful co-moving scaffold, maintain a rolling reference trajectory, estimate nonnegative one-step gain envelopes $\widehat K_t$, and then estimate a conservative window envelope for products of length $W$. The primary falsification statistic is not the spectral radius of a best-fit one-step matrix but whether the estimated window product is dominated by some public envelope $\bar K$ with $\rho(\bar K)<1$. In genuinely open-ended regimes, switch to the incremental protocol of Theorem~\ref{thm:incremental_open}: estimate how accepted capability increments push the governance variables over each window, and test whether the resulting spillover envelope remains summable. The theory predicts that runs satisfying either certified interface obey the relevant governance ceilings, while runs violating it should exhibit more frequent drift blow-ups or debt excursions.

\paragraph{Protocol P2: class-conditional one-sided shadow certification under adaptive routing and two clocks.} On challenged cohorts, publish a confidence-budget manifest $\{\alpha_{h,\mathcal C,\mathfrak B}^{\mathrm{chal}},\alpha_{h,\mathcal C,\mathfrak B}^{\mathrm{tr}},n_{\min}(h,\mathcal C,\mathfrak B)\}$, build block-frozen class/block-conditional doubly robust scores, and compute replayable lower confidence sequences
\[
\underline M^{\mathrm{chal}}_{t,n}(h;\mathcal C_t,\mathfrak B_t)
=
\widehat{\mathcal M}_{t,n}^{\mathrm{chal}}(h;\mathcal C_t,\mathfrak B_t)
-\operatorname{rad}_{t,n}^{\mathrm{chal}}(h,\alpha_{h,\mathcal C_t,\mathfrak B_t}^{\mathrm{chal}};\mathcal C_t,\mathfrak B_t)
-\delta_h^{\mathrm{design}}(\mathcal C_t,\mathfrak B_t).
\]
Separately estimate a Tier-2 transport-calibration family $(a_h(\mathcal C_t,\mathfrak B_t),b_h(\mathcal C_t,\mathfrak B_t),\delta_h^{\mathrm{tr}}(\mathcal C_t,\mathfrak B_t))$ from matched long-horizon replay or audited transport certificates, with the transport residual envelope itself certified under the budget $\alpha_{h,\mathcal C_t,\mathfrak B_t}^{\mathrm{tr}}$. The contamination budget $\widehat\Delta_t^{\Shadow}+L_{\Shadow}\widehat S_t^{(L)}$, the staleness budget $\widehat\Omega_h^{\mathrm{stale}}(f_{\Shadow})$, and the interaction budget $\widehat\Gamma_{t,h}^{\mathrm{int}}$ are replayable proof-carrying upper envelopes under separate logged certificates and therefore do not consume the statistical budget reserved for challenge or transport confidence sequences. Compute the operational certified lower bound
\[
\widehat L_t
\Defeq
\sum_{h=1}^{H}\beta^{h-1}\Bigl[
\begin{aligned}
& a_h(\mathcal C_t,\mathfrak B_t)\,\underline M^{\mathrm{chal}}_{t,n}(h;\mathcal C_t,\mathfrak B_t)
-b_h(\mathcal C_t,\mathfrak B_t)\\
&\quad -\delta_h^{\mathrm{tr}}(\mathcal C_t,\mathfrak B_t)
-\widehat\Delta_t^{\Shadow}
-L_{\Shadow}\widehat S_t^{(L)}\\
&\quad -\widehat\Omega_h^{\mathrm{stale}}(f_{\Shadow})
-\widehat\Gamma_{t,h}^{\mathrm{int}}
\end{aligned}
\Bigr].
\]
If the current pair $(\mathcal C_t,\mathfrak B_t)$ has insufficient matured support, or fewer than $n_{\min}(h,\mathcal C_t,\mathfrak B_t)$ usable matured cohorts at the required horizon, Interface A is fail-closed and the proposal must be routed to challenge, proof escalation, or sandbox. The theory predicts that positive values of $\widehat L_t$ correlate with future slow-lane benefit, whereas leaving the viability band of Theorem~\ref{thm:two_clock} or violating the speed limit from Proposition~\ref{prop:shadow_speed} should sharply degrade sign reliability.

\paragraph{Protocol P3: queue-slack and repair-throughput intervention.} Vary proposal throttling, challenge throughput, and explicit repair throughput so that both the predictable load ratio $\Lambda_t/\mu$ and the strict-conservation balance $\bar\lambda_L+\lambda_q\bar q-(1-\eps_{\mathrm{rep}})\mu^{\mathrm{rep}}$ cross zero. The theory predicts two distinct thresholds: one for verification congestion and one for repair throughput. Below both thresholds, backlog and debt are controlled; above the verification threshold backlog grows, and above the repair threshold strict-conservation debt drifts upward even if bugs are found quickly.

\paragraph{Protocol P3b: Red-Queen verification and governance audit.} Estimate capability-dependent verification complexity $c^{\mathrm{ver}}(K_t^{\mathrm{cap}})$ from replay time, proof-check time, environment rebuild cost, and challenge-depth cost at each capability level. Separately estimate governance recomputation cost $c_t^{\mathrm{gov}}$ for semantic-volume, routing-sensitivity, replay-floor, and integrity-envelope checks, together with certificate-verification cost $\widetilde c_t^{\mathrm{gov}}$ and fallback-audit fraction $\sigma_t^{\mathrm{gov}}$. Compare predictable load $\Lambda_t$ with effective service $P_t^{\mathrm{verif}}/(c^{\mathrm{ver}}(K_t^{\mathrm{cap}})+\widehat c_t^{\mathrm{gov}}(u_t,\pi_t))$. The theory predicts sharp queue deterioration once $\Lambda_t(c^{\mathrm{ver}}(K_t^{\mathrm{cap}})+\widehat c_t^{\mathrm{gov}}(u_t,\pi_t))$ persistently exceeds physical verification budget $P_t^{\mathrm{verif}}$.

\paragraph{Protocol P3c: physical resource-escrow audit.} Log the total physical budget $P_t^{\mathrm{tot}}$, the reserved verification share $P_t^{\mathrm{verif}}$, the ratio $\rho_t^{\mathrm{verif}}$, and the attestation status of the allocator controlling this split. The theory predicts that runs without a protected lower floor on $\rho_t^{\mathrm{verif}}$ can starve replay and shadow service even when software-side acceptance predicates look healthy, whereas runs satisfying the resource-escrow invariant should keep the verification clocks alive whenever workload slack remains positive.

\paragraph{Protocol P4: anti-masking ablation.} Compare cohort-escrow accounting against aggregate netting on mixed-quality proposal streams. The key endpoint is downstream rollback or contradiction-resolution need. The theory predicts that proposal-local escrow debt has higher precision for harmful proposals because unrelated good cohorts cannot cancel them.

\paragraph{Protocol P5: reserve-conservation, robust-semantic-volume, routing, replay-floor, and semantic-horizon audit.} Re-run the same admissible promotion, demotion, parser-update, and environment-update schedule under two bookkeeping policies: one that preserves tags, replay floors, robust-semantic-volume thresholds, and heterodoxy routing load via refreezing and forced-attention checks, and one that does not. In the first policy, log replay compatibility $G_t^{\mathrm{replay}}$, endogenous semantic floor $\underline S_t^{\mathrm{sem}}$, contradiction mass $R_t^{\mathrm{contra}}$, basis-specific semantic volumes $\mathcal V_t^{(k)}$, robust semantic volume $\bar{\mathcal V}_t^{\mathrm{contra}}$, and heterodoxy routing load $\mathfrak A_t^{\Het}$. The theory predicts invariance of $R_t^{\mathrm{contra}}$ under the tag-preserving policy up to genuine entry and expiry noise, bounded loss of $\bar{\mathcal V}_t^{\mathrm{contra}}$ under spam rejection and representation capture by any strict subset of bases, maintenance of the replay floor on retained audit items, but gradual decay of effective evidentiary weight consistent with Theorem~\ref{thm:finite_horizon}. A successful audit should therefore show replay preservation, basis-robust semantic diversity, and nonzero routing sensitivity rather than mere storage.

\paragraph{Minimal logging contract.} At minimum, \emph{raw proposal logs} should include proposal identifier and time, parent version, evaluator hash before and after edit, parser-family version set, routed lane, declared writable scope, declared dependency-closure size, queue snapshot, replay-compatibility score, contradiction-tag status, provenance depth, allocator lineage identifier, migration-attestation transcript hash, reserved verification ratio before and after migration, and lane-transition record. \emph{Derived governance metrics} should include claim mass, systemic risk bond, challenge sampling probability, realized challenge delay, matured shadow outcome, matched retirement credit, secondary-liability charge, explicit retirement action, contradiction fingerprints for each encoder family, robust-semantic-volume marginal gain or loss, heterodoxy routing sensitivity, dropped and regenerated contradiction mass, and dropped and regenerated robust semantic volume. \emph{Audit artifacts} should include constitutional proof-language tag, proof length, proof-check runtime and memory, checker identifier, PMU-measured wall-clock and energy, proof-preemption flag and reason, scrubber family identifier, scrubber randomness attestation or commitment, attacker-conditioned entropy estimate for scrubber seeds, empirical scrubber-capacity estimate, and garbage-collection event identifier. The point is to separate what must be logged directly from what may be recomputed offline and from what is itself an auditable certificate.

\paragraph{Primary metrics.} Estimated window-product spectral radius, yardstick drift ceiling violations, certified lower stable-gain bound, shadow contamination budget, singularity margins $\Xi^{\Shadow}-A^{\mathrm{adv}}$ and $\Xi^{\Replay}-A^{\mathrm{adv}}$, Red-Queen load ratio $\Lambda_t(c^{\mathrm{ver}}(K_t^{\mathrm{cap}})+\widehat c_t^{\mathrm{gov}}(u_t,\pi_t))/P_t^{\mathrm{verif}}$, queue mean and tail, proposal-local peak debt, systemic-risk-bond excursions, causal-intervention-capacity excursions $\Gamma_t^{\mathrm{cic}}(u_t)$, replay-floor violations, contradiction reserve conservation error, robust semantic-volume floor violations, heterodoxy routing-load violations, resource-escrow ratio violations $\rho_t^{\mathrm{verif}}<\rho_{\min}$, and long-horizon rollback rate.

\paragraph{Decision rule mapping.} A practical deployment policy suggested by the theory is to distinguish four interfaces explicitly and to expose a non-binary action alphabet. Interface 0 is base admissibility: replay success, declared cost budget, side-effect well-definedness, mandatory challenge routing for ruler edits, and summary typing or context-preservation clauses must all hold. Interface A is certified stable gain: the calibrated lower stable-gain bound must be nonnegative after subtracting shadow contamination, staleness, and endogenous integrity budgets. Interface B is governance debt and blast-radius safety: proposal-local or bundle-level peak debt, including the systemic risk bond at issuance and any interaction debt for coupled edits, must stay below threshold. Interface C is governance viability: challenge-lane safety margin, Red-Queen verification slack, contradiction reserve, robust semantic volume, heterodoxy routing load, replay floor, semantic-retention floor, proof-budget envelope, and resource-escrow clauses must all remain nonnegative on the admitted horizon. In operational terms the action set should be
\[
\{\textsf{fast-accept},\textsf{defer-challenge},\textsf{escalate-proof},\textsf{sandbox},\textsf{promote},\textsf{reject}\}.
\]
Fast acceptance requires the master admissibility rule of Proposition~\ref{prop:master_admissibility}. Failure of fast admission does \emph{not} by itself imply permanent rejection: evaluator edits or proof-heavy proposals may instead be deferred to protected challenge, escalated to a richer proof lane, or sandboxed under tighter scope until they satisfy Definition~\ref{def:promotion_eligible}. A practical autonomous agent can implement this by maintaining replayable estimates of both $\Phi_t$ and $\Phi_t^{\mathrm{net}}$, by running the selector-invariant review contract together with bounded-wait completion from Definition~\ref{def:promotion_eligible}, and by exposing the protected resource counters needed for Definition~\ref{def:service_scaling} and Definition~\ref{def:net_growth_metric}. The resulting deployment logic is intentionally stricter than ``positive claim'' because the paper's whole point is that claims alone are cheap theater, whereas bounded-delay delayed certification and resource-coupled service can still be growth-promoting.
\section{Limitations}
This paper is intentionally narrow. It is not a full theory of semantic truth, parser stability, ontology transport, constitutional self-revision, or shared epistemic commons governance, though it does make the replay-compatibility and gauge-transport interfaces explicit. It does not prove indefinite self-improvement, universal safety, or absolute sovereignty. It also does not claim that the canonical admissibility instance is uniquely correct. The contribution is instead a first-principles interface theory for one particular layer: recursive self-improvement when the evaluator itself drifts endogenously. Its strongest positive claims are explicitly attached to the observable comparison contracts introduced in Definitions~\ref{def:gp_class} and \ref{def:gp_net_class}; outside those contracts the paper does not assert realized growth promotion.

The framework is model-architecture-agnostic in the limited sense that it does not assume a particular optimizer, neural architecture, or training rule. It is not instrumentation-free. Replayable comparison, challenge lanes, proof lanes, protected-service counters, and resource escrows are substantive structural requirements, not cosmetic implementation details. Likewise full admissibility is certificate-based rather than oracle-based: the theory distinguishes the latent quantity of oracle stable gain from the replayable lower certificates that are actually allowed to govern admission.

Several objects are deliberately handled as conservative public envelopes rather than as exact latent quantities. In particular, the audit-diversity reserves $\Xi_t^{\Shadow},\Xi_t^{\Replay}$, attack-reach envelope $A_t^{\mathrm{adv}}$, intervention-capacity envelope $\Gamma_t^{\mathrm{cic}}(u_t)$, and endogenous semantic floor $\underline S_t^{\mathrm{sem}}$ must still be instantiated by concrete proxy constructions in any deployment. The theorems say what follows once such envelopes are replayably published; they do not eliminate the engineering burden of constructing them conservatively.

The semantic-volume layer also inherits representation risk. Cross-basis robust semantic volume is designed to blunt single-basis capture, but its usefulness still depends on at least one sufficiently faithful uncaptured basis and on replayable transport certificates when the basis family changes. Likewise the hidden-state no-go theorem is an identifiability impossibility, not a claim that practical challenge design is futile; it says only that some ambiguities cannot be broken from the fast public lane alone. The public envelopes $\Gamma_t^{\mathrm{cic}}$, $A_t^{\mathrm{adv}}$, $\Xi_t^{\Shadow}$, $\Xi_t^{\Replay}$, and $\underline S_t^{\mathrm{sem}}$ remain conservative wrappers whose operational sharpness depends on the published proxy construction, but the gate discipline of Definition~\ref{def:proxy_tiers} now restricts fast deployment to Tier-1 and Tier-2 quantities so that heuristic Tier-3 signals cannot by themselves manufacture admissibility. The quantitative shadow-identifiability proposition is deliberately stronger than what arbitrary deployments can guarantee: it assumes positive challenged-subsequence exposure, bounded delay, and martingale-type concentration on matured challenged cohorts. The new challenge-service proposition, bounded-exposure entry control, cap-invariance proposition, and selection-liveness theorem narrow the gap between queue theory and identifiability, but practical deployments still need explicit scheduler, selector, bounded-wait completion, admission-control, surrogate-construction, and instrumentation choices to instantiate their constants. Full admissibility is also intentionally narrower than mere lane admission: delayed challenge, richer-proof, and sandbox actions are modeled as nondeployment states and do not enter the productive frontier until the delayed certificates and review actions required for actual deployment have been completed. Finally, the state variable $\Pi_t$ allows policy manifests to drift endogenously, but the ambient transition-law class is still fixed up to the replayable parameterizations declared by the model; unrestricted self-rewriting of the law class itself would require an additional meta-dynamics layer.

Finally, proof-carrying governance and selective recomputation can themselves become bottlenecks if deployed without explicit lane design. Requiring public certificates, protected audit families, resource escrow, and bounded-delay challenge positivity may delay fast admission for many edits that are in fact good, and may produce false negatives when good proposals are proof-heavy, expensive to replay, disadvantaged by hardware asymmetry, or stabilized by sign-canceling dynamics that lie outside the conservative small-gain wrapper. The framework therefore needs the richer action alphabet formalized above, together with review-locked deployment selection, bounded-delay proof and promotion service, cap-maintaining intake together with bounded-exposure entry control, service scaling with success, replayable comparison-sound surrogate selection, and the governance-investment interface: challenge routing, proof escalation, sandboxing, and promotive governance edits are what prevent bounded verification discipline from degenerating into blanket anti-growth rejection when the ruler can rewrite itself. The choice among challenge, proof, sandbox, and immediate deployment is itself a separate routing-control problem; the present paper provides admissibility and liveness contracts for those lanes, not an optimal global router. A full welfare analysis of this false-negative trade-off is outside the present paper.

\section{Conclusion}
Recursive self-improvement becomes substantially harder once the ruler moves with the object being improved. Benchmarks decay, parser registries drift, challenge queues backlog, memory substrates are rewritten, and even the proof language used to certify safety can become a bottleneck. The right abstraction is therefore not one-step optimization but coupled admission under moving evaluative interfaces. This paper formalizes that layer as a first-principles interface theory, model-architecture-neutral but infrastructure- and instrumentation-explicit, conditional on published replayable interfaces, and sharpens the separation between four interfaces that are easy to conflate: base admissibility, certified stable gain, debt-bounded stability, and governance safety, whose conjunction yields full admissibility. It also separates oracle stable gain from the replayable lower certificates that alone are allowed to enter admission.

Several consequences follow. Window-contractive gain envelopes bound deviation from a co-moving scaffold and yield explicit yardstick-drift ceilings. Observable pseudometrics together with replayable gauge transport replace raw parameter norms in the open-ended theorem, so bounded governability no longer depends on coordinate choice or on silently changing the ruler. Delayed shadow lanes become certifying objects only when contamination, staleness, challenge coverage, and class-conditional transport calibration to the slow utility surrogate are all controlled. Contradiction must survive not merely as stored mass but as basis-robust, routed, inspectable disagreement with persistent obligations and finite-budget protected family exposure. Debt must include structural blast radius and bundled interaction debt or globally dangerous rewrites can masquerade as many small local wins. Forgetting must be compensated by regenerated contradiction mass or the system can legally burn down its own audit memory while remaining superficially compliant.

The paper's positive scientific claim is therefore conditional rather than universal: it is a first-principles sufficient-condition \emph{interface theory} of growth promotion on explicit observable comparison contracts, with admissibility-operator monotonicity as the core theorem and realized-path comparison only as an explicit transition-class corollary rather than the primary source of novelty. Stable recursive improvement requires disagreement to survive as a live governance input, audit diversity to stay ahead of attack reach, verification throughput and protected service lanes to scale with verification complexity and proposal throughput, proof-heavy proposals to be handled by proof-carrying escalation rather than left to an unmetered bottleneck, and replacement of stale evidence to be paid for under the current ruler with a governance-burden accounting that lower-bounds real protected resource use. Under those observable conditions the theory yields falsifiable interfaces rather than rhetoric. The fixed-threshold frontier theorem makes the positive content explicit: stronger replay, challenge, promotion, proof, reserve, and escrow slack expands the set of productive upgrades that can be admitted without destabilizing the loop and without appealing to deregulation. The frontier-expansion theorem then shows that some edits with zero immediate productive increment are still promotive because they enlarge future admissible frontiers under the same thresholds. The selector-invariant delayed-deployment theorem complements this by showing that delayed challenge need not be a bureaucratic graveyard: once a cohort acquires a positive delayed certificate, silent exposure delay, target-moving, and unreviewed rejection are all blocked by bounded-exposure entry and selector-invariant review, and bounded-delay deployment follows once the separate bounded-wait completion clause is added unless an audited certificate validates genuine ineligibility. The richer-proof-lane theorem extends the same logic to proof-heavy proposals beyond the current fast proof regime by tying bounded deployability to proof-carrying escalation, total replayable checkers, thermodynamic escrow, and cap-maintaining proof-lane intake rather than to wishful thinking. The resource-coupled service-scaling law then shows that success need not collapse those lanes provided service is backed by published protected budgets and replayable intake control. Finally, the surrogate-selector theorem shows that these comparisons do not require exact online solution of the one-step frontier maximization problem; replayable witness-dominating lower envelopes together with admissibility-operator monotonicity suffice, and the exact comparator case appears as the zero-distortion special case of the interval-comparator framework. Finally, the realized-path and resource-priced comparison corollaries state the stronger claim needed by an autonomous agent designer only in their correct explicit scope: they are transition-class corollaries built on the operator-monotonicity layer, and on the observable comparison contracts of Definitions~\ref{def:gp_class} and \ref{def:gp_net_class}, for explicit regime-policy pairs whose plus selector dominates replayably certified comparison witnesses, governance investments improve not only static admissibility sets but the realized raw and net productive trajectories themselves. When the interfaces fail, the failure mode is not mysterious: the loop loses identifiability, loses audit mass, loses deployment liveness, loses protected budget, or loses throughput before it loses rhetoric. For autonomous agents, that is the practical reason to implement the framework: it converts self-editing from an exercise in local score maximization into a reproducible upgrade protocol whose successful edits stay portable across future versions, whose governance improvements are certified investments in larger future growth frontiers even after accounting for governance cost, and whose failed edits leave auditable debris instead of mythology.


\appendix
\section*{In-file Machine Appendix}
{\tiny
\begin{verbatim}
YAML_APPENDIX_BEGIN
canonical_source: "This appendix is the sole canonical machine-readable summary embedded in the source file."
anchor_index:
  def:challenge_design: "blocked randomized challenge design with class/block target and fail-closed support rule"
  def:chal_defects: "design-defect envelopes for positivity, ignorability, support mismatch, and clipping bias"
  prop:chal_defect_comp: "common-scale additive composition of challenge-design defects"
  def:interface_operators: "interface-indexed single-gate admissibility operators"
  def:post_action_frontier: "action-induced next-step regime and well-typed post-action frontier scores"
  thm:dr_shadow_cert: "finite-sample lower certificate for class/block challenge target"
  prop:chal_radius: "empirical-Bernstein executable instantiation for ordered matured sample"
  def:transport_calibration: "class/block transport calibration from challenge margin to slow utility"
  prop:transport_tier2: "matched-replay certificate implies Tier-2 transport family"
  prop:transport_radius: "upper CS for transport residual"
  prop:challenge_to_shadow: "operational and optional oracle retained-margin certificates"
  def:feasible_order: "formal feasible monotonicity order on fixed-threshold regimes"
  thm:operator_monotone: "fixed-threshold admissibility-operator monotonicity"
  thm:familywise_exposure: "finite-budget protected reserve exposure with challenge coupling"
  ex:toy_ruleredit: "hard-core toy instantiating class/block challenge and one-step monotonicity"
  prop:toy_ruleredit_mono: "toy one-step monotonicity under feasible order"

theorem_dependency:
  thm:dr_shadow_cert:
    depends_on: [def:challenge_design, def:chal_defects, prop:chal_defect_comp, def:confidence_manifest]
    outputs: [underline_M_chal, delta_design]
  prop:chal_defect_comp:
    depends_on: [def:chal_defects]
    outputs: [delta_design_common_scale]
  prop:chal_radius:
    depends_on: [thm:dr_shadow_cert]
    outputs: [chal_radius_executable, sample_filtration]
  prop:transport_tier2:
    depends_on: [def:transport_calibration, def:proxy_tiers]
    outputs: [tier2_transport_family]
  prop:transport_radius:
    depends_on: [def:transport_calibration, prop:transport_tier2]
    outputs: [delta_tr_executable]
  prop:challenge_to_shadow:
    depends_on: [thm:dr_shadow_cert, def:transport_calibration, def:class_homogeneity, prop:transport_radius]
    outputs: [underline_M_op, underline_M_oracle_optional]
  prop:shadow_bridge:
    depends_on: [def:full_admissibility, prop:challenge_to_shadow]
    outputs: [lower_oracle_stable_gain_bound]
  cor:shadow_cert:
    depends_on: [prop:shadow_bridge, prop:challenge_to_shadow, def:cohort_escrow]
    outputs: [interface_A_operational_sufficiency, oracle_upgrade_optional]
  prop:mono_base:
    depends_on: [def:frontier, def:interface_operators, def:feasible_order]
  prop:mono_gain:
    depends_on: [def:frontier, def:interface_operators, def:proxy_tiers, prop:challenge_to_shadow, def:feasible_order]
  prop:mono_debt:
    depends_on: [def:frontier, def:interface_operators, def:cohort_escrow, def:feasible_order]
  prop:mono_gov:
    depends_on: [def:frontier, def:interface_operators, def:proxy_tiers, def:feasible_order]
  thm:operator_monotone:
    depends_on: [def:interface_operators, prop:mono_base, prop:mono_gain, prop:mono_debt, prop:mono_gov]
  cor:realized_growth:
    depends_on: [thm:operator_monotone, def:gp_class, def:post_action_frontier, def:witness_selector, cor:service_scaling]
    outputs: [pathwise_raw_dominance]
  cor:net_realized_growth:
    depends_on: [thm:operator_monotone, def:gp_net_class, def:post_action_frontier, cor:service_scaling]
    outputs: [pathwise_net_dominance]
  thm:familywise_exposure:
    depends_on: [def:forced_attention]
  thm:shadow_debt:
    depends_on: [def:cohort_escrow, ex:toy_multi]
  prop:toy_ruleredit_mono:
    depends_on: [ex:toy_ruleredit, def:feasible_order, def:challenge_design, def:confidence_manifest]

object_lineage:
  M_t_oracle_h:
    meaning: "cohort-specific oracle retained margin"
    defined_at: "def:retained_margin"
  M_chal_class_block_h:
    meaning: "class/block retained challenge margin on matured issued support"
    defined_at: "def:challenge_design"
  M_t_op_h:
    meaning: "operational class/block retained-margin target"
    defined_at: "def:transport_calibration"
  underline_M_chal:
    meaning: "finite-sample lower confidence sequence for class/block challenge target"
    defined_at: "thm:dr_shadow_cert"
  underline_M_op:
    meaning: "admission-facing lower certificate for operational retained-margin target"
    defined_at: "prop:challenge_to_shadow"
  underline_M_oracle_optional:
    meaning: "oracle-margin lower certificate requiring class-homogeneity envelope"
    defined_at: "prop:challenge_to_shadow"
  delta_tr:
    meaning: "transport residual envelope certified from matched replay on the same matured support"
    defined_at: "prop:transport_radius"
  delta_class:
    meaning: "gap from class/block operational target to cohort-specific oracle margin"
    defined_at: "def:class_homogeneity"
  delta_cmp:
    meaning: "witnesswise comparator distortion handled only at stable-gain bridge"
    defined_at: "def:transport_calibration"

protocol_map:
  P2:
    inputs: [underline_M_chal, delta_tr, shadow_contamination_cert, staleness_cert, interaction_cert]
    statistical_budget: [alpha_chal, alpha_tr]
    deterministic_budgets: [shadow_contamination_cert, staleness_cert, interaction_cert]
    output: [L_hat_t]

policy_fail_closed:
  empty_or_small_class_support: "Interface A inconclusive; route to challenge, proof, or sandbox when matured support is below n_min(h,C,B)."
  unsupported_transport_family: "Tier-2 transport calibration unavailable; no fast gain certificate."
  comparator_incompatible: "No stable-gain certificate under current witness family."

notation_addendum:
  published: "Declared in the active manifest and binding for audit."
  replayable: "Recoverable by deterministic replay from declared artifacts and kernels."
  observable: "Either directly logged or replay-reconstructible from published artifacts."
  operational_target: "Admission-facing class/block target, distinct from cohort-specific oracle target."
  challenge_observed_score: "Y_obs = A * O * Y_ret on the matured issued support."
  deterministic_budget: "A proof-carrying or replay-verifiable upper envelope that does not consume statistical alpha."
YAML_APPENDIX_END
\end{verbatim}
}

\begin{thebibliography}{99}

\bibitem{DatasetShift2008}
J.~Qui\~nonero-Candela, M.~Sugiyama, A.~Schwaighofer, and N.~D. Lawrence, editors.
\newblock \emph{Dataset Shift in Machine Learning}.
\newblock MIT Press, 2008. DOI: 10.7551/mitpress/9780262170055.001.0001.

\bibitem{ManheimGarrabrant2018}
D.~Manheim and S.~Garrabrant.
\newblock Categorizing Variants of Goodhart's Law.
\newblock \emph{arXiv preprint arXiv:1803.04585}, 2018.

\bibitem{HespanhaMorse1999ADT}
J.~P. Hespanha and A.~S. Morse.
\newblock Stability of switched systems with average dwell-time.
\newblock In \emph{Proceedings of the 38th IEEE Conference on Decision and Control}, volume 3, pages 2655--2660, 1999. DOI: 10.1109/CDC.1999.831330.

\bibitem{SelfEvolvingSurvey2025}
H.-A. Gao, J. Geng, W. Hua, M. Hu, X. Juan, H. Liu, S. Liu, J. Qiu, X. Qi, Y. Wu, H. Wang, H. Xiao, Y. Zhou, S. Zhang, J. Zhang, J. Xiang, Y. Fang, Q. Zhao, D. Liu, Q. Ren, C. Qian, Z. Wang, M. Hu, H. Wang, Q. Wu, H. Ji, and M. Wang.
\newblock A Survey of Self-Evolving Agents: On Path to Artificial Super Intelligence.
\newblock \emph{arXiv preprint arXiv:2507.21046}, 2025.

\bibitem{HorvitzThompson1952}
D.~G. Horvitz and D.~J. Thompson.
\newblock A Generalization of Sampling Without Replacement from a Finite Universe.
\newblock \emph{Journal of the American Statistical Association}, 47(260):663--685, 1952. DOI: 10.1080/01621459.1952.10483446.

\bibitem{Schmidhuber2007}
J.~Schmidhuber.
\newblock G\"odel Machines: Fully Self-Referential Optimal Universal Self-Improvers.
\newblock In B.~Goertzel, C.~Pennachin, and M.~Geisweiller, editors, \emph{Artificial General Intelligence}, pages 199--226. Springer, 2007.

\bibitem{Garrabrant2016}
S.~Garrabrant, T.~Benson-Tilsen, A.~Critch, N.~Soares, and J.~Taylor.
\newblock Logical Induction.
\newblock \emph{arXiv preprint arXiv:1609.03543}, 2016.

\bibitem{Freedman1975}
D.~A. Freedman.
\newblock On Tail Probabilities for Martingales.
\newblock \emph{The Annals of Probability}, 3(1):100--118, 1975. DOI: 10.1214/aop/1176996452.

\bibitem{Neely2010}
M.~J. Neely.
\newblock \emph{Stochastic Network Optimization with Application to Communication and Queueing Systems}.
\newblock Synthesis Lectures on Communication Networks, 3(1):1--211, Morgan \& Claypool, 2010. DOI: 10.2200/S00271ED1V01Y201006CNT007.

\bibitem{Tibshirani2019}
R.~J. Tibshirani, R.~F. Barber, E.~J. Cand\`es, and A.~Ramdas.
\newblock Conformal Prediction Under Covariate Shift.
\newblock In \emph{Advances in Neural Information Processing Systems 32 (NeurIPS 2019)}, 2019.

\bibitem{Yang2024}
Y.~Yang, A.~K. Kuchibhotla, and E.~Tchetgen Tchetgen.
\newblock Doubly Robust Calibration of Prediction Sets under Covariate Shift.
\newblock \emph{Journal of the Royal Statistical Society Series B: Statistical Methodology}, 86(4):943--965, 2024. DOI: 10.1093/jrsssb/qkae009.

\bibitem{HowardRamdasMcAuliffeSekhon2021}
S.~R. Howard, A.~Ramdas, J.~McAuliffe, and J.~Sekhon.
\newblock Time-uniform, nonparametric, nonasymptotic confidence sequences.
\newblock \emph{The Annals of Statistics}, 49(2):1055--1080, 2021. DOI: 10.1214/20-AOS1991.

\bibitem{RamdasGrunwaldVovkShafer2023}
A.~Ramdas, P.~Gr\"unwald, V.~Vovk, and G.~Shafer.
\newblock Game-Theoretic Statistics and Safe Anytime-Valid Inference.
\newblock \emph{Statistical Science}, 38(4):576--597, 2023. DOI: 10.1214/23-STS894.

\bibitem{WassermanRamdasBalakrishnan2020}
L.~Wasserman, A.~Ramdas, and S.~Balakrishnan.
\newblock Universal inference.
\newblock \emph{Proceedings of the National Academy of Sciences}, 117(29):16880--16890, 2020. DOI: 10.1073/pnas.1922664117.

\bibitem{ZhanHadadHirshbergAthey2021}
R.~Zhan, V.~Hadad, D.~A. Hirshberg, and S.~Athey.
\newblock Off-Policy Evaluation via Adaptive Weighting with Data from Contextual Bandits.
\newblock \emph{arXiv preprint arXiv:2106.02029}, 2021.

\bibitem{JinCandes2026}
Y.~Jin and E.~J. Cand\`es.
\newblock Model-free selective inference under covariate shift via weighted conformal p-values.
\newblock \emph{Biometrika}, 113(1):asaf066, 2026. DOI: 10.1093/biomet/asaf066.


\bibitem{JiangTeelPraly1994}
Z.-P. Jiang, A.~R. Teel, and L. Praly.
\newblock Small-Gain Theorem for ISS Systems and Applications.
\newblock \emph{Mathematics of Control, Signals, and Systems}, 7(2):95--120, 1994. DOI: 10.1007/BF01211469.

\bibitem{JiangWang2001}
Z.-P. Jiang and Y. Wang.
\newblock Input-to-state stability for discrete-time nonlinear systems.
\newblock \emph{Automatica}, 37(6):857--869, 2001. DOI: 10.1016/S0005-1098(01)00028-0.

\bibitem{Jonkers2024}
J. Jonkers, G. Van Wallendael, L. Duchateau, and S. Van Hoecke.
\newblock Conformal Predictive Systems Under Covariate Shift.
\newblock In \emph{Proceedings of the Thirteenth Symposium on Conformal and Probabilistic Prediction with Applications}, PMLR 230:406--423, 2024.


\bibitem{LiberzonMorse1999}
D.~Liberzon and A.~S. Morse.
\newblock Basic Problems in Stability and Design of Switched Systems.
\newblock \emph{IEEE Control Systems}, 19(5):59--70, 1999. DOI: 10.1109/37.793443.

\bibitem{ShortenWirth2007}
R.~Shorten, F.~Wirth, O.~Mason, K.~Wulff, and C.~King.
\newblock Stability Criteria for Switched and Hybrid Systems.
\newblock \emph{SIAM Review}, 49(4):545--592, 2007. DOI: 10.1137/05063516X.

\bibitem{LinAntsaklis2009}
H.~Lin and P.~J. Antsaklis.
\newblock Stability and Stabilizability of Switched Linear Systems: A Survey of Recent Results.
\newblock \emph{IEEE Transactions on Automatic Control}, 54(2):308--322, 2009. DOI: 10.1109/TAC.2008.2012009.

\bibitem{Yin2011IncorrectFixes}
Z.~Yin, D.~Yuan, Y.~Zhou, S.~Pasupathy, and L.~Bairavasundaram.
\newblock How Do Fixes Become Bugs? A Comprehensive Characteristic Study on Incorrect Fixes in Commercial and Open Source Operating Systems.
\newblock In \emph{Proceedings of the 19th ACM SIGSOFT Symposium and the 13th European Conference on Foundations of Software Engineering (ESEC/FSE 2011)}, pages 26--36, 2011. DOI: 10.1145/2025113.2025121.

\bibitem{MaesBermejo2024}
M.~Maes-Bermejo, A.~Serebrenik, M.~Gallego, F.~Gort\'azar, and J.~M. Gonz\'alez-Barahona.
\newblock Hunting bugs: Towards an automated approach to identifying which change caused a bug through regression testing.
\newblock \emph{Empirical Software Engineering}, 29:66, 2024. DOI: 10.1007/s10664-024-10479-z.

\bibitem{Matthews2010Preservation}
B.~Matthews, A.~Shaon, J.~Bicarregui, and C.~Jones.
\newblock A Framework for Software Preservation.
\newblock \emph{International Journal of Digital Curation}, 5(1):91--105, 2010. DOI: 10.2218/ijdc.v5i1.145.

\bibitem{Butler2023RB}
S.~Butler, J.~Gamalielsson, B.~Lundell, C.~Brax, A.~Mattsson, T.~Gustavsson, J.~Feist, B.~Kvarnstr\"om, and E.~L\"onroth.
\newblock On business adoption and use of reproducible builds for open and closed source software.
\newblock \emph{Software Quality Journal}, 31:687--719, 2023. DOI: 10.1007/s11219-022-09607-z.

\bibitem{Malka2024SpaceTime}
J.~Malka, S.~Zacchiroli, and T.~Zimmermann.
\newblock Reproducibility of Build Environments through Space and Time.
\newblock \emph{arXiv preprint arXiv:2402.00424}, 2024.

\bibitem{Perkel2020TenYear}
J.~M. Perkel.
\newblock Challenge to scientists: does your ten-year-old code still run?
\newblock \emph{Nature}, 584(7822):656--658, 2020. DOI: 10.1038/d41586-020-02462-7.

\bibitem{TakahashiBH2026}
K.~Takahashi.
\newblock AI Benchmark Half-Life in Recursive Corpora: A Theory of Validity Decay under Semantic Leakage and Regeneration.
\newblock Zenodo, 2026. DOI: 10.5281/zenodo.18954286.

\bibitem{TakahashiSearch2026}
K.~Takahashi.
\newblock Search Stability under Finite Context: A Minimal Theory of Adequacy Preservation, Compression, and Reset in Long-Running Agents.
\newblock Zenodo, 2026. DOI: 10.5281/zenodo.18905242.

\bibitem{TakahashiPV2026}
K.~Takahashi.
\newblock Proposal-Veto Balance for Observable-Only Autonomous Intelligence: Stability Thresholds, Identifiability Limits, and Commit-Window Effects.
\newblock Zenodo, 2026. DOI: 10.5281/zenodo.18883290.

\bibitem{TakahashiMTE2026}
K.~Takahashi.
\newblock Metrology-Theoretic Epistemics Engine (MTE): Observable-Only Metrology for Long-Horizon Autonomous Intelligence.
\newblock Zenodo, 2026. DOI: 10.5281/zenodo.18845340.


\bibitem{TakahashiSEC2026}
K.~Takahashi.
\newblock Sovereign Epistemic Commons under No-Meta Governance.
\newblock Zenodo, 2026. DOI: 10.5281/zenodo.18997828.

\bibitem{Angeli2002}
D.~Angeli.
\newblock A Lyapunov approach to incremental stability properties.
\newblock \emph{IEEE Transactions on Automatic Control}, 47(3):410--421, 2002. DOI: 10.1109/9.989067.

\bibitem{Angeli2009}
D.~Angeli.
\newblock Further results on incremental input-to-state stability.
\newblock \emph{IEEE Transactions on Automatic Control}, 54(6):1386--1391, 2009. DOI: 10.1109/TAC.2009.2015561.

\bibitem{Gu2019BadNets}
T.~Gu, K.~Liu, B.~Dolan-Gavitt, and S.~Garg.
\newblock BadNets: Evaluating backdooring attacks on deep neural networks.
\newblock \emph{IEEE Access}, 7:47230--47244, 2019. DOI: 10.1109/ACCESS.2019.2909068.

\bibitem{Li2024BackdoorSurvey}
Y.~Li, Y.~Jiang, Z.~Li, and S.-T. Xia.
\newblock Backdoor learning: A survey.
\newblock \emph{IEEE Transactions on Neural Networks and Learning Systems}, 35(1):5--22, 2024. DOI: 10.1109/TNNLS.2022.3182979.

\bibitem{Avizienis1985}
A.~Avizienis.
\newblock The N-version approach to fault-tolerant software.
\newblock \emph{IEEE Transactions on Software Engineering}, 11(12):1491--1501, 1985.

\bibitem{KnightLeveson1986}
J.~C. Knight and N.~G. Leveson.
\newblock An Experimental Evaluation of the Assumption of Independence in Multiversion Programming.
\newblock \emph{IEEE Transactions on Software Engineering}, 12(1):96--109, 1986.

\bibitem{AcharyaRobinson2011}
M.~Acharya and B.~Robinson.
\newblock Practical change impact analysis based on static program slicing for industrial software systems.
\newblock In \emph{Proceedings of the 33rd International Conference on Software Engineering (ICSE 2011)}, pages 746--755. ACM, 2011. DOI: 10.1145/1985793.1985898.

\bibitem{PalmskogCelikGligoric2020}
K.~Palmskog, A.~Celik, and M.~Gligoric.
\newblock Practical Machine-Checked Formalization of Change Impact Analysis.
\newblock In \emph{Tools and Algorithms for the Construction and Analysis of Systems (TACAS 2020)}, LNCS 12079, pages 137--157. Springer, 2020. DOI: 10.1007/978-3-030-45237-7\_9.

\bibitem{KuleszaTaskar2012}
A.~Kulesza and B.~Taskar.
\newblock Determinantal Point Processes for Machine Learning.
\newblock \emph{Foundations and Trends in Machine Learning}, 5(2--3):123--286, 2012. DOI: 10.1561/2200000044.


\bibitem{Pearl2009}
J.~Pearl.
\newblock \emph{Causality: Models, Reasoning, and Inference}.
\newblock 2nd edition, Cambridge University Press, 2009. ISBN: 978-0521895606.

\bibitem{JanzingSchoelkopf2010}
D.~Janzing and B.~Sch{\"o}lkopf.
\newblock Causal Inference Using the Algorithmic Markov Condition.
\newblock \emph{IEEE Transactions on Information Theory}, 56(10):5168--5194, 2010. DOI: 10.1109/TIT.2010.2060095.

\bibitem{Necula1997}
G.~C. Necula.
\newblock Proof-Carrying Code.
\newblock In \emph{Proceedings of the 24th ACM SIGPLAN-SIGACT Symposium on Principles of Programming Languages (POPL 1997)}, pages 106--119, 1997.

\bibitem{Appel2001}
A.~W. Appel.
\newblock Foundational Proof-Carrying Code.
\newblock In \emph{Proceedings of the 16th Annual IEEE Symposium on Logic in Computer Science (LICS 2001)}, pages 247--256, 2001. DOI: 10.1109/LICS.2001.932501.

\bibitem{Tishby1999}
N.~Tishby, F.~C. Pereira, and W.~Bialek.
\newblock The Information Bottleneck Method.
\newblock In \emph{Proceedings of the 37th Annual Allerton Conference on Communication, Control, and Computing}, 1999.

\bibitem{RATS2023}
H.~Birkholz, D.~Thaler, M.~Richardson, N.~Smith, and W.~Pan.
\newblock Remote ATtestation procedureS (RATS) Architecture.
\newblock RFC 9334, Internet Engineering Task Force, 2023.

\bibitem{ShepherdMarkantonakis2024}
C.~Shepherd and K.~Markantonakis.
\newblock \emph{Trusted Execution Environments}.
\newblock Springer, Cham, 2024. DOI: 10.1007/978-3-031-55561-9.

\bibitem{NIST90B2018}
M.~S\"onmez Turan, E.~Barker, J.~Kelsey, K.~McKay, M.~Baish, and M.~Boyle.
\newblock \emph{Recommendation for the Entropy Sources Used for Random Bit Generation}.
\newblock NIST Special Publication 800-90B, 2018. DOI: 10.6028/NIST.SP.800-90B.

\bibitem{Wilhelm2008WCET}
R.~Wilhelm, J.~Engblom, A.~Ermedahl, N.~Holsti, S.~Thesing, D.~B. Whalley, G.~Bernat, C.~Ferdinand, R.~Heckmann, T.~Mitra, F.~Mueller, I.~Puaut, P.~P. Puschner, J.~Staschulat, and P.~Stenstr\"om.
\newblock The worst-case execution-time problem---overview of methods and survey of tools.
\newblock \emph{ACM Transactions on Embedded Computing Systems}, 7(3):36:1--36:53, 2008. DOI: 10.1145/1347375.1347389.

\bibitem{Kocher1996Timing}
P.~C. Kocher.
\newblock Timing Attacks on Implementations of Diffie-Hellman, RSA, DSS, and Other Systems.
\newblock In \emph{Advances in Cryptology---CRYPTO '96}, LNCS 1109, pages 104--113. Springer, 1996. DOI: 10.1007/3-540-68697-5\_9.



\bibitem{Jungers2009}
R.~M. Jungers.
\newblock \emph{The Joint Spectral Radius: Theory and Applications}.
\newblock Lecture Notes in Control and Information Sciences 385, Springer, Berlin, Heidelberg, 2009. DOI: 10.1007/978-3-540-95980-9.

\bibitem{Williams1991}
D.~Williams.
\newblock \emph{Probability with Martingales}.
\newblock Cambridge Mathematical Textbooks, Cambridge University Press, 1991. DOI: 10.1017/CBO9780511813658.

\bibitem{CoverThomas2006}
T.~M. Cover and J.~A. Thomas.
\newblock \emph{Elements of Information Theory}.
\newblock 2nd edition, Wiley-Interscience, 2006. DOI: 10.1002/047174882X.

\bibitem{BarakEtAl2011LHL}
B.~Barak, Y.~Dodis, H.~Krawczyk, O.~Pereira, K.~Pietrzak, F.-X. Standaert, and Y.~Yu.
\newblock Leftover Hash Lemma, Revisited.
\newblock In P.~Rogaway, editor, \emph{Advances in Cryptology---CRYPTO 2011}, LNCS 6841, pages 1--20. Springer, 2011.

\bibitem{DiCosmoZacchiroli2017}
R.~Di Cosmo and S.~Zacchiroli.
\newblock Software Heritage: Why and How to Preserve Software Source Code.
\newblock In \emph{Proceedings of iPRES 2017: 14th International Conference on Digital Preservation}, Kyoto, Japan, 2017.

\end{thebibliography}


\end{document}
